---
title: "MatProt PRS Time-Varying Analysis"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection/reports/risk_scores")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"))) #setting up pathway

library(tidyverse)
library(conflicted)
library(survival)
library(survminer)

```


# **Overview**

MatProt PRS time-varying analysis. Specifically focused on checking proportional
hazards assumption in models for primary variable of interest (fdr_4level), as
well as potentially running a model that acounts for time-varying hazard and
understanding mom vs. dad during following.

Current approach:

+ cox.zph(coxme) - cox.zph doesn't check random effects for hazards
+ addition of interaction fdr_4level:log(time-to-event)
  + does not converge for T1D progression outcome
  
Other options:

+ *TODO: add from slides!*


**More details:**

From cox.zph documentation:

Random effects terms such a frailty or random effects in a coxme model are not
checked for proportional hazards, rather they are treated as a fixed offset in
model.

*TODO: not sure what this means for us? Will see what it looks like anyway -*
*Kirk says this should be okay.*


# **Load Data**

```{r}
# Model objects - coxme --------------------------------------------------------
#TODO: switch all these paths after re-run
mod_ia <- readRDS(
  "Maternal_Protection/data/results/famhist_prs/model_IA_fupIA_fdr_4level-PC1-PC2-sex-cc.rds")
mod_ia_grs2x <- readRDS(
  "Maternal_Protection/data/results/famhist_prs/model_IA_fupIA_fdr_4level-PC1-PC2-sex-cc-GRS2x.rds")

mod_t1d <- readRDS(
  "Maternal_Protection/data/results/famhist_prs/model_T1D_fupT1D_fdr_4level-PC1-PC2-sex-cc.rds")
mod_t1d_grs2x <- readRDS(
  "Maternal_Protection/data/results/famhist_prs/model_T1D_fupT1D_fdr_4level-PC1-PC2-sex-cc-GRS2x.rds")

mod_t1dprog <- readRDS(
  "Maternal_Protection/data/results/famhist_prs/coxme_T1D_prog_fdr_dad_fdr_4level-PC1-PC2-sex-cc-fupIA-mAA_at_IA.rds")
mod_t1dprog_grs2x <- readRDS(
  "Maternal_Protection/data/results/famhist_prs/coxme_T1D_prog_fdr_dad_fdr_4level-PC1-PC2-sex-cc-fupIA-mAA_at_IA-GRS2x.rds")

# Model objects - coxph with tt(fdr_4level) linear ----------------------------
#TODO: this currently doesn't differentiate which transform or which FDR was
# reference! Will fix once finalize what want to look at

# coxph_tt <- c("IA", "T1D", "T1D_prog")
# coxph_tt_files <- unlist(lapply(coxph_tt, function(x) {
#   list.files(
#     path = "Maternal_Protection/data/results/famhist_prs/",
#     pattern = paste0(
#       "coxph_tt_linear_fdr_4level_",
#       x
#     ),
#     full.names = TRUE
#   )
# }))
# 
# coxph_tt_files <- unique(grep("rds", coxph_tt_files, value = T))
# 
# # Read files
# coxph_tt_mod <- lapply(unique(grep("rds", coxph_tt_files, value = T)), readRDS)
# names(coxph_tt_mod) <- sapply(coxph_tt_files, function(file) {
#   base_name <- sub(
#     ".*coxph_tt_linear_fdr_4level_(.*?)_fdr.*",
#     "\\1",
#     file)
#   suffix <- if (grepl("GRS2x", file)) "_GRS2x" else ""
#   paste0(base_name, suffix)
# })


# Model objects - coxph with tt(fdr_4level) log ----------------------------
#TODO: this currently doesn't differentiate which transform or which FDR was
# reference! Will fix once finalize what want to look at

coxph_tt <- c("IA", "T1D", "T1D_prog")
coxph_tt_files <- unlist(lapply(coxph_tt, function(x) {
  list.files(
    path = "Maternal_Protection/data/results/famhist_prs/",
    pattern = paste0(
      "coxph_tt_log_fdr_4level_",
      x
    ),
    full.names = TRUE
  )
}))

coxph_tt_files <- unique(grep("rds", coxph_tt_files, value = T))

# Read files
coxph_tt_mod <- lapply(unique(grep("rds", coxph_tt_files, value = T)), readRDS)
names(coxph_tt_mod) <- sapply(coxph_tt_files, function(file) {
  base_name <- sub(
    ".*coxph_tt_log_fdr_4level_(.*?)_fdr.*",
    "\\1",
    file)
  suffix <- if (grepl("GRS2x", file)) "_GRS2x" else ""
  paste0(base_name, suffix)
})


# Pheno files ------------------------------------------------------------------
pheno_ia <- readRDS(
  "Maternal_Protection/data/teddy/famhist_prs/pheno_all_ia.rds")


# tt results linear and log dad as reference
linear <- read_csv(
  "Maternal_Protection/data/results/famhist_prs/coxph_tt_linear_fdr_4level_IA_T1D_T1D_prog_fdr_dad_ref_GRS2x_teddy_all.csv")
log <- read_csv(
  "Maternal_Protection/data/results/famhist_prs/coxph_tt_log_fdr_4level_IA_T1D_T1D_prog_fdr_dad_ref_GRS2x_teddy_all.csv")

```

# **Functions**

```{r}
show_cox_zph <- function(mod) {
  mod_cox_zph <- cox.zph(mod)
  print(mod_cox_zph)
  
  mod_vars <- grep("GLOBAL", rownames(mod_cox_zph$table), invert = T, value = T)
  
  for (var in mod_vars) {
    print(ggcoxzph(mod_cox_zph, var = c(var)))
  }
  
  # Return just p-values and columns
  table <- as.matrix(mod_cox_zph$table)
  p <- as.data.frame(t(table[, "p"]))
  return(p)
}

```



# **Summary cox.zph on coxme models**

## **IA**

### **No GRS2x**


```{r out.width = '50%', fig.show = 'hold'}

zph_ia <- show_cox_zph(mod_ia)

```

### **With GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_ia_grs2x <- show_cox_zph(mod_ia_grs2x)

```


## **T1D**

### **No GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1d <- show_cox_zph(mod_t1d)

```


### **With GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1d_grs2x <- show_cox_zph(mod_t1d_grs2x)

```



## **T1D Progression**

### **No GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1dprog <- show_cox_zph(mod_t1dprog)

```

### **With GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1dprog_grs2x <- show_cox_zph(mod_t1dprog_grs2x)

```

## **Summary - ZPH P-Values All Models**

```{r}

result <- bind_rows(
  zph_ia %>% dplyr::mutate(model = "IA"),
  zph_ia_grs2x %>% dplyr::mutate(model = "IA-GRS2x"),
  zph_t1d %>% dplyr::mutate(model = "T1D"),
  zph_t1d_grs2x %>% dplyr::mutate(model = "T1D-GRS2x"),
  zph_t1dprog %>% dplyr::mutate(model = "T1Dprog"),
  zph_t1dprog_grs2x %>% dplyr::mutate(model = "T1Dprog-GRS2x"),
)

result %>% relocate(model, GLOBAL)

```



# **Plot coxph models with tt(fdr_4level)**

Based on code from Kirk.

### Version for Dad as Reference

#### **No GRS2x**

```{r}
mod <- coxph_tt_mod$IA

# Extract coefficients and variance-covariance matrix from stratified model
coefs <- coef(mod)
vcov_mat <- vcov(mod)

# Create time sequence
times <- c(0.5, 1, 2, 5, 10, 15)
# times <- seq(0.5, 15, by = 0.1)

# Function to calculate HR and CI at each time point
calc_hr_ci <- function(beta_main, beta_tt, var_main, var_tt, cov_main_tt, times) {
  log_hr <- beta_main + beta_tt * log(times)
  se_log_hr <- sqrt(var_main + (log(times))^2 * var_tt + 2 * log(times) * cov_main_tt)
  
  data.frame(
    time = times,
    hr = exp(log_hr),
    lower = exp(log_hr - 1.96 * se_log_hr),
    upper = exp(log_hr + 1.96 * se_log_hr)
  )
}

# None vs. Dad
var_main <- "fdr_4levelNone"
var_tt <- "tt(fdr_4level)None"
none_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
none_hr$cohort <- "None vs Dad"

# Mom vs. Dad
var_main <- "fdr_4levelMom"
var_tt <- "tt(fdr_4level)Mom"
mom_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
mom_hr$cohort <- "Mom vs Dad"

# Sib vs. Dad
var_main <- "fdr_4levelSib"
var_tt <- "tt(fdr_4level)Sib"
sib_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
sib_hr$cohort <- "Sib vs Dad"

# Combine and plot
hr_data <- rbind(none_hr, mom_hr, sib_hr)

ggplot(hr_data, aes(x = time, y = hr, color = cohort, fill = cohort)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  scale_y_log10(breaks = c(0.5, 1, 2, 4, 8, 16)) +
  scale_x_continuous(breaks = c(0, 1, 2, 5, 10, 15)) +
  labs(
    x = "Follow-up Time (Years)",
    y = "Islet Autoimmunity Hazard Ratio",
    title = "Time-Varying IA Hazard Ratios for Birth Cohort Effects",
    subtitle = "Reference: Dad",
    caption = "Y-axis displayed on log scale",
    color = "Comparison",
    fill = "Comparison"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```





#### **With GRS2x**

```{r}
mod <- coxph_tt_mod$IA_GRS2x

# Extract coefficients and variance-covariance matrix from stratified model
coefs <- coef(mod)
vcov_mat <- vcov(mod)

# Create time sequence
times <- c(0.5, 1, 2, 5, 10, 15)
# times <- seq(0.5, 15, by = 0.1)

# Function to calculate HR and CI at each time point
calc_hr_ci <- function(beta_main, beta_tt, var_main, var_tt, cov_main_tt, times) {
  log_hr <- beta_main + beta_tt * log(times)
  se_log_hr <- sqrt(var_main + (log(times))^2 * var_tt + 2 * log(times) * cov_main_tt)
  
  data.frame(
    time = times,
    hr = exp(log_hr),
    lower = exp(log_hr - 1.96 * se_log_hr),
    upper = exp(log_hr + 1.96 * se_log_hr)
  )
}

# None vs. Dad
var_main <- "fdr_4levelNone"
var_tt <- "tt(fdr_4level)None"
none_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
none_hr$cohort <- "None vs Dad"

# Mom vs. Dad
var_main <- "fdr_4levelMom"
var_tt <- "tt(fdr_4level)Mom"
mom_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
mom_hr$cohort <- "Mom vs Dad"

# Sib vs. Dad
var_main <- "fdr_4levelSib"
var_tt <- "tt(fdr_4level)Sib"
sib_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
sib_hr$cohort <- "Sib vs Dad"

# Combine and plot
hr_data <- rbind(none_hr, mom_hr, sib_hr)

ggplot(hr_data, aes(x = time, y = hr, color = cohort, fill = cohort)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  scale_y_log10(breaks = c(0.5, 1, 2, 4, 8, 16)) +
  scale_x_continuous(breaks = c(0, 1, 2, 5, 10, 15)) +
  labs(
    x = "Follow-up Time (Years)",
    y = "Islet Autoimmunity Hazard Ratio",
    title = "Time-Varying IA Hazard Ratios for Birth Cohort Effects",
    subtitle = "Reference: Dad",
    caption = "Y-axis displayed on log scale",
    color = "Comparison",
    fill = "Comparison"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```






### Version for None as Reference

```{r}
mod <- coxph_tt_mod$IA

# Extract coefficients and variance-covariance matrix from stratified model
coefs <- coef(mod)
vcov_mat <- vcov(mod)

# Create time sequence
times <- c(0.5, 1, 2, 5, 10, 15)
# times <- seq(0.5, 15, by = 0.1)

# Function to calculate HR and CI at each time point
calc_hr_ci <- function(beta_main, beta_tt, var_main, var_tt, cov_main_tt, times) {
  log_hr <- beta_main + beta_tt * log(times)
  se_log_hr <- sqrt(var_main + (log(times))^2 * var_tt + 2 * log(times) * cov_main_tt)
  
  data.frame(
    time = times,
    hr = exp(log_hr),
    lower = exp(log_hr - 1.96 * se_log_hr),
    upper = exp(log_hr + 1.96 * se_log_hr)
  )
}

# Dad vs. None
var_main <- "fdr_4levelDad"
var_tt <- "tt(fdr_4level)Dad"
dad_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
dad_hr$cohort <- "Dad vs None"

# Mom vs. None
var_main <- "fdr_4levelMom"
var_tt <- "tt(fdr_4level)Mom"
mom_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
mom_hr$cohort <- "Mom vs None"

# Sib vs. None
var_main <- "fdr_4levelSib"
var_tt <- "tt(fdr_4level)Sib"
sib_hr <- calc_hr_ci(
  beta_main = coefs[var_main],
  beta_tt = coefs[var_tt],
  var_main = vcov_mat[var_main, var_main],
  var_tt = vcov_mat[var_tt, var_tt],
  cov_main_tt = vcov_mat[var_main, var_tt],
  times = times
)
sib_hr$cohort <- "Sib vs None"

# Combine and plot
hr_data <- rbind(dad_hr, mom_hr, sib_hr)

ggplot(hr_data, aes(x = time, y = hr, color = cohort, fill = cohort)) +
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, color = NA) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray40") +
  scale_y_log10(breaks = c(0.5, 1, 2, 4, 8, 16)) +
  scale_x_continuous(breaks = c(0, 1, 2, 5, 10, 15)) +
  labs(
    x = "Follow-up Time (Years)",
    y = "Islet Autoimmunity Hazard Ratio",
    title = "Time-Varying IA Hazard Ratios for Birth Cohort Effects",
    subtitle = "Reference: Dad",
    caption = "Y-axis displayed on log scale",
    color = "Comparison",
    fill = "Comparison"
  ) +
  theme_bw() +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  )
```




# **Quick summary linear vs log**

```{r, message = F}

merge <- full_join(
  linear %>% dplyr::mutate(transform = "linear"),
  log %>% dplyr::mutate(transform = "log")) %>%
  dplyr::select(
    transform, outcome, nobs, term, form, estimate_fdr_none, estimate_fdr_sib, estimate_fdr_mom,
    p.value_fdr_none, p.value_fdr_mom, p.value_fdr_sib) %>%
  pivot_longer(
    cols = matches("estimate_|p.value_"),
    names_to = c(".value", "fdr"),
    names_pattern = "(estimate|p.value)_(.*)") %>%
  dplyr::mutate(or = exp(estimate))

merge
write_tsv(merge, "Maternal_Protection/data/results/famhist_prs/TMP_coxph_tt_linear_log_summary.txt")

```





# **Session Info**

```{r session_info}
sessionInfo()

```

