---
title: "MatProt PRS Time-Varying Analysis"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: |
  (function(inputFile, encoding) {
    library(here)
    source(here("R", "study_specs.R"))
    rmarkdown::render(
      inputFile,
      encoding = encoding,
      output_dir = study_specs$teddy$report_out_dir
    )
  })
---

```{r setup, include=FALSE}
library(here)
source(here("config.R"))
source(here("R/study_specs.R"))
source(here("set_paper_colors.R"))
source(here("R/utils_coxph_tt.R"))

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = study_specs$teddy$result_out_dir)

library(tidyverse)
library(conflicted)
library(survival)
library(survminer)

```


# **Overview**

MatProt PRS time-varying analysis. Specifically focused on checking proportional
hazards assumption in models for primary variable of interest (fdr_4level), as
well as potentially running a model that accounts for time-varying hazard and
allows us to understand hazard over time in FDR Mom vs. FDR Dad during followup
time. Proportional hazards violation does not invalidate the model, but instead
changes the interpretation to average instead of instantaneous.

Approach:

1. Use cox.zph to check proportional hazards assumption on original primary
  coxme() models with kinship. Note that cox.zph doesn't check random effects
  for hazards.
2. For models where fdr_4level violates proportional hazards assumption (IA)
  add time-varying component for fdr_4level to model. 
  + Because doing this in  coxme() framework using an interaction between
    fdr_4level and time caused models not to converge, primary analysis was
    shifted to using coxph() with a cluster on family ID.
  + Also evaluated strata() for other violating covariates in some models, but
    ultimately decided not to include, keeping simpler models for parsimony.

**Final decision:** final MatProt PRS paper primary models will account for
time-varying hazard in fdr_4level using a linear transform of time, based on the
cox.zph residual plots and model AIC after exploration of many time transform
options.

*More details:*

From cox.zph documentation: Random effects terms such a frailty or random
effects in a coxme() model are not checked for proportional hazards, rather they
are treated as a fixed offset in model.


# **Load Data**

```{r}
fdr_order <- c("Father", "Mother", "Sibling", "None")
outcome_order <- c("IA", "Progression", "T1D")

#####  coxme model objects -----------------------------------------------------
### IA
mc_ia <- readRDS(
  "model_objects/coxme_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc_teddy_all.rds")
mc_ia_g <- readRDS(
  "model_objects/coxme_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-GRS2x_teddy_all.rds")

### T1D
mc_t1d <- readRDS(
  "model_objects/coxme_T1D_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc_teddy_all.rds")
mc_t1d_g <- readRDS(
  "model_objects/coxme_T1D_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-GRS2x_teddy_all.rds")

### T1D prog
mc_t1dprog <- readRDS(
  "model_objects/coxme_T1D_prog_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-fupIA-mAA_at_IA_teddy_all.rds")
mc_t1dprog_g <- readRDS(
  "model_objects/coxme_T1D_prog_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-fupIA-mAA_at_IA-GRS2x_teddy_all.rds")


#####  coxph model objects -----------------------------------------------------
### IA
# No time-varying
m_ia <- readRDS(
  "model_objects/coxph_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_g <- readRDS(
  "model_objects/coxph_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
# Linear time transform
m_ia_lin <- readRDS(
  "model_objects/coxph_tt_linear_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_lin_g <- readRDS(
  "model_objects/coxph_tt_linear_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
# Log time transform
m_ia_log <- readRDS(
  "model_objects/coxph_tt_log_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_log_g <- readRDS(
  "model_objects/coxph_tt_log_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
# Spline time transform, with knots informed by PMID 25883970
m_ia_spline_3 <- readRDS(
  "model_objects/coxph_tt_spline_3knots_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_spline_3_g <- readRDS(
  "model_objects/coxph_tt_spline_3knots_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
# Spline time transform, with knots informed by PMID 25883970
m_ia_spline_5 <- readRDS(
  "model_objects/coxph_tt_spline_5knots_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_spline_5_g <- readRDS(
  "model_objects/coxph_tt_spline_5knots_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
# Spline time transform, with knots auto-determined (equally spaced quantiles)
m_ia_spline_d2 <- readRDS(
  "model_objects/coxph_tt_spline_2df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_spline_d2_g <- readRDS(
  "model_objects/coxph_tt_spline_2df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
m_ia_spline_d3 <- readRDS(
  "model_objects/coxph_tt_spline_3df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_spline_d3_g <- readRDS(
  "model_objects/coxph_tt_spline_3df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
m_ia_spline_d4 <- readRDS(
  "model_objects/coxph_tt_spline_4df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_spline_d4_g <- readRDS(
  "model_objects/coxph_tt_spline_4df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
m_ia_spline_d6 <- readRDS(
  "model_objects/coxph_tt_spline_6df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_spline_d6_g <- readRDS(
  "model_objects/coxph_tt_spline_6df_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
# Spline time transform, with single knot chosen based on knowledge of data
# and prior literature
m_ia_spline_1 <- readRDS(
  "model_objects/coxph_tt_spline_1knots_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_ia_spline_1_g <- readRDS(
  "model_objects/coxph_tt_spline_1knots_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")

### T1D
m_t1d <- readRDS(
  "model_objects/coxph_T1D_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m_t1d_g <- readRDS(
  "model_objects/coxph_T1D_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
m_t1d_strat <- readRDS(
  "model_objects/coxph_T1D_fdr_dad_ref_fdr_4level-PC1-PC2-strata(sex)-strata(cc)-cluster(FID)_teddy_all.rds")
m_t1d_g_strat <- readRDS(
  "model_objects/coxph_T1D_fdr_dad_ref_fdr_4level-PC1-PC2-strata(sex)-strata(cc)-cluster(FID)-GRS2x_teddy_all.rds")

### T1D prog
m_t1dprog <- readRDS(
  "model_objects/coxph_T1D_prog_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-fupIA-mAA_at_IA_teddy_all.rds")
m_t1dprog_g <- readRDS(
  "model_objects/coxph_T1D_prog_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-fupIA-mAA_at_IA-GRS2x_teddy_all.rds")


#####  model result files ------------------------------------------------------

# Only loading subset of files compared to all models above, used to compare
# model effect estimates across versions

### IA
ia <- read_csv("coxph_IA_fdr_dad_ref_GRS2x_teddy_all.csv") %>%
  dplyr::mutate(model = paste0(outcome, ifelse(is.na(term), "", paste0("-", term))))
ia_lin <- read_csv("coxph_tt_linear_fdr_4level_IA_fdr_dad_ref_GRS2x_teddy_all.csv") %>%
  dplyr::mutate(model = paste0(outcome, ifelse(is.na(term), "", paste0("-", term)), "-lin"))
ia_log <- read_csv("coxph_tt_log_fdr_4level_IA_fdr_dad_ref_GRS2x_teddy_all.csv") %>%
  dplyr::mutate(model = paste0(outcome, ifelse(is.na(term), "", paste0("-", term)), "-log"))
ia_spline_3knots <- read_csv("coxph_tt_spline_3knots_fdr_4level_IA_fdr_dad_ref_GRS2x_teddy_all.csv") %>%
  dplyr::mutate(model = paste0(outcome, ifelse(is.na(term), "", paste0("-", term)), "-spline-3knots"))
ia_merge <- bind_rows(
  ia, ia_lin, ia_log, ia_spline_3knots) %>%
  dplyr::relocate(model) %>%
  distinct(across(-model), .keep_all = TRUE)

### T1D
t1d <- read_csv("coxph_T1D_fdr_dad_ref_GRS2x_teddy_all.csv") %>%
  dplyr::mutate(model = paste0(outcome, ifelse(is.na(term), "", paste0("-", term))))
t1d_strat <- read_csv("coxph_T1D_fdr_dad_ref_GRS2x_strata_teddy_all.csv") %>%
  dplyr::mutate(
    model = paste0(
      outcome, ifelse(is.na(term), "", paste0("-", term)), "-strat"))
t1d_merge <- bind_rows(
  t1d, t1d_strat) %>%
  dplyr::relocate(model) %>%
  distinct(across(-model), .keep_all = TRUE)

### T1D Prog
prog_merge <- read_csv("coxph_T1D_prog_fdr_dad_ref_GRS2x_teddy_all.csv") %>%
  dplyr::mutate(model = paste0(outcome, ifelse(is.na(term), "", paste0("-", term))))


#####  pheno files -------------------------------------------------------------
ia <- readRDS(paste0(study_specs$teddy$intermed_out_dir, "/pheno_all_ia.rds"))

```

# **Functions**

```{r}
show_cox_zph <- function(mod) {
  mod_cox_zph <- cox.zph(mod)
  print(mod_cox_zph)
  
  mod_vars <- grep("GLOBAL", rownames(mod_cox_zph$table), invert = T, value = T)
  
  for (var in mod_vars) {
    print(ggcoxzph(mod_cox_zph, var = c(var)))
  }
  
  # Return just p-values and columns
  table <- as.matrix(mod_cox_zph$table)
  p <- as.data.frame(t(table[, "p"]))
  return(p)
}

# data <- m2_dad
# ref <- "dad"
# term <- "GRS2x"
make_h2_forest_plot <- function(data, ref) {
  
  fdr_all <- c("dad", "mom", "sib", "none")
  fdr <- setdiff(fdr_all, ref)
  
  cols_to_select <- c(
    "study", "model", "outcome", "model_group",
    paste0("estimate_fdr_", fdr),
    paste0("conf.low_fdr_", fdr),
    paste0("conf.high_fdr_", fdr))
  
  data_plot <- data %>%
    select(all_of(cols_to_select)) %>%
    rename(model_type = model) %>%
    pivot_longer(
      cols = matches("estimate_|conf\\.low_|conf\\.high_"),
      names_to = c(".value", "term"),
      names_pattern = "(estimate|conf\\.low|conf\\.high)_(.*)") %>%
    mutate(term = recode(
      term,
      fdr_mom = "Mother",
      fdr_sib = "Sibling",
      fdr_dad = "Father",
      fdr_none = "None")) %>%
    mutate(
      or = exp(estimate),
      or_conf_low = exp(conf.low),
      or_conf_high = exp(conf.high)) %>%
    mutate(term = factor(term, levels = rev(fdr_order))) %>%
    mutate(outcome = case_when(
      outcome == "case_t1dprog" ~ "Progression",
      outcome == "ia" ~ "IA",
      outcome == "persist_conf_ab" ~ "IA",
      outcome == "t1d" ~ "T1D")) %>%
    mutate(outcome = factor(outcome, levels = outcome_order)) %>%
    mutate(
      model_type_ind = ifelse(grepl("GRS2x", model_type), "With GRS2x", "Without GRS2x"),
      model_type_ind = factor(model_type_ind, levels = c("With GRS2x", "Without GRS2x"))
    )
  
  plot <- ggplot(
    data_plot,
    aes(x = term, y = or, ymin = or_conf_low, ymax = or_conf_high,
        color = term, fill = term, alpha = model_type_ind)
  ) + 
    geom_linerange(position = position_dodge(width = 0.5), show.legend = FALSE, size = 0.8) +
    geom_hline(yintercept = 1, lty = 2) +
    geom_point(size = 2.5, shape = 21, colour = "white", stroke = 0.5,
               position = position_dodge(width = 0.5)) +
    scale_color_manual(values = c(
      "Father" = color_fdr_dad,
      "Mother" = color_fdr_mom,
      "Sibling" = color_fdr_sib,
      "None" = color_fdr_none),
      guide = "none") +
    scale_fill_manual(values = c(
      "Father" = color_fdr_dad,
      "Mother" = color_fdr_mom,
      "Sibling" = color_fdr_sib,
      "None" = color_fdr_none),
      guide = "none") +
    scale_alpha_manual(
      values = c("Without GRS2x" = 0.4, "With GRS2x" = 1)) +
    guides(
      alpha = guide_legend(
        reverse = T,
        override.aes = list(
          shape = 21,
          size  = 3,
          fill  = "grey50",
          color = "grey50"))) +
    coord_flip() +
    theme_bw() +
    labs(
      x = "",
      y = "Hazard Ratio"
    ) +
    theme(
      legend.position = "bottom",
      strip.background = element_rect(fill = "white", color = "black"),
      axis.text.y = element_text(angle = 90, hjust = 0.5)
    ) +
    facet_wrap(~ model_group, ncol = 1, strip.position = "left")
  
  return(plot)
}

```



# **Summary PH Violations coxme Models**

## **IA**


### **No GRS2x**


```{r out.width = '50%', fig.show = 'hold'}

zph_ia <- show_cox_zph(m_ia)

```

### **With GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_ia_g <- show_cox_zph(m_ia_g)

```

### **Models with tt()**

Can't run cox.zph on models with tt(). Instead, use AIC to compare.

Would be better to do an Anova with LRT on the nested models, but can't do that
because of the robust variance due to the cluster(FID) in the models. Instead
compare all with AIC.

Note that can't compare models with strata() to models without because a
different partial likelihood is used (computed within strata).

**Look at all, stacked so with/without GRS2x values together:**

```{r}
AIC(
  m_ia, m_ia_g, m_ia_lin, m_ia_lin_g, m_ia_log, m_ia_log_g,
  m_ia_spline_3, m_ia_spline_3_g, m_ia_spline_5, m_ia_spline_5_g,
  m_ia_spline_d2, m_ia_spline_d2_g, m_ia_spline_d3, m_ia_spline_d3_g,
  m_ia_spline_d4, m_ia_spline_d4_g, m_ia_spline_d6, m_ia_spline_d6_g)
```

```{r}
AIC(m_ia_spline_1, m_ia_spline_1_g)
```



In general, GRS2x improves model fit and linear time has slightly lower AIC
than log time.

**Look at all models without GRS2x:**

```{r}
AIC(m_ia, m_ia_lin, m_ia_log, m_ia_spline)
```

All have fairly similar AIC, best is linear, then log, then spline, then no
time interaction with fdr_4level.

**Look at all models with GRS2x:**

```{r}
AIC(m_ia_g, m_ia_lin_g, m_ia_log_g, m_ia_spline_g)
```

**Look at knot ages:**

```{r}
m_ia_spline_3$tt_meta$knots
m_ia_spline_5$tt_meta$knots
m_ia_spline_3$tt_meta$boundary

```



## **T1D**

### **No GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1d <- show_cox_zph(m_t1d)

```


### **With GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1d_g <- show_cox_zph(m_t1d_g)

```


### **No GRS2x, With Strata**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1d_strat <- show_cox_zph(m_t1d_strat)

```


### **With GRS2x, With Strata**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1d_g_strat <- show_cox_zph(m_t1d_g_strat)

```


Adding the strata resolve PH violation in PC1. 


## **T1D Progression**

### **No GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1dprog <- show_cox_zph(m_t1dprog)

```

### **With GRS2x**

```{r out.width = '50%', fig.show = 'hold'}

zph_t1dprog_g <- show_cox_zph(m_t1dprog_g)

```

## **Summary - ZPH P-Values All Models**

```{r}

result <- bind_rows(
  zph_ia %>% dplyr::mutate(model = "IA"),
  zph_ia_g %>% dplyr::mutate(model = "IA-GRS2x"),
  zph_t1d %>% dplyr::mutate(model = "T1D"),
  zph_t1d_strat %>% dplyr::mutate(model = "T1D-Strat"),
  zph_t1d_g %>% dplyr::mutate(model = "T1D-GRS2x"),
  zph_t1d_g_strat %>% dplyr::mutate(model = "T1D-GRS2x-Strat"),
  zph_t1dprog %>% dplyr::mutate(model = "T1Dprog"),
  zph_t1dprog_g %>% dplyr::mutate(model = "T1Dprog-GRS2x"),
)

result <- result %>% relocate(model, GLOBAL)
result

write_tsv(result, "summary_coxzph.txt")

```



# **Main Outcomes Forest Plots of Model Versions**

### **IA**

```{r}
ia_merge <- ia_merge %>%
  dplyr::mutate(model = factor(model), levels = c(
    "IA", "IA-GRS2x",
    "IA-lin", "IA-GRS2x-lin",
    "IA-log", "IA-GRS2x-log", 
    "IA-spline-3knots", "IA-GRS2x-spline-3knots")) %>%
  dplyr::mutate(model_group = case_when(
    model == "IA" | model == "IA-GRS2x" ~ "IA",
    model == "IA-lin" | model == "IA-GRS2x-lin" ~ "IA-lin",
    model == "IA-log" | model == "IA-GRS2x-log" ~ "IA-log",
    model == "IA-spline-3knots" | model == "IA-GRS2x-spline-3knots" ~ "IA-spline-3knots"
  ))


make_h2_forest_plot(
  data = ia_merge,
  ref = "dad"
)
```


### **T1D**

```{r}
t1d_merge <- t1d_merge %>%
  dplyr::mutate(model = factor(model), levels = c(
    "T1D", "T1D-GRS2x",
    "T1D-strat", "T1D-GRS2x-strat")) %>%
  dplyr::mutate(model_group = case_when(
    model == "T1D" | model == "T1D-GRS2x" ~ "T1D",
    model == "T1D-strat" | model == "T1D-GRS2x-strat" ~ "T1D-strat"
  ))


make_h2_forest_plot(
  data = t1d_merge,
  ref = "dad"
)
```


### **T1D Prog**

```{r}
prog_merge <- prog_merge %>%
  dplyr::mutate(model = factor(model), levels = c(
    "T1D", "T1D-GRS2x")) %>%
  dplyr::mutate(model_group = case_when(
    model == "T1D" | model == "T1D-GRS2x" ~ "T1D",
  ))


make_h2_forest_plot(
  data = prog_merge,
  ref = "dad"
)
```

# **Plot IA Hazard Over Time**

Based on code originally from Kirk.

### **Linear**

```{r}
lin_tt <- get_tt_hr(
  mod = m_ia_lin,
  var = "fdr_4level",
  levels = c("None", "Mom", "Sib"),
  ref_label = "Dad",
  time_transform = identity)

plot_tt_hr(
  data = lin_tt,
  var = "fdr_4level",
  ref_label = "Dad")

```

```{r}
lin_g_tt <- get_tt_hr(
  mod = m_ia_lin_g,
  var = "fdr_4level",
  levels = c("None", "Mom", "Sib"),
  ref_label = "Dad",
  time_transform = identity)

plot_tt_hr(
  data = lin_g_tt,
  var = "fdr_4level",
  ref_label = "Dad")
```


### **Log**

```{r}
log_tt <- get_tt_hr(
  mod = m_ia_log,
  var = "fdr_4level",
  levels = c("None", "Mom", "Sib"),
  ref_label = "Dad",
  time_transform = log)

plot_tt_hr(
  data = log_tt,
  var = "fdr_4level",
  ref_label = "Dad")

```

```{r}
log_g_tt <- get_tt_hr(
  mod = m_ia_log_g,
  var = "fdr_4level",
  levels = c("None", "Mom", "Sib"),
  ref_label = "Dad",
  time_transform = log)

plot_tt_hr(
  data = log_g_tt,
  var = "fdr_4level",
  ref_label = "Dad")
```



### **Spline**

*Current code only shows version with three pre-defined knots, but same*
*plotting code can be used for all versions.*

```{r}
spline_tt <- get_tt_hr_spline(
  mod = m_ia_spline_3,
  var = "fdr_4level",
  levels = c("None", "Mom", "Sib"),
  ref_label = "Dad",
  knots = m_ia_spline_3$tt_meta$knots,
  boundary = m_ia_spline_3$tt_meta$boundary)

plot_tt_hr(
  data = spline_tt,
  var = "fdr_4level",
  ref_label = "Dad")

```

```{r}
spline_tt_g <- get_tt_hr_spline(
  mod = m_ia_spline_3_g,
  var = "fdr_4level",
  levels = c("None", "Mom", "Sib"),
  ref_label = "Dad",
  knots = m_ia_spline_3_g$tt_meta$knots,
  boundary = m_ia_spline_3_g$tt_meta$boundary)

plot_tt_hr(
  data = spline_tt_g,
  var = "fdr_4level",
  ref_label = "Dad")

```


# **Cross-Sectional Check**

Used the above HRs over time to check the last age at which T1D mom was
protective. Want to internally verify this by doing a left-censored analysis:
if start follow up at age 6, expect to see no protective effect.

```{r}
ia_cens <- ia %>%
  dplyr::filter(fupIA >= 5)
```



# **Session Info**

```{r session_info}
sessionInfo()

```
