---
title: "Manuscript_Famhist_PRS"
author: "NV, Updated by SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: |
  (function(inputFile, encoding) {
    library(here)
    source(here("R", "study_specs.R"))
    rmarkdown::render(
      inputFile,
      encoding = encoding,
      output_dir = study_specs$teddy$report_out_dir
    )
  })
---

```{r setup, include=FALSE}
library(here)
devtools::load_all()
source(here("config.R"))

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = study_specs$teddy$result_out_dir)

library(tidyverse)
library(conflicted)
library(table1)
library(ggpubr)
library(cowplot)
library(survminer)
library(ggpattern)
library(broom)
library(rstatix)

```

# **Overview**

This report generates all figures, tables, and text-referenced values for the
MatProt PRS manuscript.

*TO NOTE:* previous versions of this report included both DAISY & TEDDY, as well
as multiple data versions (HLA-DR filtering, FDR none as reference for modeling,
Non-HLA GRS2x component only analyses, and sex-stratified results.). The current
report has been cleaned up to only include the final version presented in the
manuscript. The results for all are still saved at
Maternal_Protection/data/results/famhist_prs.

# **Load Data**


```{r, message = F}
#TODO: move these to config?
fdr_order <- c("Father", "Mother", "Sibling", "None")
outcome_order <- c("T1D", "IA", "Progression")

# From step 01 - read in pheno files
t_all <- readRDS(
  paste0(study_specs$teddy$intermed_out_dir, "/pheno_all_ia.rds"))
  # same as pheno_all_t1d.rds
t_all_prog <- readRDS(
  paste0(study_specs$teddy$intermed_out_dir, "/pheno_all_prog.rds"))
t_ctrls_ia <- readRDS(
  paste0(study_specs$teddy$intermed_out_dir, "/pheno_ctrls_ia.rds"))

# From step 02 - read in hypoth 1 results
m1_dad <- read_csv(
  "lm_GRS2x_fdr_dad_ref_teddy_all.csv")
m_ia_lin_g <- readRDS(
  "model_objects/coxph_tt_linear_fdr_4level_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")

# From step 03 - read in hypoth 2 results
# Primary results - coxph
m2_dad_ia <- read_csv(
  "coxph_tt_linear_IA_fdr_dad_ref_GRS2x_teddy_all.csv")
m2_dad_t1d_prog <- read_csv(
  "coxph_T1D_T1D_prog_fdr_dad_ref_GRS2x_teddy_all.csv")
# Supplementary results - coxph sex-stratified
m2_female_ia <- read_csv("coxph_tt_linear_IA_fdr_dad_ref_GRS2x_teddy_female.csv")
m2_female_t1d_prog <- read_csv("coxph_T1D_T1D_prog_fdr_dad_ref_GRS2x_teddy_female.csv")
m2_male_ia <- read_csv("coxph_tt_linear_IA_fdr_dad_ref_GRS2x_teddy_male.csv")
m2_male_t1d_prog <- read_csv("coxph_T1D_T1D_prog_fdr_dad_ref_GRS2x_teddy_male.csv")
# CoxPH violation results - coxph no time-varying
# No time-varying
m2_ia_ph <- readRDS(
  "model_objects/coxph_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)_teddy_all.rds")
m2_ia_g_ph <- readRDS(
  "model_objects/coxph_IA_fdr_dad_ref_fdr_4level-PC1-PC2-sex-cc-cluster(FID)-GRS2x_teddy_all.rds")
# Coxph results with interaction FDR*sex
m2_ia_intxn <- read_csv(
  "coxph_tt_linear_IA_fdr_dad_ref_GRS2x_teddy_all_intxn.csv")
m2_t1d_prog_intxn <- read_csv(
  "coxph_T1D_T1D_prog_fdr_dad_ref_GRS2x_teddy_all_intxn.csv")

# From TEDDY ancestry estimation
pheno_raw <- read_tsv(study_specs$teddy$pheno_raw_path)
gmm_pop <- read_delim(study_specs$teddy$ancestry_path)

# From T1D mom & T1D dad allele freq comparison
risk_score_dir <- dirname(study_specs$teddy$grs2_path)
af_fam_grs2x <- read_delim(
  paste0(risk_score_dir, "/afreq_by_fdr_grs2x.txt"),
  col_types = cols(.default = col_character()))
af_fam_grs2x_test_m <- read.table(
  paste0(risk_score_dir, "/t1d_dad_mom_pheno_assoc_result.assoc"),
         header = T)
af_fam_grs2x_test_s <- read.table(
  paste0(risk_score_dir, "/t1d_dad_sib_pheno_assoc_result.assoc"),
         header = T)

# From GRS2x run
log_grs2x <- read_delim(
  paste0(risk_score_dir, "/score_results_with_imp_log_summary.txt"),
  col_types = cols(.default = col_character()))

# From TOPMed imputation output
tm_info <- read_tsv(
  paste0(risk_score_dir, "/chr_all_concat_grs2x.pvar"),
  comment = "##")

```


# **Figures**
## **Figure 1**

### **Fig1AB**

Boxplot of GRS2x scores and forest plot of model 1 results.


```{r}
# Prepping here for supplementary figure 1 as well
pheno_all <- t_all %>%
  dplyr::select(fdr_4level, GRS2x, Non_HLA, IA, sex) %>%
  dplyr::mutate(study = "TEDDY") %>%
  dplyr::mutate(fdr_4level = recode(
    fdr_4level,
    Mom = "Mother",
    Sib = "Sibling",
    Dad = "Father",
    None = "None")) %>%
    dplyr::mutate(fdr_4level = factor(fdr_4level, levels = fdr_order))

pheno_ctrls <- t_ctrls_ia %>%
  dplyr::select(fdr_4level, GRS2x, Non_HLA, IA, sex) %>%
  dplyr::mutate(study = "TEDDY") %>%
  dplyr::mutate(fdr_4level = recode(
    fdr_4level,
    Mom = "Mother",
    Sib = "Sibling",
    Dad = "Father",
    None = "None")) %>%
  dplyr::mutate(fdr_4level = factor(fdr_4level, levels = fdr_order))

# Set plot scales
grs2x_min <- 
  floor(min(pheno_all$GRS2x, pheno_ctrls$GRS2x) - 0.5)
  # -0.5 to allow for group size labels
grs2x_max <- 
  ceiling(max(pheno_all$GRS2x, pheno_ctrls$GRS2x) + 2)
  # +2 to allow for significance bars
```

ANOVA for F1 A:

```{r out.width = '50%', fig.show = 'hold'}
# ANOVA for F1 A - use base R to run and check assumptions, but plot using
# ggpubr/rstatix
anova_f1_a_check <- aov(GRS2x ~ fdr_4level, data = pheno_all)
print(summary(anova_f1_a_check))
plot(anova_f1_a_check)  # check residual assumptions
car::leveneTest(GRS2x ~ fdr_4level, data = pheno_all)  # check variances

# For plotting
anova_f1_a <- anova_test(GRS2x ~ fdr_4level, data = pheno_all)

```

Tukey's HSD for F1 A:

```{r}
tukey_f1_a <- pheno_all %>%
  tukey_hsd(GRS2x ~ fdr_4level) %>%
  add_xy_position(x = "fdr_4level") %>%
  dplyr::mutate(
    p.adj.label = ifelse(
      p.adj < 0.01,
      formatC(p.adj, format = "e", digits = 1),
      formatC(p.adj, format = "f", digits = 3)))
print(tukey_f1_a)

```



```{r}
f1_a <- ggplot(pheno_all, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(x = "Affected Relative") +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(values = c(
    "None" = color_fdr_none, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib)) +
  stat_pvalue_manual(
    tukey_f1_a, tip.length = 0.01, hide.ns = TRUE,
    label = "P-adj={p.adj.label}", step.increase = 0.02) +
  labs(subtitle = paste0("ANOVA, P=", anova_f1_a$p))

```

```{r}
wald_f1_b_p <- formatC(m1_dad$fdr_global_p, format = "e", digits = 1)
f1_dad <- make_h1_forest_plot(m1_dad, "dad", fdr_order)
f1_dad <- f1_dad +
  labs(subtitle =  paste0("Global Wald, P=", wald_f1_b_p))

```

```{r}
# Save combined figure
f1 <- ggarrange(
  f1_a,
  f1_dad,
  labels = c("A", "B"), nrow=1, widths = c(.5, .5))

f1

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/fig_1.pdf"),
  f1, units = "in",  width = 7, height = 3, bg = "white")
```


### **Data as Table for Text**

All results from hypothesis 1. Read in results here, reformat, and display here
to be written into manuscript draft (no table).

```{r}
m1_exp <- m1_dad %>%
  dplyr::mutate(model_type = "linear") %>%
  dplyr::relocate(model_type, .after = nobs) %>%
  dplyr::select(-which_fdr, -study)

fdr <- c("mom", "sib", "none")
cols_to_select <- c(
  "nobs", "model_type", "form",
  paste0("estimate_fdr_", fdr),
  paste0("conf.low_fdr_", fdr),
  paste0("conf.high_fdr_", fdr),
  paste0("z.value_fdr_", fdr),
  paste0("p.value_fdr_", fdr),
  paste0("std.error_fdr_", fdr))

m1_exp_mod <- m1_exp %>%
  select(all_of(cols_to_select)) %>%
  pivot_longer(
    cols = contains("_fdr_"),
    names_to = c(".value", "fdr"),
    names_pattern = 
      "(estimate|conf\\.low|conf\\.high|z\\.value|p\\.value|std\\.error)_fdr_(mom|sib|none)") %>%
  dplyr::mutate(fdr = case_when(
    fdr == "mom" ~ "Mother",
    fdr == "sib" ~ "Sibling",
    fdr == "none" ~ "None")) %>%
  dplyr::mutate(fdr = factor(fdr, levels = rev(fdr_order)))

# Write Excel file
m1_exp_mod
```



### **Simplified for Graphical Abstract**

```{r fig.width=3, fig.height=2.2}
fdr_order_abs <- c("None", "Sibling", "Father", "Mother")

pheno_all_alt <- pheno_all %>%
  dplyr::mutate(fdr_4level = factor(fdr_4level, levels = rev(fdr_order_abs)))

f1_a_abs <- ggplot(pheno_all_alt, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(x = "Affected Relative") +
  theme_classic() +
  theme(axis.ticks = element_blank(),
        legend.title = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_x_discrete(
    labels = c("None" = "Unaffected")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  scale_fill_manual(
    values = c(
      "None" = color_fdr_none_abs, "Father" = color_fdr_dad,
      "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib)) +
  coord_flip()

f1_a_abs

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/fig_1_a_graph_abstract.png"),
  f1_a_abs, units = "in",  width = 3, height = 2.2, bg = "white")
```


### **Distribution for Graphical Abstract**

Use random normal distributions to show:

1. False father distribution
2. False hypothesized left-shifted mother distribution, if remove highest GRS2x scores
2. False mother distribution, identical to father but with line at high GRS2x

```{r fig.width=3, fig.height=2.2}

data_rand_1 <- data.frame(norm = rnorm(n = 100000, mean = 0, sd = 1))

data_rand <- bind_rows(
  data_rand_1 %>%
    dplyr::mutate(
      fdr_4level = "None",
      norm = norm - 0.75),
  data_rand_1 %>%
    dplyr::mutate(
      fdr_4level = "Sibling"),
  data_rand_1 %>%
    dplyr::mutate(
      fdr_4level = "Father"),
  data_rand_1 %>%
    dplyr::mutate(
      fdr_4level = "Mother")) %>%
  dplyr::mutate(
    fdr_4level = factor(fdr_4level, levels = c(fdr_order_abs)),
    line_val = 0.3*max(norm))

dist_hypoth_abs <- ggplot(data_rand, aes(x = norm, fill = fdr_4level)) +
  geom_density(alpha = 0.75) +
  geom_vline(
    data = data_rand %>%
      dplyr::filter(fdr_4level %in% c("Mother")),
    aes(xintercept = line_val),
    linewidth = 0.5, linetype = "dashed") +
  geom_area(
    data = data_rand %>% dplyr::filter(fdr_4level == "Mother"),
    stat = "density", 
    aes(
      x = stage(norm, after_stat = scales::oob_censor(x, c(unique(data_rand$line_val), Inf))),
      y = after_stat(density)),
    fill = "black", # Change to whatever highlight color you want
    alpha = 0.5) +
  facet_grid(rows = vars(fdr_4level), scales = "free_y") +
  theme_classic() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank(),
    axis.line.x = element_blank(),
    axis.ticks = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.line.y = element_blank(),
    legend.position = "none",
    strip.background = element_blank(),
    strip.text = element_blank()) +
  scale_fill_manual(values = c(
    "None" = color_fdr_none_abs, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib))

dist_hypoth_abs

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/dist_hypoth_graph_abstract.png"),
  dist_hypoth_abs, units = "in",  width = 3, height = 2.2, bg = "white")

```


## **Figure 2**

Forest plot of hypothesis 2 model results.

### Fig2A

```{r}
m2_merge <- dplyr::bind_rows(
  m2_dad_t1d_prog,
  m2_dad_ia) %>%
  dplyr::mutate(outcome = case_when(
    str_detect(form, "T1D,") ~ "T1D",
    str_detect(form, "fupT1D_prog") ~ "Progression",
    TRUE ~ outcome)) %>%
  dplyr::select(-which_fdr, -study)


f2_a <- make_h2_forest_plot_manuscript(
  m2_merge, "dad", "GRS2x", fdr_order, outcome_order)

```


### Fig2B

HR over time plot for IA model with GRS2x adjustment.

```{r}
lin_tt_g <- get_tt_hr(
  mod = m_ia_lin_g,
  var = "fdr_4level",
  levels = c("None", "Mom", "Sib"),
  ref_label = "Dad",
  time_transform = identity,
  times = seq(0, 15, 1))

# Use same pheno as for IA (t_all)
# lin_tt_g_n <- get_n_remain(
#   data = t_all,
#   var = "fdr_4level",
#   time_var = "fupIA")

# Option: can add data_n = ling_tt_g_n to get sample size at each time point
f2_b <- plot_tt_hr_manuscript(
  data = lin_tt_g,
  var = "fdr_4level",
  ref_label = "Dad",
  fdr_order = fdr_order,
  times = seq(0, 15, 1),
  add_last_matprot = T) +
  labs(x = "Age (Years)")

```


```{r}
# Save combined figure
f2 <- plot_grid(
  f2_a + theme(
  legend.text = element_text(margin = margin(l = 4))),
  f2_b + theme(
  legend.text = element_text(margin = margin(l = 4))),
  ncol = 1,
  align = "v",
  axis = "lr", 
  labels = c("A", "B"))

f2

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/fig_2.pdf"),
  f2, units = "in",  width = 7, height = 5, bg = "white")

```

### **Simplified for Graphical Abstract**

Figure 2A:

```{r fig.width=3, fig.height=2.2}
m2_merge_ia <- m2_merge %>%
  dplyr::filter(outcome == "IA")

f2_a_abs <- make_h2_forest_plot_manuscript(
  m2_merge_ia, "dad", "GRS2x", fdr_order_abs, outcome_order) +
  geom_point(size=3, shape=21, colour="white", stroke = 0.5,
           position=position_dodge(width = 0.5)) +
  theme_classic() +  # to match rest of graphical abstract themes
  theme(
    legend.position = "bottom",
    legend.direction = "vertical",
    axis.title.x = element_blank(),
    axis.text.y = element_text(angle = 0),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.text = element_blank(),
    strip.background = element_blank(),
    legend.margin = margin(t = -5, r = 0, b = 0, l = 0),
    legend.box.margin = margin(0, 0, 0, 0),
    legend.key.size = unit(0.6, "lines"),
    legend.box.spacing = unit(0, "pt")) +
  scale_x_discrete(
    labels = c("None" = "Unaffected")) +
  scale_fill_manual(values = c(
    "None" = color_fdr_none_abs, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib), guide = "none") +
  scale_color_manual(values = c(
    "None" = color_fdr_none_abs, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib), guide = "none")

f2_a_abs

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/fig_2_a_graph_abstract.png"),
  f2_a_abs, units = "in",  width = 3, height = 2.2, bg = "white")
```



Figure 2B:

```{r fig.width=3, fig.height=2.2}
f2_b_abs <- f2_b
f2_b_abs$layers <- f2_b_abs$layers[!sapply(
  f2_b_abs$layers,
  function(x) inherits(x$geom, "GeomText")
)]


f2_b_abs <- f2_b_abs +
  theme_classic() +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "white", color = "black")) +
  scale_x_continuous(breaks = c(0,3,6,9,12)) +
  scale_fill_manual(values = c(
    "None" = color_fdr_none_abs, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib),
    labels = c("None" = "Unaffected")) +
  scale_color_manual(values = c(
    "None" = color_fdr_none_abs, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib),
    labels = c("None" = "Unaffected"))


f2_b_abs

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/fig_2_b_graph_abstract.png"),
  f2_b_abs, units = "in",  width = 3, height = 2.2, bg = "white")
```


## **Supplemental Figure 1**

Consort diagram of study inclusions/exclusions made directly in Powerpoint.

## **Supplemental Figure 2**

Box plots of GRS2x components by family history.

```{r}
# Set plot scales
non_hla_min <- 
  floor(min(pheno_all$Non_HLA, pheno_ctrls$Non_HLA) - 0.5)
  # -0.5 to allow for group size labels
non_hla_max <- 
  ceiling(max(pheno_all$Non_HLA, pheno_ctrls$Non_HLA) + 0.5)
  # +0.5 to allow for significance bars

```


### FigS1A

T-test for FS1 A:

```{r}
pheno_all <- pheno_all %>%
  dplyr::mutate(IA = as.factor(IA))
t_fs1_a <- t_test(GRS2x ~ IA, data = pheno_all) %>%
    dplyr::mutate(
      p.label = ifelse(
        p < 0.01,
        formatC(p, format = "e", digits = 1),
        formatC(p, format = "f", digits = 3)))
print(t_fs1_a)

```


```{r}
sf1_a <- ggplot(pheno_all, aes(x = IA, y = GRS2x, fill = as.factor(IA))) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(fill ="IA") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(
    values = c("1" = color_case, "0" = color_control),
    labels = c("1" = "Yes", "0" = "No")) +
  stat_pvalue_manual(
    t_fs1_a, y.position = grs2x_max - 1.5, tip.length = 0.01, label = "P={p.label}")

```


### FigS1B

GRS2x by sex.

T-test for FS1 B:

```{r}
t_fs1_b <- t_test(GRS2x ~ sex, data = pheno_all) %>%
  dplyr::mutate(
    p.label = ifelse(
      p < 0.01,
      formatC(p, format = "e", digits = 1),
      formatC(p, format = "f", digits = 3)))
print(t_fs1_b)

```


```{r}
# male = 1, female = 0
sf1_b <- ggplot(pheno_all, aes(x = sex, y = GRS2x, fill = sex)) +
  geom_violin(draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(fill ="sex") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(
    values = c("1" = color_case, "0" = color_control),
    labels = c("1" = "Male", "0" = "Female")) +
  stat_pvalue_manual(
    t_fs1_b, y.position = grs2x_max - 1.5, tip.length = 0.01, label = "P={p.label}")

```


Also write out sex (male = 1, female = 0) means so can report in text:

```{r}
table1(~ GRS2x | sex, data=pheno_all)
```


### FigS1C

ANOVA for FS1 C:

```{r out.width = '50%', fig.show = 'hold'}
# ANOVA for F1 A - use base R to run and check assumptions, but plot using
# ggpubr/rstatix
anova_fs1_c_check <- aov(GRS2x ~ fdr_4level, data = pheno_ctrls)
print(summary(anova_fs1_c_check))
plot(anova_fs1_c_check)  # check residual assumptions
car::leveneTest(GRS2x ~ fdr_4level, data = pheno_ctrls)  # check variances

# For plotting
anova_fs1_c <- anova_test(GRS2x ~ fdr_4level, data = pheno_ctrls)

```

```{r}
sf1_c <- ggplot(pheno_ctrls, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75)) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(values = c(
    "None" = color_fdr_none, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib)) +
  labs(subtitle = sprintf("ANOVA, P=%.3g", anova_fs1_c$p))

```

### FigS1D

ANOVA for FS1 D:

```{r out.width = '50%', fig.show = 'hold'}
# ANOVA for F1 A - use base R to run and check assumptions, but plot using
# ggpubr/rstatix
anova_fs1_d_check <- aov(Non_HLA ~ fdr_4level, data = pheno_all)
print(summary(anova_fs1_d_check))
plot(anova_fs1_d_check)  # check residual assumptions
car::leveneTest(Non_HLA ~ fdr_4level, data = pheno_all)  # check variances

# For plotting
anova_fs1_d <- anova_test(Non_HLA ~ fdr_4level, data = pheno_all)

```

Tukey's HSD for FS1 D:

```{r}
tukey_fs1_d <- pheno_all %>%
  tukey_hsd(Non_HLA ~ fdr_4level) %>%
  add_xy_position(x = "fdr_4level") %>%
  dplyr::mutate(
    p.adj.label = ifelse(
      p.adj < 0.01,
      formatC(p.adj, format = "e", digits = 1),
      formatC(p.adj, format = "f", digits = 3)))
print(tukey_fs1_d)

```

```{r}
sf1_d <- ggplot(pheno_all, aes(x = fdr_4level, y = Non_HLA, fill = fdr_4level)) +
  geom_violin(alpha = 0.75, draw_quantiles = c(0.25, 0.5, 0.75)) +
  labs(y = "Non-HLA GRS2x") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(non_hla_min, non_hla_max, by = 2), limits = c(non_hla_min, non_hla_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = non_hla_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(values = c(
    "None" = color_fdr_none, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib)) +
  stat_pvalue_manual(
    tukey_fs1_d, tip.length = 0.01, hide.ns = TRUE,
    label = "P-adj={p.adj.label}", step.increase = 0.08) +
  labs(subtitle = sprintf("ANOVA, P=%.3g", anova_fs1_d$p))

```

### Combine & Write Out

```{r}
sf1_ab <- ggarrange(
  sf1_a, sf1_b,
  labels = c("A", "B"), nrow=1, widths = c(.5, .5))
sf1_cd <- ggarrange(
  sf1_c, sf1_d,
  labels = c("C", "D"), nrow=1, widths = c(.5, .5),
  common.legend = T, legend = "bottom")

sf1 <- ggarrange(
  sf1_ab,
  sf1_cd,
  nrow=2, widths = c(.5, .5), heights = c(1,1))

sf1

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/supp_fig_2.pdf"),
  sf1, units = "in",  width = 7, height = 6, bg = "white")

```


## **Supplemental Figure 3**

```{r}
# Prep data
m2_female <- dplyr::bind_rows(
  m2_female_ia,
  m2_female_t1d_prog) %>%
  dplyr::mutate(outcome = case_when(
    str_detect(form, "T1D,") ~ "T1D",
    str_detect(form, "fupT1D_prog") ~ "Progression",
    TRUE ~ outcome)) %>%
  dplyr::select(-which_fdr, -study)
m2_male <- dplyr::bind_rows(
  m2_male_ia,
  m2_male_t1d_prog) %>%
  dplyr::mutate(outcome = case_when(
    str_detect(form, "T1D,") ~ "T1D",
    str_detect(form, "fupT1D_prog") ~ "Progression",
    TRUE ~ outcome)) %>%
  dplyr::select(-which_fdr, -study)

# Get common limits
get_max_hr <- function(data) {
  val <- data %>%
    dplyr::select(starts_with("conf.high_fdr")) %>%
    unlist() %>%
    max(na.rm = TRUE)
  return(exp(val))
}
get_min_hr <- function(data) {
  val <- data %>%
    dplyr::select(starts_with("conf.low_fdr")) %>%
    unlist() %>%
    min(na.rm = TRUE)
  return(exp(val))
}
max_hr <- round(max(
  get_max_hr(m2_female), get_max_hr(m2_male)))
min_hr <- round(min(
  get_min_hr(m2_female), get_min_hr(m2_male)))

```


Sex-stratified version of results in Figure 2A.

```{r}
sf2_a <- make_h2_forest_plot_manuscript(
  m2_female, "dad", "GRS2x", fdr_order, outcome_order) +
  scale_y_continuous(limits = c(min_hr, max_hr))
sf2_b <- make_h2_forest_plot_manuscript(
  m2_male, "dad", "GRS2x", fdr_order, outcome_order) +
  scale_y_continuous(limits = c(min_hr, max_hr))

```

```{r}
sf2 <- ggarrange(
  sf2_a +
    labs(y = ""),
  sf2_b,
  labels = c("A", "B"), nrow=2,
  common.legend = T, legend = "right")

sf2

ggsave(
  paste0(study_specs$teddy$manuscript_out_dir, "/figures/supp_fig_3.pdf"),
  sf2, units = "in",  width = 7, height = 4, bg = "white")

```


### Frequency Table of Sample Size

Also want to see sample sizes stratified by offspring sex in addition to
outcome and affected relative (already in table 1).

**Female:**

```{r}
t_all_f <- t_all %>%
  dplyr::filter(sex == 0)

table1(
  ~ as.factor(IA) + as.factor(T1D) | fdr_4level,
  data = t_all_f)

```


**Male:**

```{r}
t_all_m <- t_all %>%
  dplyr::filter(sex == 1)

table1(
  ~ as.factor(IA) + as.factor(T1D) | fdr_4level,
  data = t_all_m)

```

### Models with Sex*FDR Interaction

Check p-value for the Wald test of all levels of sex*FDR interaction.

*TODO: if keep this section, need to switch from manual rename of .csv files!*

**IA:**

```{r}
m2_ia_intxn %>% dplyr::select(outcome, term, grep("global", colnames(.)))

```


**T1D & Prog:**

```{r}
m2_t1d_prog_intxn%>%
  dplyr::mutate(outcome = case_when(
    str_detect(form, "T1D,") ~ "T1D",
    str_detect(form, "fupT1D_prog") ~ "Progression",
    TRUE ~ outcome)) %>%
  dplyr::select(outcome, term, grep("global", colnames(.)))
```
# **Tables**
## **Table 1**

```{r}
# Combine datasets and add Group column
t_all_ia <- t_all %>%
  dplyr::filter(IA == 1)
t_all_t1d <- t_all %>%
  dplyr::filter(T1D == 1)
pheno <- rbind(
  t_ctrls_ia %>% dplyr::mutate(group = "Unaffected"),
  t_all_ia %>% dplyr::mutate(group = "IA"),
  t_all_t1d %>% dplyr::mutate(group = "T1D"))

# Set variable labels
label(pheno$sex) <- "Female, N (%)"
label(pheno$fdr_4level) <- "Affected relative, N (%)"
label(pheno$dr34) <- "High-risk HLA DR3/4, N (%)"
label(pheno$GRS2x) <- "GRS2x, Mean (SD)"
label(pheno$cc) <- "Clinical center, N (%)"

# Create the table
t1 <- table1(
  ~ sex + fdr_4level + dr34 + GRS2x + cc + fupIA + fupT1D | group,
  data = pheno, overall = FALSE)

# Convert table1 object to data frame
t1_df <- data.frame(lapply(as.data.frame(t1), as.character), stringsAsFactors = FALSE)
colnames(t1_df) <- c("level", colnames(t1_df)[-1])
t1_df$var_name <- apply(t1_df[, -1, drop = FALSE], 1, function(x) any(x == ""))
t1_df_mod <- t1_df %>%
  mutate(
    label = if_else(var_name, level, NA)) %>%
  fill(label) %>%
  dplyr::filter(!var_name) %>%
  dplyr::mutate(level = trimws(level))

# Prepare for export
# clinical center: 1=COL, 2=GEO, 3=WAS, 4=FIN, 5=GER, 6=SWE
t1_df_mod_exp <- t1_df_mod %>%
  dplyr::select(label, level, Unaffected, IA, T1D) %>%
  dplyr::mutate(label = ifelse(is.na(label), "N (%)", label)) %>%
  dplyr::filter(!(label == "Female, N (%)" & level == 1)) %>% # only list female
  dplyr::filter(!(label == "High-risk HLA DR3/4, N (%)" & level == 0)) %>%
  dplyr::filter(!level == "Median [Min, Max]") %>%
  dplyr::mutate(level = case_when(
    label == "Clinical center, N (%)" & level == 1 ~ "COL",
    label == "Clinical center, N (%)" & level == 2 ~ "GEO",
    label == "Clinical center, N (%)" & level == 3 ~ "WAS",
    label == "Clinical center, N (%)" & level == 4 ~ "FIN",
    label == "Clinical center, N (%)" & level == 5 ~ "GER",
    label == "Clinical center, N (%)" & level == 6 ~ "SWE",
    TRUE ~ level)) %>%
  dplyr::mutate(level = case_when(
    level == "Dad" ~ "Father",
    level == "Mom" ~ "Mother",
    level == "Sib" ~ "Sibling",
    TRUE ~ level)) %>%
  dplyr::mutate(across(everything(), ~ gsub("\\(N=([0-9]+)\\)", "\\1", .)))

# Add percentages to group Ns
cols <- c("Unaffected", "IA", "T1D")
rows <- which(t1_df_mod_exp$label == "N (%)")
vals <- as.numeric(t1_df_mod_exp[rows, cols])
percs <- round(vals / nrow(t_all) * 100, 1)
t1_df_mod_exp[rows, cols] <-
  paste0(vals, " (", percs, "%)")

# Move all follow up time into single variable
t1_df_mod_exp$T1D <- ifelse(
    t1_df_mod_exp$label == "fupIA",
    t1_df_mod_exp$T1D[t1_df_mod_exp$label == "fupT1D"],
    t1_df_mod_exp$T1D)
t1_df_mod_exp$label[t1_df_mod_exp$label == "fupIA"] <-
  "Time-to-event or last follow-up (years), Mean (SD)"
t1_df_mod_exp <- t1_df_mod_exp[t1_df_mod_exp$label != "fupT1D", ]

# Write out
write_csv(
  t1_df_mod_exp,
  paste0(study_specs$teddy$manuscript_out_dir, "/tables/table_1.csv"))

```



## **Supplemental Table 1**

Want to show source of GRS2x variants as well as AFs split by T1D-affected
mom vs. T1D-affected dad.

```{r}
# Add in variant source
tm_info <- tm_info %>%
  dplyr::mutate(chr_pos = paste0(`#CHROM`, ":", POS)) %>%
  dplyr::mutate(
    Typed   = grepl("TYPED", INFO),
    Imputed = grepl("IMPUTED", INFO),
    SNP_Category = case_when(
      Typed & Imputed  ~ "typed_and_tm_r3_imp",
      Typed & !Imputed ~ "typed",
      !Typed & Imputed ~ "tm_r3_imp",
      TRUE             ~ NA))

# Add missing grs2x imputed SNPs to TOPMed imputation
imp_grs2x <- log_grs2x %>%
  dplyr::filter(grs2x_imp == "imputed") %>%
  dplyr::filter(!chr_pos %in% tm_info$chr_pos) %>%
  dplyr::mutate(SNP_Category = "grs2x_imp")
all_snps <- full_join(
  tm_info,
  imp_grs2x) %>%
  dplyr::select(chr_pos, SNP_Category)

# Merge this info into table with SNPs & AFs
af_fam_grs2x_src <- left_join(
  af_fam_grs2x,
  all_snps,
  by = c("POSITION_HG38" = "chr_pos"))

# Merge in PLINK statistical test results
af_fam_grs2x_src_test_tmp <- left_join(
  af_fam_grs2x_src,
  af_fam_grs2x_test_m %>%
    dplyr::select(SNP, P) %>%
    dplyr::rename(P_mother = P),
  by = c("ID" = "SNP"))
af_fam_grs2x_src_test <- left_join(
  af_fam_grs2x_src_test_tmp,
  af_fam_grs2x_test_s %>%
    dplyr::select(SNP, P) %>%
    dplyr::rename(P_sibling = P),
  by = c("ID" = "SNP"))

# Format for export
af_fam_grs2x_src_test_exp <- af_fam_grs2x_src_test %>%
  dplyr::select(-POSITION_HG38, -STRAND, -REPLACED) %>%
  dplyr::relocate(
    c(ALT_FREQS_father, ALT_FREQS_mother, ALT_FREQS_sibling), .after = ALT_FREQ) %>%
  dplyr::relocate(
    ID, .after = COMPONENT) %>%
  dplyr::relocate(
    P_mother, P_sibling, .after = ALT_FREQS_sibling) %>%
  dplyr::relocate(
    EFFECT_ALELLE, .after = RSID) %>%
  dplyr::rename(
    alt_freqs_grs2x = ALT_FREQ,
    source = SNP_Category,
    id_hg38 = ID,
    alt_freqs_mother_p_value = P_mother,
    alt_freqs_sibling_p_value = P_sibling) %>%
  dplyr::mutate(
    alt_freqs_grs2x = as.numeric(alt_freqs_grs2x),
    ALT_FREQS_mother = as.numeric(ALT_FREQS_mother),
    ALT_FREQS_father = as.numeric(ALT_FREQS_father),
    ALT_FREQS_sibling = as.numeric(ALT_FREQS_sibling)) %>%
  dplyr::mutate(across(where(is.numeric), ~ signif(.x, 3)))
colnames(af_fam_grs2x_src_test_exp) <- tolower(colnames(af_fam_grs2x_src_test_exp))

# Write Excel file
write_csv(
  af_fam_grs2x_src_test_exp,
  paste0(study_specs$teddy$manuscript_out_dir, "/tables/supp_table_1.csv"))

```


### **Data for Text**

Want to summarize the TOPMed imputation quality to include in the text.

```{r}
tm_info_r2 <- as.numeric(str_extract(tm_info$INFO, "(?<=R2=)\\d+(?:\\.\\d+)?"))
summary(tm_info_r2)
print(paste0("Mean = ", mean(tm_info_r2)))
```


## **Supplemental Table 2**

All results from hypothesis 2. Read in results here, reformat, and write out
to manuscript tables folder.

```{r}
fdr <- c("mom", "sib", "none")
cols_to_select <- c(
  "outcome", "term", "nobs", "form", "fdr_global_p",
  paste0("estimate_fdr_", fdr),
  paste0("conf.low_fdr_", fdr),
  paste0("conf.high_fdr_", fdr),
  paste0("z.value_fdr_", fdr),
  paste0("p.value_fdr_", fdr),
  paste0("std.error_fdr_", fdr))
m2_exp <- m2_merge %>%
  select(all_of(cols_to_select)) %>%
  pivot_longer(
    cols = matches(
      "estimate_|conf\\.low_|conf\\.high_|z\\.value_|p\\.value_|std\\.error_"),
    names_to = c(".value", "fdr"),
    names_pattern = 
      "(estimate|conf\\.low|conf\\.high|z\\.value|p\\.value|std\\.error)_(.*)") %>%
  dplyr::mutate(
    hr = exp(estimate),
    hr_conf_low = exp(conf.low), # so showing HR, not beta/estimate
    hr_conf_high = exp(conf.high)) %>%  
  dplyr::relocate(hr, hr_conf_low, hr_conf_high, .after = conf.high) %>%
  dplyr::relocate(fdr_global_p, .after = p.value) %>%
  dplyr::mutate(fdr = case_when(
    fdr == "fdr_mom" ~ "Mother",
    fdr == "fdr_sib" ~ "Sibling",
    fdr == "fdr_none" ~ "None")) %>%
  dplyr::mutate(fdr = factor(fdr, levels = rev(fdr_order))) %>%
  dplyr::rename(
    affected_relative = fdr,
    grs2x_adjustment = term,
    affected_relative_overall_p_value = fdr_global_p) %>%
  dplyr::mutate(grs2x_adjustment = ifelse(is.na(grs2x_adjustment), "Without", "With"))  %>%
  dplyr::mutate(across(-nobs & where(is.numeric), ~ signif(.x, 3))) %>%
  dplyr::mutate(
    outcome = factor(outcome, levels = outcome_order)) %>%
  dplyr::arrange(outcome)
colnames(m2_exp) <- gsub("\\.", "_", colnames(m2_exp))

# Write Excel file
write_csv(
  m2_exp,
  paste0(study_specs$teddy$manuscript_out_dir, "/tables/supp_table_2.csv"))

```


# **Results Only In Text**

## **Follow-Up Time**

Want follow-up time of whole analysis population. This should be best
represented by the "fupT1D" variable that captures T1D case date or control
last visit date:

```{r}
summary(t_all$fupT1D)
```

## **Ancestry Estimation**

Want to describe the Ns from each non-EUR 1000G population that were removed.

```{r, message = F}
# Re-running part of the utils_teddy.R logic here, if more time, revisit and
# re-work so can use function(s) here
pheno_intermed <- prep_teddy_base(study_specs$teddy)

gmm_pop_filt <- gmm_pop %>%
  dplyr::filter(IID %in% pheno_intermed$ID)
table(gmm_pop_filt$pop_gmm)
  
```


## **GRS2x Range**

The updated table 1 only shows unaffected, IA, and T1D, but old paper results
show range for all `r n_distinct(t_all$ID)`:

+ Summary: `r summary(t_all$GRS2x)`
+ Standard deviation: `r sd(t_all$GRS2x)`


## **CoxPH Violation**

More complete analysis was completed in time_varying_analysis.Rmd, but just
reporting the IA model Cox proportional hazards violation here.

```{r}
show_cox_zph(m2_ia_ph)
```


# **Session Info**

```{r session_info}
sessionInfo()

```
