---
title: "Manuscript_Famhist_PRS"
author: "NV, Updated by SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: |
  (function(inputFile, encoding) {
    library(here)
    source(here("R", "study_specs.R"))
    rmarkdown::render(
      inputFile,
      encoding = encoding,
      output_dir = study_specs$teddy$report_out_dir
    )
  })
---

```{r setup, include=FALSE}

library(here)
source("config.R")
source(here("config.R"))
source(here("R/study_specs.R"))
source(here("set_paper_colors.R"))

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = study_specs$teddy$result_out_dir)

library(tidyverse)
library(table1)
library(ggpubr)
library(haven)
library(dplyr)
library(tidyr)
library(janitor)
library(ggplot2)
library(knitr)
library(khroma)
library(writexl)
library(sas7bdat)
library(readxl)
library(data.table)
library(tidytext)
library(kableExtra)

```

# **Overview**

This report generates all figures, tables, and text-referenced values for the
MatProt PRS manuscript.

*TO NOTE:* previous versions of this report included both DAISY & TEDDY, as well
as multiple data versions (HLA-DR filtering, FDR none as reference for modeling,
Non-HLA GRS2x component only analyses, and sex-stratified results.). The current
report has been cleaned up to only include the final version presented in the
manuscript. The results for all are still saved at
Maternal_Protection/data/results/famhist_prs.

### Subset Genetic Input

GRS2x variant selection and genotype subsetting were performed prior to analysis
(see `scripts/extract_grs2x_variants.*`).

*TODO: move this to somewhere else?*


# **Load Data**


```{r, message = F}
fdr_order <- c("Father", "Mother", "Sibling", "None")
outcome_order <- c("IA", "Progression", "T1D")

# From step 01 - read in pheno files
t_all <- readRDS(
  paste0(study_specs$teddy$intermed_out_dir, "/pheno_all_ia.rds"))
  # same as pheno_all_t1d.rds
t_all_prog <- readRDS(
  paste0(study_specs$teddy$intermed_out_dir, "/pheno_all_prog.rds"))
t_ctrls_ia <- readRDS(
  paste0(study_specs$teddy$intermed_out_dir, "/pheno_ctrls_ia.rds"))

# # From step 02 - read in hypoth 1 results
# m1_dad <- read_csv(
#   "Maternal_Protection/data/results/famhist_prs/lm_GRS2x_fdr_dad_ref_daisy_teddy_all.csv")

# # From step 03 - read in hypoth 2 results
# m2_dad <- read_csv(
#   "Maternal_Protection/data/results/famhist_prs/lm_outcome_fdr_dad_ref_GRS2x_daisy_teddy_all.csv")




# # GRS2x TEDDY log
# grs2x_teddy <- fread(
#   "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/tm_r3_imp/full_qc/score_results_with_imp_log_summary.txt")
# 
# # TOPMed R3 information
# tmr3_teddy <- read_tsv(
#   "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/chr_all_concat_grs2x.pvar",
#   comment = "##")
# 
# # Read in male/female AFs
# af_male <- read_delim(
# "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/chr_all_concat_grs2x_male.afreq")
# af_female <- read_delim(
# "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/chr_all_concat_grs2x_female.afreq")
# 
# # Read in T1D Mom/Dad AFs
# af_d <- read_delim(
# "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/chr_all_concat_grs2x_t1d_dad.afreq")
# af_m <- read_delim(
# "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/chr_all_concat_grs2x_t1d_mom.afreq")

```



# **Figure 1**

Boxplot of GRS2x scores and forest plot of model 1 results.

```{r}
# data <- m1_dad
# ref <- "dad"
make_h1_forest_plot <- function(data, ref) {
  
  # Map reference to the other terms
  fdr_all <- c("dad", "mom", "sib", "none")
  fdr <- setdiff(fdr_all, ref)
  
  # Dynamically select columns
  cols_to_select <- c(
    "study",  # "group",
    paste0("estimate_fdr_", fdr),
    paste0("conf.low_fdr_", fdr),
    paste0("conf.high_fdr_", fdr))
  
  data_plot <- data %>%
    select(all_of(cols_to_select)) %>%
    pivot_longer(
      cols = matches("estimate_|conf\\.low_|conf\\.high_"),
      names_to = c(".value", "term"),
      names_pattern = "(estimate|conf\\.low|conf\\.high)_(.*)") %>%
    dplyr::mutate(term = recode(
      term,
      fdr_mom = "Mother",
      fdr_sib = "Sibling",
      fdr_dad = "Father",
      fdr_none = "None")) %>%
    dplyr::mutate(term = factor(term, levels = rev(fdr_order)))
  
  # Build plot
  plot <- ggplot(data_plot, aes(
    x = term, y = estimate, ymin = conf.low, ymax = conf.high, color = term, fill = term)) +
    geom_linerange(position = position_dodge(width = 0.5), show.legend = FALSE, size = 0.8) +
    geom_hline(yintercept = 0, lty = 2) +
    geom_point(size = 2.5, shape = 21, colour = "white", stroke = 0.5,
               position = position_dodge(width = 0.5)) +
    scale_fill_manual(values = c("Father" = color_fdr_dad,
                                 "Mother" = color_fdr_mom,
                                 "Sibling" = color_fdr_sib,
                                 "None" = color_fdr_none)) +
    scale_color_manual(values = c("Father" = color_fdr_dad,
                                  "Mother" = color_fdr_mom,
                                  "Sibling" = color_fdr_sib,
                                  "None" = color_fdr_none)) +
    coord_flip() +
    theme_bw() +
    labs(x = "", y = "Difference in GRS2x", fill = "FDR") +
    theme(legend.position = "none",
          strip.background = element_rect(fill = "white", color = "black"),
          axis.text.y = element_text(angle = 90, hjust = 0.5))
  
  return(plot)
  
}

```

### **Boxplot**

```{r}
# Prepping here for supplementary figure 1 as well
pheno_all <- t_all %>%
  dplyr::select(fdr_4level, GRS2x, Non_HLA, ia, persist_conf_ab, sex) %>%
  dplyr::mutate(study = "TEDDY") %>%
  dplyr::mutate(fdr_4level = recode(
    fdr_4level,
    Mom = "Mother",
    Sib = "Sibling",
    Dad = "Father",
    None = "None")) %>%
    dplyr::mutate(fdr_4level = factor(fdr_4level, levels = fdr_order))

pheno_ctrls <- t_ctrls_ia %>%
  dplyr::select(fdr_4level, GRS2x, Non_HLA, ia, persist_conf_ab, sex) %>%
  dplyr::mutate(study = "TEDDY") %>%
  dplyr::mutate(fdr_4level = recode(
    fdr_4level,
    Mom = "Mother",
    Sib = "Sibling",
    Dad = "Father",
    None = "None")) %>%
  dplyr::mutate(fdr_4level = factor(fdr_4level, levels = fdr_order))

# Set plot scales
grs2x_min <- 
  floor(min(pheno_all$GRS2x, pheno_ctrls$GRS2x) - 0.5)
  # -0.5 to allow for group size labels
grs2x_max <- 
  ceiling(max(pheno_all$GRS2x, pheno_ctrls$GRS2x))
```


```{r}
f1_a <- ggplot(pheno_all, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  # geom_boxplot(varwidth = TRUE, width = 1) +
  geom_boxplot(alpha = 0.75) +
  labs(x = "") +
  theme_bw() +
  theme(# axis.title.x = element_blank(),
        # axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "none",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(values = c(
    "None" = color_fdr_none, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib))

f1_a
```

### **Dad as Ref**

```{r}
# Remove DAISY, filter to only analysis in "all", not "ctrls
m1_dad_filt <- m1_dad %>%
  dplyr::filter(study == "teddy")

f1_dad <- make_h1_forest_plot(m1_dad_filt, "dad")
f1_dad

```

```{r}
# Save combined figure
f1 <- ggarrange(
  f1_a,
  f1_dad,
  labels = c("A", "B"), nrow=1, widths = c(.5, .5))

f1

ggsave(
  "Maternal_Protection/dissemination/famhist_prs_paper/figures/fig_1_hypoth_1_forest_plots_dad_ref.png",
  f1, units = "in",  width = 7, height = 3, bg = "white")
```



# **Figure 2**

Forest plot of hypothesis 2 model results.

```{r}
# data <- m2_dad
# ref <- "dad"
# term <- "GRS2x"
make_h2_forest_plot <- function(data, ref, term) {
  
  fdr_all <- c("dad", "mom", "sib", "none")
  fdr <- setdiff(fdr_all, ref)
  
  cols_to_select <- c(
    "study", "term", "outcome",
    paste0("estimate_fdr_", fdr),
    paste0("conf.low_fdr_", fdr),
    paste0("conf.high_fdr_", fdr))
  
  data_plot <- data %>%
    select(all_of(cols_to_select)) %>%
    dplyr::rename(model_type = term) %>%
    pivot_longer(
      cols = matches("estimate_|conf\\.low_|conf\\.high_"),
      names_to = c(".value", "term"),
      names_pattern = "(estimate|conf\\.low|conf\\.high)_(.*)") %>%
    dplyr::mutate(term = recode(
      term,
      fdr_mom = "Mother",
      fdr_sib = "Sibling",
      fdr_dad = "Father",
      fdr_none = "None")) %>%
    dplyr::mutate(
      or = exp(estimate),
      or_conf_low = exp(conf.low),
      or_conf_high = exp(conf.high)) %>%  # so plotting OR, not beta
    dplyr::mutate(term = factor(term, levels = rev(fdr_order))) %>%
    dplyr::mutate(outcome = case_when(
      outcome == "case_t1dprog" ~ "Progression",
      outcome == "ia" ~ "IA",
      outcome == "persist_conf_ab" ~ "IA",
      outcome == "t1d" ~ "T1D"))  %>%
    dplyr::mutate(outcome = factor(outcome, levels = outcome_order))
  
  data_plot <- data_plot %>%
    dplyr::filter(is.na(model_type) | model_type == !!term) %>%
    dplyr::filter(!duplicated(.)) %>%  # remove extra identical rows from without term
    dplyr::mutate(model_type_ind = ifelse(is.na(model_type), "Without GRS2x", "With GRS2x")) %>%
    dplyr::mutate(model_type_ind = factor(model_type_ind, levels = c("With GRS2x", "Without GRS2x")))

  plot <- ggplot(
    data_plot,
    aes(x=term, y=or, ymin=or_conf_low, ymax=or_conf_high, color=term,
        fill=term, alpha=model_type_ind)) + 
    geom_linerange(position=position_dodge(width = 0.5), show.legend = F, size=0.8) +
    geom_hline(yintercept = 1, lty=2) +
    geom_point(size=2.5, shape=21, colour="white", stroke = 0.5,
               position=position_dodge(width = 0.5)) +
    scale_color_manual(values = c(
      "Father" = color_fdr_dad,
      "Mother" = color_fdr_mom,
      "Sibling" = color_fdr_sib,
      "None" = color_fdr_none),
      guide = "none") +
    scale_fill_manual(values = c(
      "Father" = color_fdr_dad,
      "Mother" = color_fdr_mom,
      "Sibling" = color_fdr_sib,
      "None" = color_fdr_none),
      guide = "none") +
    scale_alpha_manual(
      values = c("Without GRS2x" = 0.4, "With GRS2x" = 1)) +
    guides(
    alpha = guide_legend(
      reverse = T,
      override.aes = list(
        shape = 21,
        size  = 3,
        fill  = "grey50",
        color = "grey50"))) +
    coord_flip() +
    theme_bw() +
    labs(
      x = "",
      y = "Hazard Ratio",
      alpha = "",
    ) +
    scale_x_discrete(labels = c("teddy" = "TEDDY","daisy" = "DAISY")) +
    theme(legend.position = "bottom",
          strip.background = element_rect(fill = "white", color = "black"),
          axis.text.y = element_text(angle = 90, hjust = 0.5)) +
    facet_wrap(~outcome)
  
  return(plot)
  
}

```

## **With/Without GRS2x**

### **Dad as Ref**

```{r}
m2_dad_filt <- m2_dad %>%
  dplyr::filter(study == "teddy")

f2_dad <- make_h2_forest_plot(m2_dad_filt, "dad", "GRS2x")
f2_dad

ggsave(
  "Maternal_Protection/dissemination/famhist_prs_paper/figures/fig_2_hypoth_2_grs2x_forest_plots_dad_ref.png",
  f2_dad, units = "in",  width = 7, height = 3, bg = "white")

```


# **Supplemental Figure 3**

Box plots of GRS2x components by family history.

```{r}
# Set plot scales
non_hla_min <- 
  floor(min(pheno_all$Non_HLA, pheno_ctrls$Non_HLA) - 0.5)
  # -0.5 to allow for group size labels
non_hla_max <- 
  ceiling(max(pheno_all$Non_HLA, pheno_ctrls$Non_HLA))

```


### FigS3A

```{r}
sf3_a <- ggplot(pheno_ctrls, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  # geom_boxplot(varwidth = TRUE, width = 1) +
  geom_boxplot(alpha = 0.75) +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(values = c(
    "None" = color_fdr_none, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib))

```

### FigS3B

```{r}
sf3_b <- ggplot(pheno_all, aes(x = fdr_4level, y = Non_HLA, fill = fdr_4level)) +
  # geom_boxplot(varwidth = TRUE, width = 1) +
  geom_boxplot(alpha = 0.75) +
  labs(y = "Non-HLA GRS2x") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(non_hla_min, non_hla_max, by = 2), limits = c(non_hla_min, non_hla_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = non_hla_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(values = c(
    "None" = color_fdr_none, "Father" = color_fdr_dad,
    "Mother" = color_fdr_mom, "Sibling" = color_fdr_sib))

```

### FigS3C

IA in cases and controls - at Jill's recommendation calling these IA yes and IA
no instead of cases/controls since working in full cohort. Using persist_conf_ab
as IA here since rest of paper is defined this way as well.

```{r}
sf3_c <- ggplot(pheno_all, aes(x = persist_conf_ab, y = GRS2x, fill = as.factor(persist_conf_ab))) +
  # geom_boxplot(varwidth = TRUE, width = 1) +
  geom_boxplot() +
  labs(fill ="IA") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(
    values = c("1" = color_case, "0" = color_control),
    labels = c("1" = "Yes", "0" = "No"))

```


Also write out IA case and control means so can report in text:

```{r}
table1(~ GRS2x | persist_conf_ab, data=pheno_all)
```

### FigS1D

GRS2x by sex.

```{r}
# male = 1, female = 0
sf3_d <- ggplot(pheno_all, aes(x = sex, y = GRS2x, fill = as.factor(sex))) +
  # geom_boxplot(varwidth = TRUE, width = 1) +
  geom_boxplot() +
  labs(fill ="sex") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "bottom",
        strip.background = element_rect(fill = "white", color = "black")) +
  scale_y_continuous(
    breaks = seq(grs2x_min, grs2x_max, by = 2), limits = c(grs2x_min, grs2x_max)) +
  stat_summary(
    fun.data = function(x) data.frame(y = grs2x_min, label = length(x)),
    geom = "text") +
  scale_fill_manual(
    values = c("1" = color_case, "0" = color_control),
    labels = c("1" = "Male", "0" = "Female"))


```


Also write out sex (male = 1, female = 0) means so can report in text:

```{r}
table1(~ GRS2x | sex, data=pheno_all)
```




### Combine & Write Out


```{r}
sf3_ab <- ggarrange(
  sf3_a, sf3_b,
  labels = c("A", "B"), nrow=1, widths = c(.5, .5),
  common.legend = T, legend = "bottom")
sf3_cd <- ggarrange(
  sf3_c, sf3_d,
  labels = c("C", "D"), nrow=1, widths = c(.5, .5))

sf3 <- ggarrange(
  sf3_ab,
  sf3_cd,
  nrow=2, widths = c(.5, .5), heights = c(1,1))

sf3

ggsave(
  "Maternal_Protection/dissemination/famhist_prs_paper/figures/supp_fig_3_grs2x_boxplots.png",
  sf3, units = "in",  width = 7, height = 6, bg = "white")
```





# **Check Age at IA/T1D by FDR**

High-risk HLA more common in early onset T1D cases.
Higher miscarriage in women with early-onset T1D.
Selective loss of highest risk fetuses.
So would expect to see older onset in mom children that live?

```{r}
t_all_ia <- t_all %>% dplyr::filter(ia == 1)  # 814
ggplot(t_all_ia, aes(x=fdr_4level, y=fup4_L)) +
  geom_boxplot()

t_all_t1d <- t_all %>% dplyr::filter(t1d == 1)  # 369
ggplot(t_all_t1d, aes(x=fdr_4level, y=fup_t1d)) +
  geom_boxplot()


```




# **Table 1**

```{r}
# Combine datasets and add Group column
pheno <- t_all_prog %>%
  mutate(group = case_when(
    #TODO: continue here!! ####################################
  ))

# Set variable labels
label(pheno_all$sex) <- "Sex, N (%)"
label(pheno_all$persist_conf_ab) <- "IA, N (%)"
label(pheno_all$mAA_at_ia) <- "Multiple Autoantibodies at IA Seroconversion, N (%)"
label(pheno_all$t1d) <- "T1D, N (%)"
label(pheno_all$case_t1dprog) <- "Progression, N (%)"
label(pheno_all$fdr_4level) <- "FDR, N (%)"
label(pheno_all$dr34) <- "DR3/4, N (%)"

# Create the table
t1 <- table1(
  ~ sex + fdr_4level + dr34 + GRS2x + mAA_at_ia + as.factor(persist_conf_ab) + as.factor(t1d) +
    as.factor(case_t1dprog) | group,
  data = pheno_all, overall = FALSE)
t1

# Convert table1 object to data frame
t1_df <- as.data.frame(t1)

# Define output folder and filename
output_file <- "Maternal_Protection/dissemination/famhist_prs_paper/tables/table_1.xlsx"

# Write Excel file
write_xlsx(t1_df, path = output_file)
```



# **(Not Used) Supplemental Table 1 - AFs by Child Sex**

Want to show source of GRS2x variants as well as AFs split by sex.

First, get range of typed & TOPMed R3 imputed SNP Rsq. If wanted MAF, should
really re-calculate in subset as these files have in all TEDDY.

```{r}
tmr3_teddy <- tmr3_teddy %>%
  dplyr::mutate(
    R2 = as.numeric(str_extract(INFO, "(?<=R2=)[0-9.]+")))

```

Now, look at GRS2x SNP weight and allele frequencies in TEDDY, split by sex.

### Split Male/Female

*This section only run once*

```{r, eval = F}
t_all_male <- t_all %>%
  dplyr::filter(sex == 1) %>%
  dplyr::select(ID) %>%
  left_join(t_fam, by = c("ID" = "IID")) %>%
  dplyr::relocate(FID, .before = ID)
t_all_female <- t_all %>%
  dplyr::filter(sex == 0) %>%
  dplyr::select(ID) %>%
  left_join(t_fam, by = c("ID" = "IID")) %>%
  dplyr::relocate(FID, .before = ID)

write_tsv(
  t_all_male, "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/male_ids.txt",
  col_names = F)
write_tsv(
  t_all_female, "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/female_ids.txt",
  col_names = F)
```


```{bash, eval = F}
cd Immunogenetics_T1D/genetics/teddy_r01

plink2 --pfile risk_scores/grs2x/chr_all_concat_grs2x \
  --keep risk_scores/grs2x/male_ids.txt \
  --freq \
  --out risk_scores/grs2x/chr_all_concat_grs2x_male
  
plink2 --pfile risk_scores/grs2x/chr_all_concat_grs2x \
  --keep risk_scores/grs2x/female_ids.txt \
  --freq \
  --out risk_scores/grs2x/chr_all_concat_grs2x_female

```

### Make Table

```{r}
# First, get AFs split by sex
af_comb <- inner_join(
  af_female %>% dplyr::select(-`#CHROM`, -OBS_CT),
  af_male %>% dplyr::select(-`#CHROM`, -OBS_CT),
  by = c("ID", "REF", "ALT"),
  suffix = c("_female", "_male"))  # all 64, so male/female same ALT

af_comb_grs2x <- left_join(
  grs2x_snps %>% dplyr::mutate(ID = paste0(POSITION_HG38, ":", REF, ":", ALT)),
  af_comb,
  by = c("ID", "REF", "ALT"))

```


```{r}
# Now add in variant source
tmr3_teddy_sub <- tmr3_teddy %>%
  dplyr::mutate(chr_pos = paste0(`#CHROM`, ":", POS)) %>%
  dplyr::mutate(
    Typed   = grepl("TYPED", INFO),
    Imputed = grepl("IMPUTED", INFO),
    SNP_Category = case_when(
      Typed & Imputed  ~ "typed_and_tm_r3_imp",
      Typed & !Imputed ~ "typed",
      !Typed & Imputed ~ "tm_r3_imp",
      TRUE             ~ NA))

# Add missing grs2x imputed SNPs to TEDDY
grs2x_imputed_snps <- grs2x_teddy %>%
  filter(grs2x_imp == "imputed") %>%
  filter(!chr_pos %in% tmr3_teddy_sub$chr_pos) %>%
  mutate(SNP_Category = "grs2x_imp")
all_snps <- full_join(
  tmr3_teddy_sub,
  grs2x_imputed_snps) %>%
  dplyr::select(chr_pos, SNP_Category)

# Merge this info into table with SNPs & AFs
af_comb_grs2x_src <- left_join(
  af_comb_grs2x,
  all_snps,
  by = c("POSITION_HG38" = "chr_pos"))

# Format for export
af_comb_grs2x_src_exp <- af_comb_grs2x_src %>%
  dplyr::select(-POSITION_HG38, -STRAND, -REPLACED) %>%
  dplyr::relocate(
    c(ALT_FREQS_female, ALT_FREQS_male), .after = ALT_FREQ) %>%
  dplyr::relocate(
    ID, .after = COMPONENT) %>%
  dplyr::rename(
    ALT_FREQS_grs2x = ALT_FREQ,
    source = SNP_Category,
    id_hg38 = ID)

colnames(af_comb_grs2x_src_exp) <- tolower(colnames(af_comb_grs2x_src_exp))

```


# **Supplemental Table 1 - AFs by T1D-affected Mom vs. Dad**

Want to show source of GRS2x variants as well as AFs split by T1D-affected
mom vs. T1D-affected dad.

### Split T1D Mom/T1D Dad

*This section only run once*

```{r, eval = F}
t_all_d <- t_all %>%
  dplyr::filter(fdr_4level == "Dad") %>%
  dplyr::select(ID) %>%
  left_join(t_fam, by = c("ID" = "IID")) %>%
  dplyr::relocate(FID, .before = ID)
t_all_m <- t_all %>%
  dplyr::filter(fdr_4level == "Mom") %>%
  dplyr::select(ID) %>%
  left_join(t_fam, by = c("ID" = "IID")) %>%
  dplyr::relocate(FID, .before = ID)

write_tsv(
  t_all_d, "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/t1d_dad_ids.txt",
  col_names = F)
write_tsv(
  t_all_m, "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/t1d_mom_ids.txt",
  col_names = F)
```


```{bash, eval = F}
cd Immunogenetics_T1D/genetics/teddy_r01

plink2 --pfile risk_scores/grs2x/chr_all_concat_grs2x \
  --keep risk_scores/grs2x/t1d_dad_ids.txt \
  --freq \
  --out risk_scores/grs2x/chr_all_concat_grs2x_t1d_dad
  
plink2 --pfile risk_scores/grs2x/chr_all_concat_grs2x \
  --keep risk_scores/grs2x/t1d_mom_ids.txt \
  --freq \
  --out risk_scores/grs2x/chr_all_concat_grs2x_t1d_mom

```

### Make Table

```{r}
# First, get AFs split by sex
af_fam <- inner_join(
  af_m %>% dplyr::select(-`#CHROM`, -OBS_CT),
  af_d %>% dplyr::select(-`#CHROM`, -OBS_CT),
  by = c("ID", "REF", "ALT"),
  suffix = c("_mom", "_dad"))  # all 64, so same ALT

af_fam_grs2x <- left_join(
  grs2x_snps %>% dplyr::mutate(ID = paste0(POSITION_HG38, ":", REF, ":", ALT)),
  af_fam,
  by = c("ID", "REF", "ALT"))

```


```{r}
# Now add in variant source
tmr3_teddy_sub <- tmr3_teddy %>%
  dplyr::mutate(chr_pos = paste0(`#CHROM`, ":", POS)) %>%
  dplyr::mutate(
    Typed   = grepl("TYPED", INFO),
    Imputed = grepl("IMPUTED", INFO),
    SNP_Category = case_when(
      Typed & Imputed  ~ "typed_and_tm_r3_imp",
      Typed & !Imputed ~ "typed",
      !Typed & Imputed ~ "tm_r3_imp",
      TRUE             ~ NA))

# Add missing grs2x imputed SNPs to TEDDY
grs2x_imputed_snps <- grs2x_teddy %>%
  filter(grs2x_imp == "imputed") %>%
  filter(!chr_pos %in% tmr3_teddy_sub$chr_pos) %>%
  mutate(SNP_Category = "grs2x_imp")
all_snps <- full_join(
  tmr3_teddy_sub,
  grs2x_imputed_snps) %>%
  dplyr::select(chr_pos, SNP_Category)

# Merge this info into table with SNPs & AFs
af_fam_grs2x_src <- left_join(
  af_fam_grs2x,
  all_snps,
  by = c("POSITION_HG38" = "chr_pos"))

# Format for export
af_fam_grs2x_src_exp <- af_fam_grs2x_src %>%
  dplyr::select(-POSITION_HG38, -STRAND, -REPLACED) %>%
  dplyr::relocate(
    c(ALT_FREQS_mom, ALT_FREQS_dad), .after = ALT_FREQ) %>%
  dplyr::relocate(
    ID, .after = COMPONENT) %>%
  dplyr::rename(
    ALT_FREQS_grs2x = ALT_FREQ,
    source = SNP_Category,
    id_hg38 = ID)

colnames(af_fam_grs2x_src_exp) <- tolower(colnames(af_fam_grs2x_src_exp))

# Write Excel file
write_csv(
  af_fam_grs2x_src_exp %>% mutate(across(where(is.numeric), ~ round(.x, 3))),
  "Maternal_Protection/dissemination/famhist_prs_paper/tables/supp_table_1.csv")

```



# **Supplemental Table 2**

All results from hypothesis 1. Read in results here, reformat, and write out
to manuscript tables folder.


```{r}
m1_exp <- m1_dad_filt %>%
  dplyr::mutate(model_type = "linear") %>%
  dplyr::relocate(model_type, .after = nobs) %>%
  dplyr::select(-which_fdr, -study)

fdr <- c("mom", "sib", "none")
cols_to_select <- c(
  "nobs", "model_type", "form",
  paste0("estimate_fdr_", fdr),
  paste0("conf.low_fdr_", fdr),
  paste0("conf.high_fdr_", fdr),
  paste0("z.value_fdr_", fdr),
  paste0("p.value_fdr_", fdr),
  paste0("std.error_fdr_", fdr))

m1_exp_mod <- m1_exp %>%
  select(all_of(cols_to_select)) %>%
  pivot_longer(
    cols = contains("_fdr_"),
    names_to = c(".value", "fdr"),
    names_pattern = 
      "(estimate|conf\\.low|conf\\.high|z\\.value|p\\.value|std\\.error)_fdr_(mom|sib|none)") %>%
  dplyr::mutate(fdr = case_when(
    fdr == "mom" ~ "Mother",
    fdr == "sib" ~ "Sibling",
    fdr == "none" ~ "None")) %>%
  dplyr::mutate(fdr = factor(fdr, levels = rev(fdr_order)))

# Write Excel file
write_csv(
  m1_exp_mod,
  "Maternal_Protection/dissemination/famhist_prs_paper/tables/supp_table_2.csv")
```


# **Supplemental Table 3**

All results from hypothesis 2. Read in results here, reformat, and write out
to manuscript tables folder.

```{r}
m2_exp <- m2_dad_filt %>%
  dplyr::mutate(outcome = case_when(
    outcome == "case_t1dprog" ~ "Progression",
    outcome == "ia" ~ "IA",
    outcome == "persist_conf_ab" ~ "IA",
    outcome == "t1d" ~ "T1D")) %>%
  dplyr::mutate(model_type = "cox_prop_hazards") %>%
  dplyr::relocate(model_type, .after = nobs) %>%
  dplyr::select(-which_fdr, -study) %>%
  dplyr::filter(is.na(term) | term == "GRS2x") %>%
  dplyr::filter(!duplicated(.)) %>% # remove extra identical rows from without term
  select(where(~ !all(is.na(.))))  # remove columns from non-GRS2x term

fdr <- c("mom", "sib", "none")
cols_to_select <- c(
  "outcome", "term", "nobs", "model_type", "form",
  paste0("estimate_fdr_", fdr),
  paste0("conf.low_fdr_", fdr),
  paste0("conf.high_fdr_", fdr),
  paste0("z.value_fdr_", fdr),
  paste0("p.value_fdr_", fdr),
  paste0("std.error_fdr_", fdr))
m2_exp <- m2_exp %>%
  select(all_of(cols_to_select)) %>%
  pivot_longer(
    cols = matches(
      "estimate_|conf\\.low_|conf\\.high_|z\\.value_|p\\.value_|std\\.error_"),
    names_to = c(".value", "fdr"),
    names_pattern = 
      "(estimate|conf\\.low|conf\\.high|z\\.value|p\\.value|std\\.error)_(.*)") %>%
  dplyr::mutate(
    hr = exp(estimate),
    hr_conf_low = exp(conf.low),
    hr_conf_high = exp(conf.high)) # >%  # so plotting OR, not beta
  
m2_exp <- m2_exp %>%
  dplyr::relocate(hr, hr_conf_low, hr_conf_high, .after = conf.high) %>%
  dplyr::mutate(fdr = case_when(
    fdr == "fdr_mom" ~ "Mother",
    fdr == "fdr_sib" ~ "Sibling",
    fdr == "fdr_none" ~ "None")) %>%
  dplyr::mutate(fdr = factor(fdr, levels = rev(fdr_order)))

# Write Excel file
write_csv(
  m2_exp,
  "Maternal_Protection/dissemination/famhist_prs_paper/tables/supp_table_3.csv")
```


# **Session Info**

```{r session_info}
sessionInfo()

```
