---
title: "Teddy Immuno T1D GRS2x"
author: "NV, Updated by SDS"
date: "Last compiled on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Immunogenetics_T1D/genetics/reports")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"))) #setting up pathway

# Load libraries
library(tidyverse)
library(haven)
library(conflicted)
library(dplyr)
library(ggplot2)
library(patchwork)
library(openxlsx)
library(data.table)
library(GGally)
library(ggcorrplot)
library(ggpubr)

```


# **Overview**

+ Genetics: Immunogenetics_T1D/genetics/teddy_r01/tm_r3_imp/imputed_clean_maf0_rsq0.3
  + *Updated 20251118 to use "full_qc" TOPMed R3 input:*
    + *Immunogenetics_T1D/genetics/teddy_r01/imputation/tm_r3_imp/exome_chip/full_qc/imputed_clean_maf0_rsq0.3/chr_all_concat*
+ Code: immuno_t1d/risk_scores/teddy_r01
+ Results: Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x
    + *Updated 20251118 to save previous results at grs2x/tm_r3_imp/hwe_in_all and*
      *updated results at grs2x/tm_r3_imp/full_qc.*
+ Pheno: Immunogenetics_T1D/raw/teddy_r01/2025-03-05/rj_immuno_all_demo_masked.sas7bdat

# **Load Data**

*TO NOTE: this section loads data that is generated in bash chunks below.*

```{r}
log_file <- read_delim(
  "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/tm_r3_imp/full_qc/score_results.log",
  delim = " - ", col_names = F)

log_file_imp <- read_delim(
  "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/tm_r3_imp/full_qc/score_results_with_imp.log",
  delim = " - ", col_names = F)

SNPs_input_impute <- fread(
  "Immunogenetics_T1D/genetics/teddy_r01/imputation/tm_r3_imp/exome_chip/full_qc/imputed_clean_maf0_rsq0.3/chr_all_concat.pvar",
  skip = "#CHROM"  # skip comment line to get clean column names
)

# Additional for specific checks
chr6_info_full_qc <- read_tsv(gzfile(
  "Immunogenetics_T1D/genetics/teddy_r01/imputation/tm_r3_imp/exome_chip/full_qc/imputed/chr6.info.gz"),
  comment = "##")

```


# **GRS2x Calculation**

Using mamba installation of PRSedm:

mamba install sethsh7::prsedm conda-forge::python bioconda::pysam

### **No GRS2x Imputation**

```{zsh engine.opts= '-i', eval=F}
mamba activate prsedm

in="Immunogenetics_T1D/genetics/teddy_r01/imputation/tm_r3_imp/exome_chip/full_qc/imputed_clean_maf0_rsq0.3/chr_all_concat.vcf.gz"

out="Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/tm_r3_imp/full_qc/score_results"

# Run PRSedm
prsedm --vcf "$in" \
       --col GT \
       --build hg38 \
       --scores t1dgrs2-luckett25 \
       --output "${out}.csv" > "${out}.log" 2>&1

```

Check Log File:

```{r}
# Add "cluster" information to log file before removing text
snps_log <- log_file %>%
  dplyr::mutate(group = ifelse(str_detect(X4, "Generating"), gsub(".+ for", "", X4), NA)) %>%
  fill(group, .direction = "down") %>%
  dplyr::mutate(grs2x_imp = case_when(
    str_detect(X4, "Imputing") ~ "imputed",
    str_detect(X4, "not found and impute=off") ~ "not_found",
    TRUE ~ "used"))  %>%
  dplyr::mutate(chr_pos = X4) %>%
  dplyr::filter(str_detect(chr_pos, "[[:digit:]]+:[[:digit:]]+")) %>%
  dplyr::mutate(chr_pos = gsub(".+ ", "", gsub("\\.", "", gsub("\\.\\.\\.", "", chr_pos)))) %>%
  group_by(chr_pos) %>%
  slice(
    if (n() > 1) {
      which(grs2x_imp %in% c("imputed", "not_found"))
    } else {
      1
    }
  ) %>%
  ungroup() %>%
  dplyr::select(group, chr_pos, grs2x_imp) %>%
  dplyr::mutate(group = gsub("^ ", "", gsub("\\.\\.\\.", "", group)))

snps_log

write_tsv(
  snps_log,
  "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/tm_r3_imp/full_qc/score_results_log_summary.txt")

```


3 missing variants were not imputed - why is this more than previous TM R3 IMP?
All are chr6 variants, so look at info files from hwe_in_all imputation and
full_qc imputation:

+ 6:32644083
+ 6:32615766
+ 6:32630313

```{r}
pos_list <- c(32644083, 32615766, 32630313)

chr6_info_full_qc_filt <- chr6_info_full_qc %>%
  dplyr::filter(POS %in% pos_list) %>%
  dplyr::select(POS, ID, INFO)

chr6_info_hwe_all_filt <- chr6_info_hwe_all %>%
  dplyr::filter(POS %in% pos_list) %>%
  dplyr::select(POS, ID, INFO)

chr6_info <- full_join(
  chr6_info_full_qc_filt,
  chr6_info_hwe_all_filt,
  by = c("POS", "ID"),
  suffix = c("_full_qc", "_hwe_all"))

chr6_info
```

Likely makes sense for variants 32615766 (already low R2 in hwe_all) and
32644083 (the previously missing variant, no INFO available due to
server-applied R2 threshold). Not sure why 32630313 would have such a big swing,
but hopefully is not influential since AF is so low. Is there a way to check
this on PRSedm?

### **With GRS2x Imputation**

Only supplying reference for the chromosome where variant was not found. Need
to remove "chr" prefix from TOPMed freeze 8 files for PRSedm.

```{zsh engine.opts= '-i', eval=F}
mamba activate bcftools

bcftools annotate --rename-chrs rename_add_chr_prefix.txt -O z -o \
  chr6.TOPMed_freeze_8.all.rename_chr.vcf.gz chr6.TOPMed_freeze_8.all.vcf.gz
tabix chr6.TOPMed_freeze_8.all.rename_chr.vcf.gz
```


```{zsh engine.opts= '-i', eval=F}
mamba activate prsedm

in="Immunogenetics_T1D/genetics/teddy_r01/imputation/tm_r3_imp/exome_chip/full_qc/imputed_clean_maf0_rsq0.3/chr_all_concat.vcf.gz"
out="Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/tm_r3_imp/full_qc/score_results_with_imp"
refvcf="Maternal_Protection/raw/genome_maps/TOPMed_freeze8/chr6.TOPMed_freeze_8.all.rename_chr.vcf.gz"

# Run PRSedm
prsedm --vcf "$in" \
       --col GT \
       --build hg38 \
       --scores t1dgrs2-luckett25 \
       --impute \
       --refvcf "$refvcf" \
       --output "${out}.csv" > "${out}.log" 2>&1

```

Check Log File:

```{r}
# Add "cluster" information to log file before removing text
snps_log_imp <- log_file_imp %>%
  dplyr::mutate(group = ifelse(str_detect(X4, "Generating"), gsub(".+ for", "", X4), NA)) %>%
  fill(group, .direction = "down") %>%
  dplyr::mutate(grs2x_imp = case_when(
    str_detect(X4, "Imputing") ~ "imputed",
    str_detect(X4, "not found and impute=off") ~ "not_found",
    TRUE ~ "used"))  %>%
  dplyr::mutate(chr_pos = X4) %>%
  dplyr::filter(str_detect(chr_pos, "[[:digit:]]+:[[:digit:]]+")) %>%
  dplyr::mutate(chr_pos = gsub(".+ ", "", gsub("\\.", "", gsub("\\.\\.\\.", "", chr_pos)))) %>%
  group_by(chr_pos) %>%
  slice(
    if (n() > 1) {
      which(grs2x_imp %in% c("imputed", "not_found"))
    } else {
      1
    }
  ) %>%
  ungroup() %>%
  dplyr::select(group, chr_pos, grs2x_imp) %>%
  dplyr::mutate(group = gsub("^ ", "", gsub("\\.\\.\\.", "", group)))

snps_log_imp

write_tsv(
  snps_log_imp,
  "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/tm_r3_imp/full_qc/score_results_with_imp_log_summary.txt")

```


# **SNP Quality Check**

Comparing with .pvar file from input data for the quality of SNPs used.

```{r}
head(SNPs_input_impute)

```

```{r}
# Make chr:pos in imputed data
SNPs_input_impute$chr_pos = paste0(
  SNPs_input_impute$`#CHROM`, ":", SNPs_input_impute$POS)

# Subset imputed data to 67 SNPs
SNPs_input_subset <- SNPs_input_impute %>%
  dplyr::filter(chr_pos %in% snps_log$chr_pos)

# Join to get cluster info alongside INFO field
SNP_quality_check <- snps_log %>%
  left_join(SNPs_input_subset, by = c("chr_pos"))

# View quality info
SNP_quality_check[, c("#CHROM", "POS", "group", "INFO")]


# Extract MAF and R2 from INFO column using regular expressions
SNP_quality_check <- SNP_quality_check %>%
  mutate(
    MAF_value = as.numeric(str_extract(INFO, "(?<=MAF=)[^;]+")),
    R2_value  = as.numeric(str_extract(INFO, "(?<=R2=)[^;]+")))

```

Summary of MAF:

`r summary(SNP_quality_check$MAF_value)`

Summary of R2:

`r summary(SNP_quality_check$R2_value)`


# **Session Info**

```{r session_info}
sessionInfo()
```

```{r}

# Check how many IIDs overlap
overlap_iids <- intersect(prs_scores_imputed$IID, GRS2x_matprot_imputed$IID)

# Number of overlapping IIDs
length(overlap_iids)

# View which IIDs overlap
print(overlap_iids)


# Visualization
library(VennDiagram)
library(grid)

# Open a new plotting page
grid.newpage()

# Draw Venn diagram
draw.pairwise.venn(
  area1 = nrow(prs_scores_imputed),
  area2 = nrow(GRS2x_matprot_imputed),
  cross.area = length(overlap_iids),
  category = c("ImmunoT1D", "Mat_prot"),
  fill = c("skyblue", "orange"),
  alpha = 0.5,
  cex = 2,
  cat.cex = 1.5,
  cat.pos = c(-20, 20)
)

```



```{r}

# Check how many IDs overlap
overlap_ids <- intersect(prs_scores_imputed$ID, GRS2x_matprot_imputed$ID)

# Number of overlapping IDs
length(overlap_ids)

# View which IDs overlap
print(overlap_ids)


# Visualization
library(VennDiagram)
library(grid)

# Open a new plotting page
grid.newpage()

# Draw Venn diagram
draw.pairwise.venn(
  area1 = nrow(prs_scores_imputed),
  area2 = nrow(GRS2x_matprot_imputed),
  cross.area = length(overlap_ids),
  category = c("ImmunoT1D", "Mat_prot"),
  fill = c("skyblue", "orange"),
  alpha = 0.5,
  cex = 2,
  cat.cex = 1.5,
  cat.pos = c(-20, 20)
)

```
