---
title: "check_genetics_teddy_daisy"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = 
  paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection/reports/TEDDYreports/teddy_risk_scores")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB")))

library(tidyverse)
library(conflicted)
library(readxl)
library(kableExtra)
library(vcfR)
library(DT)
library(ggvenn)
library(biomaRt)


```

# **Overview**

This report was prompted by VERY different GRS2 SNP availability in TEDDY
(calculated in explore_matprot/teddy_risk_scores/calc_teddy_risk_scores.Rmd on
GitHub) compared to DAISY (calculated in
DAISY/genetics/risk_scores/code/calc_daisy_grs2.Rmd). Specifically looking at
the Excel file (T1DGRS67_TOPMED_hg38.xlsx) with all haplogenotypes and SNPs in
the GRS2 (PMID 30655379) from the HLA haplotype pipeline on
[GitHub](https://github.com/sethsh7/hla-prs-toolkit).


**Outline:**

1. Compare Unimputed Exome Data
2. Compare TOPMed Imputed Data *TODO!*

# **Load Data**

```{r load_data, message = F}
# Load exome chip data
raw_bim_t <- read_delim(
  "Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_SNP_exome_masked.bim",
  col_names = c("chr", "id", "cm", "pos", "a1", "a2"))
raw_bim_d <- read_delim(
  "DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/clean/daisyexome.bim",
  col_names = c("chr", "id", "cm", "pos", "a1", "a2"))

# Load MAF for exome chip data
raw_maf_t <- read_delim(
  "Maternal_Protection/data/genetics/teddy/exome/Norris_SNP_exome_masked_freq.afreq")
raw_maf_d <- read_delim(
  "DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/clean/daisyexome_freq.afreq")

# Research GRS2 files
# hla prs toolkit Excel mapping file
map_tk <- read_xlsx(
  "Maternal_Protection/background/teddy_risk_scores/from_hla-prs-toolkit/T1DGRS67_TOPMED_hg38.xlsx")

```

# **1. Compare Unimputed Exome Data**

## **Inputs**

TEDDY: Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_SNP_exome_masked

DAISY:

+ DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/clean/daisyexome
+ DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/daisyraw0

Not sure which DAISY files to use for comparison - raw or cleaned. Can't find
documentation in TEDDY release about cleaning for TEDDY. Check number of SNPS:

```{bash}
wc -l Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_SNP_exome_masked.bim
wc -l DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/clean/daisyexome.bim
wc -l DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/daisyraw0.bim
```

Based on this, move forward with the cleaned daisyexome files for comparison.

## **Compare Variant Overlap**

*To note: both raw datasets are in hg19.*

### **Compare by RSID**

DAISY has non-standard ID formats:

```{r}
head(raw_bim_d)
```

TEDDY appears to be more standard:

```{r}
head(raw_bim_t)
```

Check if all variant names are unique in both files:

+ `r n_distinct(raw_bim_t$id)` / `r nrow(raw_bim_t$id)` distinct in TEDDY
+ `r n_distinct(raw_bim_d$id)` / `r nrow(raw_bim_d$id)` distinct in DAISY

Check overlap as is:

```{r}
raw_bim_merge_id <- inner_join(raw_bim_t, raw_bim_d, by = c("id"), suffix = c("_t", "_d"))
count_raw_bim_merge_id <- nrow(raw_bim_merge_id)

```

There are only `r count_raw_bim_merge_id` overlapping variants by the ID column.
This could be due to differences in formatting.

Do any of the variants mapped by ID have different positions?

```{r}
raw_bim_merge_id %>% dplyr::filter(pos_t != pos_d)
```

### **Compare by Chromosome and Position**

Check if all variant positions are unique in both files:

+ `r n_distinct(raw_bim_t$pos)` / `r nrow(raw_bim_t$id)` distinct in TEDDY
+ `r n_distinct(raw_bim_d$pos)` / `r nrow(raw_bim_d$id)` distinct in DAISY

There are more duplicate positions in the DAISY dataset than in the TEDDY
dataset. Keep this in mind as merge below will have two entries for one variant
since merging by chr and pos.

```{r}
raw_bim_merge_pos <- inner_join(raw_bim_t, raw_bim_d, by = c("chr", "pos"),
                                suffix = c("_t", "_d"))
count_raw_bim_merge_pos <- nrow(raw_bim_merge_pos)
count_raw_bim_merge_pos_dist <- n_distinct(raw_bim_merge_pos$pos)

```

There are many more overlapping when join by chromosome and position:

+ rows in merge: `r count_raw_bim_merge_pos`
+ distinct positions in merge: `r count_raw_bim_merge_pos_dist`

How many of these also have the same allele 1 and allele 2 (in order)?

```{r}
count_raw_bim_merge_pos_a_match <-
  nrow(raw_bim_merge_pos %>% dplyr::filter(a1_t == a1_d & a2_t == a2_d))
```

`r count_raw_bim_merge_pos_a_match` rows have the same allele 1 and allele
2 - the vast majority but not all. 

The increase in overlap when mapping by position instead of RSID is partially
driven by non-RSID variant names in DAISY:

```{r}
head(raw_bim_merge_pos %>% dplyr::filter(id_t != id_d))
```

Increase also due to the duplicate positions in DAISY with different variant
names (so the join wasn't one-to-one):

```{r}
raw_bim_d_pos_dup_list <- raw_bim_d %>%
  dplyr::filter(duplicated(pos))
raw_bim_d_pos_dups <- raw_bim_d %>%
  dplyr::filter(pos %in% raw_bim_d_pos_dup_list$pos)

raw_bim_d_pos_dups_a_check <- raw_bim_d_pos_dups %>%
  group_by(pos) %>%
  summarize(a1_count = n_distinct(a1),
            a2_count = n_distinct(a2))
count_raw_bim_d_pos_dups_a_same <- sum(
  raw_bim_d_pos_dups_a_check$a1_count == 1 &
    raw_bim_d_pos_dups_a_check$a2_count == 1)

head(raw_bim_d_pos_dups)

```

There are `r nrow(raw_bim_d_pos_dups)` rows with variants with duplicate
positions in the raw DAISY data. Of these rows,
`r count_raw_bim_d_pos_dups_a_same` have the same ref and alt as well. 

**Venn Diagram:**

*To note: numbers don't exactly add up to totals due to duplicate positions*
*present in both datasets.*

```{r}
raw_bim_t <- raw_bim_t %>%
  dplyr::mutate(chrpos = paste0(chr, ":", pos))
raw_bim_d <- raw_bim_d %>%
  dplyr::mutate(chrpos = paste0(chr, ":", pos))

ggvenn(x <- list("TEDDY chr:pos" = raw_bim_t$chrpos, "DAISY chr:pos" = raw_bim_d$chrpos),
       c("TEDDY chr:pos", "DAISY chr:pos"), auto_scale = T)
```

## **Compare Variant MAF**

### **All Variants**

First, calculate MAF in PLINK2:

```{zsh, eval = F}
# Run once, then reload results
plink2 --bfile Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_SNP_exome_masked \
  --freq --nonfounders \
  --out Maternal_Protection/data/genetics/teddy/exome/Norris_SNP_exome_masked_freq

plink2 --bfile DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/clean/daisyexome \
  --freq --nonfounders \
  --out DAISY/genetics/daisy_ask_genetics/daisyask_exome_array/clean/daisyexome_freq

```

Show correlation plot of MAF for all variants:

```{r fig.show='hold', out.width='50%'}
# Need to map position informatio from ID for each study
raw_maf_t <- inner_join(raw_maf_t, raw_bim_t, by = c("ID" = "id"))
raw_maf_d <- inner_join(raw_maf_d, raw_bim_d, by = c("ID" = "id"))

raw_bim_maf_merge_pos <- inner_join(raw_maf_t, raw_maf_d, by = c("chr", "pos"),
                                suffix = c("_t", "_d"))

ggplot(raw_bim_maf_merge_pos %>% dplyr::filter(chr != 6), aes(ALT_FREQS_t, ALT_FREQS_d)) +
  geom_point() +
  xlim(0,1) + 
  ylim(0,1) +
  coord_fixed() +
  labs(title = "All Chr Except Chr6")

ggplot(raw_bim_maf_merge_pos %>% dplyr::filter(chr == 6), aes(ALT_FREQS_t, ALT_FREQS_d)) +
  geom_point() +
  xlim(0,1) + 
  ylim(0,1) +
  coord_fixed() +
  labs(title = "Chr6 Only")
```

*To note: DAISY's hard cutoff at 0.5 likely indicates used of a PLINK1.9*
*command run without --keep-allele-order.*

On chromosome 6, seems like there are a lot of variants with very low frequency
in TEDDY that have higher frequency in DAISY. Does that mean anything for us?

### **Common Variants**

Defining common variants as those with MAF > 5% in either study.

```{r}
raw_bim_maf_merge_pos_filt <- raw_bim_maf_merge_pos %>%
  dplyr::filter(ALT_FREQS_t >= 0.05 | ALT_FREQS_d >= 0.05)
count_raw_bim_maf_merge_pos_filt <- nrow(raw_bim_maf_merge_pos_filt)
```

This reduces variant list from `r count_raw_bim_merge_pos` to
`r count_raw_bim_maf_merge_pos_filt`.

Make density plot of the MAF difference (TEDDY - DAISY):

```{r fig.show='hold', out.width='50%'}
raw_bim_maf_merge_pos_filt <- raw_bim_maf_merge_pos_filt %>%
  dplyr::mutate(maf_diff = ALT_FREQS_t - ALT_FREQS_d)

ggplot(raw_bim_maf_merge_pos_filt %>% dplyr::filter(chr != 6), aes(x = maf_diff)) +
  geom_density() +
  labs(title = "All Chr Except Chr6")

ggplot(raw_bim_maf_merge_pos_filt %>% dplyr::filter(chr == 6), aes(x = maf_diff)) +
  geom_density() +
  labs(title = "Chr6 Only")

```

Density plot maybe not best visualization, but shows that there is little
overall difference between the studies.

## **Compare Overlap with GRS2 SNPs**

Even though it was imputed data for both studies that was/will be used to
calculate the GRS2, want to see what the overlap of unimputed SNPs is for both
studies, especially since saw additional typed SNPs found in DAISY that weren't
found in TEDDY.

### **Map by RSID**

*To note: the map is in hg38, so this check will only map by RSID.*

```{r}
map_tk_raw_bim_t <- left_join(map_tk %>% dplyr::select(COMPONENT, RSID),
                              raw_bim_t %>% dplyr::select(id, chrpos, a1, a2),
                              by = c("RSID" = "id"))
map_tk_raw_bim_d <- left_join(map_tk %>% dplyr::select(COMPONENT, RSID),
                              raw_bim_d %>% dplyr::select(id, chrpos, a1, a2),
                              by = c("RSID" = "id"))
map_tk_raw_merge <- inner_join(map_tk_raw_bim_t, map_tk_raw_bim_d,
                               by = c("COMPONENT", "RSID"),
                               suffix = c("_t", "_d"))

head(map_tk_raw_merge, n = 100)
```

Even with the DAISY data having less consistent RSID naming, mapping by RSID
for DAISY has `r sum(!is.na(map_tk_raw_bim_d$chrpos))` overlapping SNPs while
TEDDY only has `r sum(!is.na(map_tk_raw_bim_t$chrpos))`. DAISY also has 3 of the
HLA-DQ SNPs present, while TEDDY only has 1; 2 of the HLA SNPs while TEDDY has 1.

This explains the different genotyped SNPs in the TOPMed imputation results
that I saw present in DAISY but absent in TEDDY when trying to calculate th
GRS2. Also discussed with Randi this paper calculating risk scores including
the GRS2 in TEDDY (PMID 32770166)[https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7556983/]
which could explain why TEDDY is missing the HLA-DQ SNP tags. See supplemental
Table 7 for list of HLA-DQ SNPs excluded from the calculation due to the TEDDY
study inclusion criteria.

### **Map by hg19 Position**

First, going to use biomaRt to convert RSIDs from GRS2 file to hg19 chr:pos.
Then, re-map the above SNPs and see if more matches are found.

*To show table, arbitrarily removing one entry for those variants that map*
*twice due to duplicate positions.*

```{r}
# Get mart for hg19
mart <- useMart("ENSEMBL_MART_SNP", dataset = "hsapiens_snp", host = 'https://grch37.ensembl.org')

# Define your SNPs of interest
map_tk <- map_tk %>%
  dplyr::mutate(chr = gsub(":.+$", "", POSITION_BRAVO),
                pos = gsub("^.+:", "", POSITION_BRAVO))
snps <- map_tk$RSID
# Shouldn't search chr:pos as well since using hg19 mart but only have chr:pos
# in hg38. Just do chr, and if lose any RSIDs will check those by hand
chr <- map_tk$chr
# pos <- map_tk$pos

# Query to retrieve nearby genes for the SNP
result <- getBM(
  attributes = c(
    "refsnp_id",
    "chr_name",
    "chrom_start",
    "chrom_end",
    "allele"),
  filters = list(
    snp_filter = snps,
    chr_name = chr),
  mart = mart)

# Have a lot of additional matches. Filter first by chromosome
result <- result %>%
  dplyr::rename(pos_hg19 = chrom_start) %>%
  dplyr::select(-chrom_end)

# n_distinct(result$refsnp_id)  # 67, got result for all!

map_tk_hg19 <- inner_join(
  map_tk %>% dplyr::select(-chr, -pos), result,
  by = c("RSID" = "refsnp_id"))

# Finally, join with data
map_tk_hg19_raw_bim_t <- left_join(
  map_tk_hg19 %>% dplyr::select(COMPONENT, RSID, chr_name, pos_hg19),
  raw_bim_t %>% dplyr::select(id, chr, pos, a1, a2,),
  by = c("pos_hg19" = "pos"))
map_tk_hg19_raw_bim_d <- left_join(
  map_tk_hg19 %>% dplyr::select(COMPONENT, RSID, chr_name, pos_hg19),
  raw_bim_d %>% dplyr::select(id, chr, pos, a1, a2),
  by = c("pos_hg19" = "pos"))
map_tk_hg19_raw_merge <- inner_join(
  map_tk_hg19_raw_bim_t,
  map_tk_hg19_raw_bim_d,
  by = c("COMPONENT", "RSID"),
  suffix = c("_t", "_d"))

map_tk_hg19_raw_merge_no_dups <- map_tk_hg19_raw_merge %>%
  dplyr::filter(!duplicated(pos_hg19_d))

head(map_tk_hg19_raw_merge_no_dups, n = 100)
```

No additional HLA-DQ or HLA SNPs matched in either dataset. However, looks like
multiple DAISY "exm" IDs matched in non-HLA. Matching by hg19 position, DAISY
has `r sum(!is.na(map_tk_hg19_raw_merge_no_dups$id_d))` overlapping SNPs while
TEDDY has the same amount as mapped by RSID:
`r sum(!is.na(map_tk_hg19_raw_merge_no_dups$id_t))`

Again, in this subset of SNPs that are on the GRS2,  many of the "exm" DAISY
variants also have RSID named variants at the same position. Since our TOPMed
imputation pipeline removes duplicates, we likely don't need to do anything,
but Randi may reach out to DAISY about it.

```{r}
d_dups <- map_tk_hg19_raw_bim_d %>% dplyr::filter(duplicated(pos_hg19))
```

Do the duplicate entries have the same MAF?

```{r}
raw_maf_d %>% dplyr::filter(pos %in% d_dups$pos_hg19)
```

Yes, except for position 1610551 on chr20 that is slightly different.

### **Compare Variant MAF for GRS2 SNPs Mapped by hg19 Position**

*To note: for this comparison, arbitrarily removing one of the position*
*duplicates since frequencies nearly identical.*

```{r fig.show='hold', out.width='50%'}
# Add MAF to merge
temp1 <- left_join(
  map_tk_hg19_raw_merge_no_dups %>% dplyr::select(COMPONENT, RSID, chr_name_t, id_t),
  raw_maf_t %>% dplyr::select(ID, ALT_FREQS, REF, ALT),
  by = c("id_t" = "ID"))
temp2 <- left_join(
  map_tk_hg19_raw_merge_no_dups %>% dplyr::select(COMPONENT, RSID, chr_name_d, id_d),
  raw_maf_d %>% dplyr::select(ID, ALT_FREQS, REF, ALT),
  by = c("id_d" = "ID"))
  
map_tk_hg19_raw_merge_mod <- inner_join(
  temp1, temp2, by = c("COMPONENT", "RSID"), suffix = c("_t", "_d"))

ggplot(map_tk_hg19_raw_merge_mod %>% dplyr::filter(chr_name_t != 6), aes(ALT_FREQS_t, ALT_FREQS_d)) +
  geom_point() +
  xlim(0,1) + 
  ylim(0,1) +
  coord_fixed() +
  labs(title = "All Chr Except Chr6")

ggplot(map_tk_hg19_raw_merge_mod %>% dplyr::filter(chr_name_t == 6), aes(ALT_FREQS_t, ALT_FREQS_d)) +
  geom_point() +
  xlim(0,1) + 
  ylim(0,1) +
  coord_fixed() +
  labs(title = "Chr6 Only")
```

# **2. Compare TOPMed Imputed Data**

*Still TODO.*

# **Session Info**

```{r session_info}
sessionInfo()
```
