---
title: "1_fam_hist_grs2x_teddy_daisy"
author: "NV"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection/reports/risk_scores/archive")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"))) #setting up pathway

library(tidyverse)
library(table1)
library(ggpubr)
library(haven)
```

# **Overview**

This is a summary report of DAISY and TEDDY - comparing and visualizing GRS2x and phenotype before including inclusion and exclusion criteria.

# **Load Data**


```{r, message = F}

# Read in research GRS2x results

grs2x_daisy <- read_csv(
  "DAISY/genetics/daisy_ask_genetics/risk_scores/data/grs2x/tm_r3_imp/score_results.csv")

grs2x_teddy <- read_csv(
  "Immunogenetics_T1D/genetics/teddy_r01/risk_scores/grs2x/score_results_imputed.csv")

# Read in pheno files

pheno_daisy <-read_sas("DAISY/pheno/ivyomicssample.sas7bdat")


pheno_teddy<- read_sas("Immunogenetics_T1D/raw/teddy_r01/2025-09-17/rj_immuno_all_demo_masked_v2.sas7bdat")
```




# **Table 1s**

Preaparing dataset -cleaning column names
```{r}
## PHENO DAISY

# cleaning phenotype file in DAISY
pheno_daisy <- pheno_daisy %>%
  dplyr::filter(T1Dgroup != "") %>%               # remove rows with blank T1Dgroup
  group_by(ID) %>%
  slice_max(DOVISIT, with_ties = FALSE) %>% # keep latest visit per ID
  ungroup()

# Check number of unique IDs 
#length(unique(pheno_daisy$ID))- it is 178, which indicates overall cases in Daisy

# Prep data for whole section
pheno_daisy <- pheno_daisy %>%
  dplyr::mutate(T1D = factor(T1Dgroup,
                             levels = c("T1D case", "T1D control"),
                             labels = c("Case", "Control")),
                Sex = SEX)

## PHENO TEDDY
pheno_teddy <-pheno_teddy%>%
  rename(ID = RJohnson_Immuno_anc_MaskID)

pheno_teddy <- pheno_teddy %>%
  mutate(ID = as.character(ID))

pheno_teddy <- pheno_teddy %>%
  dplyr::mutate(
    T1D  = factor(t1d,
                  levels = c(1, 0),
                  labels = c("Case", "Control")))

## GRS2X in daisy

# creating ID column in grs2x_daisy from IID column
grs2x_daisy$ID <- sub("_.*", "", grs2x_daisy$IID)

grs2x_daisy <- grs2x_daisy %>%
  rename(GRS2x="t1dgrs2-luckett25_total" )

## GRS2x in teddy

grs2x_teddy$ID <- sub(".*_", "", grs2x_teddy$IID)

grs2x_teddy <- grs2x_teddy %>%
  rename(GRS2x="t1dgrs2-luckett25_total" )


```



### **DAISY**


```{r}
# Recode DMfamily into 4-level factor: none, mom, dad, sibling
pheno_daisy$fdr_4level <- factor(
  ifelse(pheno_daisy$DMfamily == "" | is.na(pheno_daisy$DMfamily), "none",
         ifelse(pheno_daisy$DMfamily == "mom", "mom",
                ifelse(pheno_daisy$DMfamily == "dad", "dad", "sibling"))),
  levels = c("none", "mom", "dad", "sibling")
)


# Set reference level to "none"
pheno_daisy$fdr_4level <- relevel(pheno_daisy$fdr_4level, ref = "none")



# table(pheno_daisy$DMfamily, pheno_daisy$fdr_4level)

# Create descriptive table by T1D 
table1(~ fdr_4level + Sex | T1D, data = pheno_daisy, overall = "Total")

# Find missing IDs in another dataset
d_miss <- setdiff(pheno_daisy$ID, grs2x_daisy$ID)


```

When merging with the risk scores results, one individual is removed:
`r d_miss`. This individual does not have genetic data.

### **TEDDY**


This table is from Phenotype file (with total participants - 8676)
```{r}


pheno_teddy <- pheno_teddy %>%
  dplyr::mutate(
    fdr_4level = case_when(
      family_mem_screen == 0 ~ "none",
      family_mem_screen == 5 ~ "mom",      # only mom
      family_mem_screen == 6 ~ "dad",      # only dad
      family_mem_screen == 7 ~ "sibling",  # only sibling
      TRUE ~ "none"  # all multi-codes (1,2,3,4) → none
    )
  )



# Make fdr_4level an ordered factor with "none" as reference
pheno_teddy$fdr_4level <- factor(pheno_teddy$fdr_4level,
                                 levels = c("none", "mom", "dad", "sibling"))

# Create descriptive table by T1D group 
table1(~ fdr_4level + Sex | T1D, data = pheno_teddy, overall = "Total")

# Find missing IDs in another dataset
d_miss <- setdiff(pheno_teddy$ID, grs2x_teddy$ID)

length(d_miss)

length(d_miss) / nrow(pheno_teddy)

```



Here, information of 461 is missing after merging, the table below summarizes the cases and controls after merging phenotype and score data in TEDDY.

```{r}

grs2x_teddy_pheno <- inner_join(grs2x_teddy, pheno_teddy, by = c("ID"))

# Create descriptive table by T1D group 
table1(~ fdr_4level + Sex | T1D, data = grs2x_teddy_pheno, overall = "Total")


```


# **Research GRS2x**


## **Plots**


```{r}
# Prep data for whole section
grs2x_daisy_pheno <- inner_join(grs2x_daisy, pheno_daisy, by = c("ID"))



# Get min/max for plots
min <- floor(min(min(grs2x_daisy_pheno$GRS2x), min(grs2x_teddy_pheno$GRS2x)))
max <- ceiling(max(max(grs2x_daisy_pheno$GRS2x), max(grs2x_teddy_pheno$GRS2x)))

```


### **By T1D Case/Control**

```{r fig.show='hold', out.width='50%'}
ggplot(grs2x_daisy_pheno, aes(x = T1D, y = GRS2x, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

ggplot(grs2x_teddy_pheno, aes(x = T1D, y = GRS2x, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

```


### **By Family History**

```{r fig.show='hold', out.width='50%'}
ggplot(grs2x_daisy_pheno, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

ggplot(grs2x_teddy_pheno, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")
```

### **By Family History - doing this for our reference not using in the study**

```{r fig.show='hold', out.width='50%'}
## PHENO DAISY
pheno1_daisy <-read_sas("DAISY/pheno/ivyomicssample.sas7bdat")

# cleaning phenotype file in DAISY
pheno1_daisy <- pheno1_daisy %>%
  group_by(ID) %>%
  slice_max(DOVISIT, with_ties = FALSE) %>% # keep latest visit per ID
  ungroup()


# Prep data for whole section
pheno1_daisy <- pheno1_daisy %>%
  dplyr::mutate(Sex = SEX)
# Recode DMfamily into 4-level factor: none, mom, dad, sibling
pheno1_daisy$fdr_4level <- factor(
  ifelse(pheno1_daisy$DMfamily == "" | is.na(pheno1_daisy$DMfamily), "none",
         ifelse(pheno1_daisy$DMfamily == "mom", "mom",
                ifelse(pheno1_daisy$DMfamily == "dad", "dad", "sibling"))),
  levels = c("none", "mom", "dad", "sibling")
)


# Set reference level to "none"
pheno1_daisy$fdr_4level <- relevel(pheno1_daisy$fdr_4level, ref = "none")


# table(pheno_daisy$fdr_4level)
# table(pheno_daisy$DMfamily, pheno_daisy$fdr_4level)



# Prep data for whole section
grs2x_daisy_pheno1 <- inner_join(grs2x_daisy, pheno1_daisy, by = c("ID"))



# Get min/max for plots
min <- floor(min(min(grs2x_daisy_pheno1$GRS2x)))
max <- ceiling(max(max(grs2x_daisy_pheno1$GRS2x)))



ggplot(grs2x_daisy_pheno1, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

ggplot(grs2x_teddy_pheno, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")
```
This is a DAISY plot for all participants (i.e., before filtering T1D cases, controls, and removing NAs). There is not much difference between this plot and the previous one, but the median for ‘Dad’ shows a slight increase, while the medians for the other FDR categories (‘None’, ‘Mom’, and ‘Sibling’) decreased.


### **By T1D Case/Control & Family History**

```{r fig.show='hold', out.width='50%'}
ggplot(grs2x_daisy_pheno, aes(x = fdr_4level, y = GRS2x, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))

ggplot(grs2x_teddy_pheno, aes(x = fdr_4level, y = GRS2x, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))
```



### **By T1D Case/Control & Family History, Faceted by Sex**

```{r}
ggplot(grs2x_daisy_pheno, aes(x = fdr_4level, y = GRS2x, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D",
       x = "FDR") +
  facet_wrap(~ Sex) + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
             geom = "text", position = position_dodge(width = 0.6))

ggplot(grs2x_teddy_pheno, aes(x = fdr_4level, y = GRS2x, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D",
       x = "FDR") +
  facet_wrap(~ Sex) + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
             geom = "text", position = position_dodge(width = 0.6))
```
# **Session Info**

```{r session_info}
sessionInfo()
```
