---
title: "Analytical Data Summary"
author: "NV, Updated by SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection/reports/risk_scores")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"))) #setting up pathway

library(tidyverse)
library(table1)
library(ggpubr)
library(haven)
library(dplyr)
library(tidyr)
library(janitor)
library(ggplot2)
library(knitr)
library(khroma)

```

# **Load Data**


```{r, message = F}
# Read in pheno files
### DAISY
pheno_daisy_all <- readRDS(
  "Maternal_Protection/data/daisy/famhist_prs/pheno_all_ia.rds")  # pheno for hypoth 1
# d_all_t1d <- readRDS(
#   "Maternal_Protection/data/daisy/famhist_prs/pheno_all_t1d.rds")
# d_all_prog <- readRDS(
#   "Maternal_Protection/data/daisy/famhist_prs/pheno_all_prog.rds")
pheno_daisy_ctrls <- readRDS(
  "Maternal_Protection/data/daisy/famhist_prs/pheno_ctrls_ia.rds")  # pheno for hypoth 1
# d_ctrls_t1d <- readRDS(
#   "Maternal_Protection/data/daisy/famhist_prs/pheno_ctrls_t1d.rds")
# d_ctrls_prog <- readRDS(
#   "Maternal_Protection/data/daisy/famhist_prs/pheno_ctrls_prog.rds")

### TEDDY
pheno_teddy_all <- readRDS(
  "Maternal_Protection/data/teddy/famhist_prs/pheno_all.rds")  # pheno for hypoth 1
# t_all_prog <- readRDS(
#   "Maternal_Protection/data/teddy/famhist_prs/pheno_all_prog.rds")
pheno_teddy_ctrls <- readRDS(
  "Maternal_Protection/data/teddy/famhist_prs/pheno_ctrls_ia.rds")
# t_ctrls_t1d <- readRDS(
#   "Maternal_Protection/data/teddy/famhist_prs/pheno_ctrls_t1d.rds")
# t_ctrls_prog <- readRDS(
#   "Maternal_Protection/data/teddy/famhist_prs/pheno_ctrls_prog.rds")

# Read GRS2x score file (61SNPs)
grs2x_61<- read.csv(
  "DAISY/genetics/daisy_ask_genetics/risk_scores/data/grs2x/tm_r3_imp/hwe_in_all/score_results_61_geno.csv")

# Pheno files with HLA group information not included in analytical pheno files
hla_daisy <- readRDS(
  "DAISY/snapshots_of_data/deidentified-snapshots-RDS/deid-casedef_daisy07172024.rds") %>%
  dplyr::select(ID, HLAGRP)
hla_teddy <- sas7bdat::read.sas7bdat(
  "Immunogenetics_T1D/raw/teddy_r01/2025-09-17/rj_immuno_all_demo_masked_v2.sas7bdat") %>%
  dplyr::select(RJohnson_Immuno_anc_MaskID, hla_category)

# Poster analysis files
poster_daisy <- readRDS(
  "Maternal_Protection/data/daisy/risk_score_analysis/pheno_ia_with_grs2_sd_eur_dr_filt_fdr_4level.rds")

```

# **Table 1s**
### **DAISY**

```{r}

# Combine datasets and add Group column
pheno_daisy <- bind_rows(
  pheno_daisy_all %>% mutate(Group = "All"),
  pheno_daisy_ctrls %>% mutate(Group = "Controls"))

# Set variable labels
label(pheno_daisy$GRS2x) <- "GRS2x Mean(SD)[Min,Max]"
label(pheno_daisy$IA) <- "IA Case, N (%)"
label(pheno_daisy$T1D) <- "T1D Case, N (%)"
label(pheno_daisy$sex) <- "Sex, N (%)"
label(pheno_daisy$fdr_4level) <- "FDR, N (%)"

# Create table
table1(~ GRS2x + IA + T1D + sex + fdr_4level | Group,
       data = pheno_daisy,
       overall = FALSE)


```


### **TEDDY**

```{r}

# Combine datasets and add group column
pheno_teddy <- bind_rows(
  pheno_teddy_all %>% mutate(Group = "All"),
  pheno_teddy_ctrls %>% mutate(Group = "Controls"))


# Set variable labels
label(pheno_teddy$GRS2x) <- "GRS2x Mean(SD)[Min,Max]"
label(pheno_teddy$IA) <- "IA Case, N (%)"
label(pheno_teddy$T1D) <- "T1D Case,N (%)"
label(pheno_teddy$sex) <- "Sex,N (%)"
label(pheno_teddy$fdr_4level) <- "FDR, N (%)"
label(pheno_teddy$IA) <- "IA Case, N (%)"
label(pheno_teddy$T1D) <- "T1D Case, N (%)"


# Create table without Total, showing only Yes for IA/T1D
table1(~ GRS2x + IA + T1D + sex + fdr_4level | Group,
       data = pheno_teddy,
       overall = FALSE)
```



# **Box plots**

```{r}
###For 61SNPs GRS2x plots
pheno_daisy_all <- pheno_daisy_all %>%
  inner_join(hla_daisy, by = c("ID"))
pheno_daisy_ctrls <- pheno_daisy_ctrls %>%
  inner_join(hla_daisy, by = c("ID"))

# Join grs2x_61 to pheno_daisy_all by IID
pheno_daisy_all_61 <- pheno_daisy_all %>%
  dplyr::select(ID, fdr_4level, HLAGRP) %>%
  left_join(
    grs2x_61 %>% dplyr::mutate(ID = gsub("^.+_", "", IID)),
    by = "ID") %>%
  rename(GRS2x = t1dgrs2.luckett25_total)

# Join grs2x_61 to pheno_daisy_ctrls by IID
pheno_daisy_ctrls_61 <- pheno_daisy_ctrls %>%
  dplyr::select(ID, fdr_4level, HLAGRP) %>%
  left_join(
    grs2x_61 %>% dplyr::mutate(ID = gsub("^.+_", "", IID)),
    by = "ID") %>%
  rename(GRS2x = t1dgrs2.luckett25_total)

# Common min/max across all DAISY datasets(67 and 61SNPs)
common_min <- floor(min(
  pheno_daisy_all$GRS2x, pheno_daisy_ctrls$GRS2x,
  pheno_daisy_all_61$GRS2x, pheno_daisy_ctrls_61$GRS2x,
  poster_daisy$GRS2, na.rm = TRUE))
common_max <- ceiling(max(
  pheno_daisy_all$GRS2x, pheno_daisy_ctrls$GRS2x,
  pheno_daisy_all_61$GRS2x, pheno_daisy_ctrls_61$GRS2x,
  poster_daisy$GRS2, na.rm = TRUE))

```

### **DAISY (TM-R3 GRS2x, 67 SNPs)**

```{r fig.show='hold', out.width='50%'}

#plot
ggplot(pheno_daisy_all, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-All",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

ggplot(pheno_daisy_ctrls, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-Control",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

```

GRS2x score is highest in siblings when compared to all other FDRs in
DAISY-ALL(Cases+control), and in case of only DAISY-control, GRS2x for none and
sibilings are on similar range.


### **DAISY (TM-R3 GRS2x, 61 SNPs)**

```{r fig.show='hold', out.width='50%'}

#plot
ggplot(pheno_daisy_all_61, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-All(61SNPs)",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

ggplot(pheno_daisy_ctrls_61, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-Control(61SNPs)",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

```

Comparing GRS2x scores with 67SNPs and 61SNPs in DAISY  for ALL vs control
group, the pattern of distribution looks similar but grs2x score is higher in
case of 67 SNPs (which explain the weight contributed by 6SNPs).

### **DAISY (TM-R3 GRS2x, 67 SNPs, DR3/3, 3/4, 4/4, & 4/X)**

```{r fig.show='hold', out.width='50%'}

pheno_daisy_all_filt <- pheno_daisy_all %>%
  dplyr::filter(HLAGRP == "DR3/3" | HLAGRP == "DR3/4" | HLAGRP == "DR4/4" | HLAGRP == "DR4/X")
pheno_daisy_ctrls_filt <- pheno_daisy_ctrls %>%
  dplyr::filter(HLAGRP == "DR3/3" | HLAGRP == "DR3/4" | HLAGRP == "DR4/4" | HLAGRP == "DR4/X")

ggplot(pheno_daisy_all_filt, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-All",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

ggplot(pheno_daisy_ctrls_filt, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-Control",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

```


### **DAISY (TM-R3 GRS2x, 61 SNPs, DR3/3, 3/4, 4/4, & 4/X)**


```{r fig.show='hold', out.width='50%'}

pheno_daisy_all_61_filt <- pheno_daisy_all_61 %>%
  dplyr::filter(HLAGRP == "DR3/3" | HLAGRP == "DR3/4" | HLAGRP == "DR4/4" | HLAGRP == "DR4/X")
pheno_daisy_ctrls_61_filt <- pheno_daisy_ctrls_61 %>%
  dplyr::filter(HLAGRP == "DR3/3" | HLAGRP == "DR3/4" | HLAGRP == "DR4/4" | HLAGRP == "DR4/X")

ggplot(pheno_daisy_all_61_filt, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-All(61 SNPs)",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

ggplot(pheno_daisy_ctrls_61_filt, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-Control(61 SNPs)",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

```





### **DAISY (Poster, 1000G GRS2, 61 SNPs, DR3/3, 3/4, 4/4, & 4/X)**


```{r, fig.show='hold', out.width='50%'}

ggplot(poster_daisy, aes(x = fdr_4level, y = GRS2, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY-All(61 SNPs)",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

```



### **TEDDY**


```{r fig.show='hold', out.width='50%'}

ggplot(pheno_teddy_all, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY-All",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

ggplot(pheno_teddy_ctrls, aes(x = fdr_4level, y = GRS2x, fill = fdr_4level)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY-Control",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(common_min, common_max, by = 2), limits = c(common_min, common_max)) +
  stat_summary(fun.data = function(x) data.frame(y = common_min, label = length(x)), 
               geom = "text")

```



# **Session Info**

```{r session_info}
sessionInfo()
```