---
title: "fam_hist_teddy_daisy"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection/reports")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection"))

library(tidyverse)
library(table1)
library(ggpubr)

```

# **Overview**

Report shows relationship between risk scores (AA 7 SNP PRS, research GRS2) and
family history in TEDDY and DAISY. Risk scores calculated for each in:

+ **TEDDY**
  + explore_matprot/teddy_risk_scores/calc_teddy_risk_scores.Rmd
  + AA 7 SNP PRS: TOPMed R3 imputed data with Michigan HLA imputed data
  + Research GRS2: 1000G phase 3 version 5 EUR imputed data, 1000G hg19 score
    list from HLA PRS toolkit
+ **DAISY**
  + DAISY/genetics/risk_scores/code/calc_daisy_grs2.Rmd & calc_daisy_prs_aa.Rmd
  + AA 7 SNP PRS: TOPMed R2 imputed data with Michigan HLA imputed data
  + Research GRS2: TOPMed R2 imputed data with Michigan HLA imputed data, TOPMed
    hg38 score list from HLA PRS toolkit
      + 6/7/24 added a version from 1000G imputed data, all 67 SNPs
      + 6/10/24 added a version from 1000G imputed data, 61 SNPs to match TEDDY

*To note: clinical GRS2 scores currently not included.*

# **Load Data**

*To note:* Using same DAISY pheno file here as used by Alex for looking into
DAISY DNAm FDR*T1D interactions. That file has 174 individuals. The IVYOmics SAS
phenotype file lists 178 individuals. 1 of the extra four in the IVYOmics file
does not have genetic data. The other do not have DNAm data.

Using TEDDY subject_pheno.csv file created for use with DNAm data.

```{r, message = F}
# Read in AA 7 SNP PRS results
aa_daisy <- read_csv(
  "../DAISY/genetics/risk_scores/data/prs/daisy_aa_7snp_prs_results.csv")
aa_teddy <- read_delim(
  "data/teddy_risk_scores/aa_7snp/scoring/teddy_aa_7snp_prs.sscore")

# Read in research GRS2 results
grs2_daisy <- read_csv(
  "../DAISY/genetics/risk_scores/data/grs2/tm_imp/daisy_grs2_results.csv")
grs2_daisy_kg <- read_tsv(
  "../DAISY/genetics/risk_scores/data/grs2/1000g/grs2_score.txt")
grs2_daisy_kg_t <- read_tsv(
  "../DAISY/genetics/risk_scores/data/grs2/1000g_teddy_snps/grs2_score.txt")
grs2_teddy <- read_tsv(
  "data/teddy_risk_scores/grs2/scoring_1000g/grs2_score.txt")

# Read in pheno files
# Instead of IVYomics sas7bdat pheno, using same pheno file as used by Alex
# for looking into DAISY DNAm FDR*T1D interactions
pheno_daisy <- read_csv("data/analysis_files/T1D_IVYclinicalData.csv")

pheno_teddy <- read.csv("data/TEDDYarrays/subject_pheno.csv")

```

# **Table 1s**

### **DAISY**

Modify DAISY pheno to have three-level FDR variable, and check that sample
sizes are up to date by checking if match reports/FDR_interaction_summary.html.

*TO NOTE: I updated "Dad" to "Dad/Sib" because that's what it was actually*
*counting - Dads + Siblings.*

```{r}
# FDR 3-level variable and re-coding of two IDs to Dad is based off of previous
# script at: code/Step8.1_logistic_fdr_interaction.R
pheno_daisy$fdr_3level = factor(
  ifelse(is.na(pheno_daisy$DMfamily), "None",
         ifelse(pheno_daisy$DMfamily == "mom/dad" |
                  pheno_daisy$DMfamily == "mom", "Mom","Dad/Sib")), 
  ordered = FALSE)
pheno_daisy$fdr_3level = relevel(pheno_daisy$fdr_3level, ref = "None")
pheno_daisy$fdr_3level[which(pheno_daisy$ID %in% c("00275-0","00459-0"))] <- "Dad/Sib"
# table(pheno_daisy$fdr_3level)
# table(pheno_daisy$DMfamily, pheno_daisy$fdr_3level)

table1(~ fdr_3level + SEX | T1Dgroup, data = pheno_daisy, overall = "Total")

d_miss <- setdiff(pheno_daisy$ID, aa_daisy$ID)
```
<br>

This matches the report, with a total of `r nrow(pheno_daisy)` individuals.
However, when merging with the risk scores results, one individual is removed:
`r d_miss`. This individual does not have genetic data.

With that individual missing & variables updated to be consistent with TEDDY
(and vice-versa), the updated Table1 is:

```{r}
# Prep data for whole section
pheno_daisy <- pheno_daisy %>%
  dplyr::mutate(T1D = factor(T1Dgroup,
                             levels = c("T1D case", "T1D control"),
                             labels = c("Case", "Control")),
                Sex = SEX)

pheno_teddy <- pheno_teddy %>%
  dplyr::mutate(T1D = factor(t1d,
                             levels = c(1, 0),
                             labels = c("Case", "Control")),
                fdr_3level = factor(case_when(
                  fdr_3level == "mom" ~ "Mom",
                  fdr_3level == "dadsib" ~ "Dad/Sib",
                  fdr_3level == "none" ~ "None"
                )))
pheno_teddy$fdr_3level = relevel(pheno_teddy$fdr_3level, ref = "None")

aa_daisy_pheno <- inner_join(aa_daisy, pheno_daisy, by = c("ID"))
aa_teddy_pheno <- inner_join(aa_teddy, pheno_teddy, by = c("IID" = "maskid")) %>%
  dplyr::mutate(AA_PRS = prs_aa_7snp_SUM)

table1(~ fdr_3level + Sex | T1D, data = aa_daisy_pheno, overall = "Total")
```


### **TEDDY**

T1D case = 1, control = 0.

```{r}
table1(~ fdr_3level + Sex | T1D, data = aa_teddy_pheno, overall = "Total")
```


# **AA 7 SNP PRS**

## **Plots**

```{r}
# Get min/max for plots
min <- floor(min(min(aa_daisy_pheno$AA_PRS), min(aa_teddy_pheno$AA_PRS)))
max <- ceiling(max(max(aa_daisy_pheno$AA_PRS), max(aa_teddy_pheno$AA_PRS)))

# Group size label position
label_pos <- min

```

### **By T1D Case/Control**

```{r fig.show='hold', out.width='50%'}

ggplot(aa_daisy_pheno, aes(x = T1D, y = AA_PRS, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

ggplot(aa_teddy_pheno, aes(x = T1D, y = AA_PRS, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

```

One-way ANOVA of AA_PRS ~ T1D case/control. Did not assume equal variances.

**DAISY:**

```{r}
oneway.test(AA_PRS ~ as.factor(T1D),
            data = aa_daisy_pheno,
            var.equal = F)
```

Significant diference between cases and controls.

**TEDDY:**

```{r}
oneway.test(AA_PRS ~ as.factor(T1D),
            data = aa_teddy_pheno,
            var.equal = F)
```

No significant difference between cases and controls.

### **By Family History**

```{r fig.show='hold', out.width='50%'}
ggplot(aa_daisy_pheno, aes(x = fdr_3level, y = AA_PRS)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

ggplot(aa_teddy_pheno, aes(x = fdr_3level, y = AA_PRS)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

```

One-way ANOVA of AA_PRS ~ 3-level FDR. Did not assume equal variances.

**DAISY:**

```{r}
oneway.test(AA_PRS ~ fdr_3level,
            data = aa_daisy_pheno,
            var.equal = F)
```

Significant difference between at least one of the FDR levels. Should I follow
up with pairwise tests?

**TEDDY:**

```{r}
oneway.test(AA_PRS ~ fdr_3level,
            data = aa_teddy_pheno,
            var.equal = F)
```

No significant difference between FDR levels. 

### **By T1D Case/Control & Family History**

```{r fig.show='hold', out.width='50%'}
ggplot(aa_daisy_pheno, aes(x = fdr_3level, y = AA_PRS, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))

ggplot(aa_teddy_pheno, aes(x = fdr_3level, y = AA_PRS, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))

```

Even when split by both T1D case/control and 3-level FDR, TEDDY has more
consistent AA_PRS scores across all groups than DAISY. This likely makes sense
due to the messy AA_PRS score distribution by case/control status (checked in
report used to generate scores). DAISY may be suggesting a survival bias since
the AA_PRS scores for Moms are much lower for cases. Not sure what it means
(if anything) that Dads control scores are similarly low.

### **By T1D Case/Control & Family History, Faceted by Sex**

```{r}
ggplot(aa_daisy_pheno, aes(x = fdr_3level, y = AA_PRS, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D",
       x = "FDR") +
  facet_wrap(~ Sex) + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
             geom = "text", position = position_dodge(width = 0.6))

ggplot(aa_teddy_pheno, aes(x = fdr_3level, y = AA_PRS, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D",
       x = "FDR") + 
  facet_wrap(~ Sex) +
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
             geom = "text", position = position_dodge(width = 0.6))
```

Not sure what to conclude. In TEDDY, there is only 1 T1D male T1D cases with a
Mom FDR!

## **Models**

*TODO: Randi - I couldn't remember if you said lm() or lme(). Using lm() here*
*as I wasn't sure what the random effects would be for lme().*

**DAISY All:**

```{r}
lm(AA_PRS ~ fdr_3level + PC1 + PC2 + PC3, data = aa_daisy_pheno)

```

**DAISY Controls Only:**

```{r}
aa_daisy_pheno_ctrl <- aa_daisy_pheno %>%
  dplyr::filter(T1D == "Control")

lm(AA_PRS ~ fdr_3level + PC1 + PC2 + PC3, data = aa_daisy_pheno_ctrl)
```

**TEDDY All:**

```{r}
lm(AA_PRS ~ fdr_3level + PC1 + PC2 + PC3, data = aa_teddy_pheno)

```

**TEDDY Controls Only:**

```{r}
aa_teddy_pheno_ctrl <- aa_teddy_pheno %>%
  dplyr::filter(T1D == "Control")

lm(AA_PRS ~ fdr_3level + PC1 + PC2 + PC3, data = aa_teddy_pheno_ctrl)
```

# **Research GRS2**

**TO NOTE: This section now compares the 1000G-calculated TEDDY and DAISY**
**scores, both with 61 SNPs to match TEDDY input.**

## **Plots**

```{r}
# Prep data for whole section
grs2_daisy_pheno <- inner_join(grs2_daisy, pheno_daisy, by = c("ID"))

grs2_daisy_pheno <- grs2_daisy_pheno %>%
  dplyr::mutate(T1D = factor(T1Dgroup,
                             levels = c("T1D case", "T1D control"),
                             labels = c("Case", "Control")))

grs2_daisy_kg_pheno <- inner_join(grs2_daisy_kg, pheno_daisy, by = c("ID"))

grs2_daisy_kg_pheno <- grs2_daisy_kg_pheno %>%
  dplyr::mutate(T1D = factor(T1Dgroup,
                             levels = c("T1D case", "T1D control"),
                             labels = c("Case", "Control")))

grs2_daisy_kg_t_pheno <- inner_join(grs2_daisy_kg_t, pheno_daisy, by = c("ID"))

grs2_daisy_kg_t_pheno <- grs2_daisy_kg_t_pheno %>%
  dplyr::mutate(T1D = factor(T1Dgroup,
                             levels = c("T1D case", "T1D control"),
                             labels = c("Case", "Control")))



grs2_teddy_pheno <- inner_join(grs2_teddy, pheno_teddy, by = c("ID" = "maskid"))

# Get min/max for plots
min <- floor(min(min(grs2_daisy_kg_t_pheno$GRS2), min(grs2_teddy_pheno$GRS2)))
max <- ceiling(max(max(grs2_daisy_kg_t_pheno$GRS2), max(grs2_teddy_pheno$GRS2)))

```

### **By T1D Case/Control**

```{r fig.show='hold', out.width='50%'}
ggplot(grs2_daisy_kg_t_pheno, aes(x = T1D, y = GRS2, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

ggplot(grs2_teddy_pheno, aes(x = T1D, y = GRS2, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

```


One-way ANOVA of GRS2 ~ T1D case/control. Did not assume equal variances.

**DAISY:**

```{r}
oneway.test(GRS2 ~ as.factor(T1D),
            data = grs2_daisy_kg_t_pheno,
            var.equal = F)
```

Significant difference between GRS2 in cases and controls.

**TEDDY:**

```{r}
oneway.test(GRS2 ~ as.factor(T1D),
            data = grs2_teddy_pheno,
            var.equal = F)
```

Also a significant difference between GRS2 cases and controls.

### **By Family History**

```{r fig.show='hold', out.width='50%'}
ggplot(grs2_daisy_kg_t_pheno, aes(x = fdr_3level, y = GRS2)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")

ggplot(grs2_teddy_pheno, aes(x = fdr_3level, y = GRS2)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text")
```

One-way ANOVA of GRS2 ~ 3-level FDR. Did not assume equal variances.

**DAISY:**

```{r}
anova_fdr_d_kg_t <- oneway.test(
  GRS2 ~ fdr_3level,
  data = grs2_daisy_kg_t_pheno,
  var.equal = F)
anova_fdr_d_kg_t
```

No significant differences between groups, although Mom median does appear
lower, but also has the smallest sample size.

**TEDDY:**

```{r}
anova_fdr_t <- oneway.test(
  GRS2 ~ fdr_3level,
  data = grs2_teddy_pheno,
  var.equal = F)
anova_fdr_t
```

Even less of a significant difference bewteen groups.

### **By T1D Case/Control & Family History**

```{r fig.show='hold', out.width='50%'}
ggplot(grs2_daisy_kg_t_pheno, aes(x = fdr_3level, y = GRS2, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))

ggplot(grs2_teddy_pheno, aes(x = fdr_3level, y = GRS2, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D",
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))
```

For DAISY, pattern appears to be similar to that of the AA_PRS, but less
extreme. TEDDY groups all look similar to each other.

### **By T1D Case/Control & Family History, Faceted by Sex**

```{r}
ggplot(grs2_daisy_kg_t_pheno, aes(x = fdr_3level, y = GRS2, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY",
       fill = "T1D",
       x = "FDR") +
  facet_wrap(~ Sex) + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
             geom = "text", position = position_dodge(width = 0.6))

ggplot(grs2_teddy_pheno, aes(x = fdr_3level, y = GRS2, fill = as.factor(T1D))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY",
       fill = "T1D",
       x = "FDR") +
  facet_wrap(~ Sex) + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
             geom = "text", position = position_dodge(width = 0.6))
```

In DAISY, unlike for AA_PRS where pattern with Mom with females and males was
visually similar, here Mom with female appears to be slightly lower than Mom
with male for both cases and controls. In TEDDY, it is hard to draw any
conclusion since there is only 1 Mom with male.


## **Models**

*TODO: Randi - I couldn't remember if you said lm() or lme(). Using lm() here*
*as I wasn't sure what the random effects would be for lme().*

**DAISY All:**

```{r}
# Add DAISY PCs to 1000g TEDDY SNPs results
pcs <- grs2_daisy %>% dplyr::select(ID, PC1, PC2, PC3)
grs2_daisy_kg_t_pheno_pcs <-
  inner_join(grs2_daisy_kg_t_pheno, pcs, by = c("ID"))


lm(GRS2 ~ fdr_3level + PC1 + PC2 + PC3, data = grs2_daisy_kg_t_pheno_pcs)

```

**DAISY Controls Only:**

```{r}
grs2_daisy_pheno_ctrl <- grs2_daisy_kg_t_pheno_pcs %>%
  dplyr::filter(T1D == "Control")

lm(GRS2 ~ fdr_3level + PC1 + PC2 + PC3, data = grs2_daisy_pheno_ctrl)
```

Intercept in controls only is lower than in all, which makes sense. Not exactly
sure how to interpret the big difference between the fdr_3levelDad/Sib
coefficient in all (0.113) compared to in controls only (-0.732). And at the
same time, fdr_3levelMom because slightly less negative in controls only.
Similar pattern as for the AA_PRS.

**TEDDY All:**

```{r}
lm(GRS2 ~ fdr_3level + PC1 + PC2 + PC3, data = grs2_teddy_pheno)

```

**TEDDY Controls Only:**

```{r}
grs2_teddy_pheno_ctrl <- grs2_teddy_pheno %>%
  dplyr::filter(T1D == "Control")

lm(GRS2 ~ fdr_3level + PC1 + PC2 + PC3, data = grs2_teddy_pheno_ctrl)
```

Same intercept pattern as DAISY, but both Dad/Sib and Mom because more positive
in controls only for TEDDY, which is a different pattern than seen for DAISY.

# **Write Out**

*For now, this section includes plots written out for grant to*
*Grants/HCT_maternal_epigenomic_targets/figures/.*

### 1. DAISY GRS2 calculated from TOPMed R2 imputation with 3 FDR groups

Not shown in above plots, only shown here.

*Min/max not set here since scale doesn't match primary output.*

```{r}
anova_fdr_d <- oneway.test(
  GRS2 ~ fdr_3level,
  data = grs2_daisy_pheno,
  var.equal = F)
anova_fdr_d_p <- round(anova_fdr_d$p.value, 3)

plot_fdr_d <- ggplot(grs2_daisy_pheno, aes(x = fdr_3level, y = GRS2)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY", x = "Family History") + 
  theme_bw() +
  # scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min - 1, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min - 1, label = length(x)), 
               geom = "text") +
  annotate("text", label = sprintf("italic(P)~'='~'%s'", anova_fdr_d_p),
            x = Inf, y = Inf, hjust = 1.2, vjust = 1.2, parse = T)
plot_fdr_d

# ggsave("../Grants/HCT_maternal_epigenomic_targets/figures/daisy_grs2_by_fdr.png",
#        plot_fdr_d,
#        width = 3, height = 2.4)

```

### 2. TEDDY GRS2 calculated from 1000G with 3 FDR groups.

Shown in plot above.

```{r}
anova_fdr_t_p <- round(anova_fdr_t$p.value, 3)

plot_fdr_t <- ggplot(grs2_teddy_pheno, aes(x = fdr_3level, y = GRS2)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "TEDDY", x = "Family History") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min - 1, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min - 1, label = length(x)), 
               geom = "text") +
  annotate("text", label = sprintf("italic(P)~'='~'%s'", anova_fdr_t_p),
            x = Inf, y = Inf, hjust = 1.2, vjust = 1.2, parse = T)
plot_fdr_t

# ggsave("../Grants/HCT_maternal_epigenomic_targets/figures/teddy_grs2_by_fdr.png",
#        plot_fdr_t,
#        width = 3, height = 2.4)

```


### 3. DAISY GRS2 calculated from 1000G with 3 FDR groups

Version with all SNPs, not shown above.

*Min/max not set here since scale doesn't match primary output.*

```{r}
anova_fdr_d_kg <- oneway.test(
  GRS2 ~ fdr_3level,
  data = grs2_daisy_kg_pheno,
  var.equal = F)
anova_fdr_d_kg_p <- round(anova_fdr_d_kg$p.value, 3)

plot_fdr_d_kg <- ggplot(grs2_daisy_kg_pheno, aes(x = fdr_3level, y = GRS2)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY", x = "Family History") + 
  theme_bw() +
  # scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min - 1, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min - 1, label = length(x)), 
               geom = "text") +
  annotate("text", label = sprintf("italic(P)~'='~'%s'", anova_fdr_d_kg_p),
            x = Inf, y = Inf, hjust = 1.2, vjust = 1.2, parse = T)
plot_fdr_d_kg

# ggsave("../Grants/HCT_maternal_epigenomic_targets/figures/daisy_grs2_1000g_by_fdr.png",
#        plot_fdr_d_kg,
#        width = 3, height = 2.4)

```

### 4. DAISY GRS2 calculated from 1000G same 61 SNPs as TEDDY with 3 FDR groups

Same plot as above.

```{r}
anova_fdr_d_kg_t_p <- round(anova_fdr_d_kg_t$p.value, 3)

plot_fdr_d_kg_t <- ggplot(grs2_daisy_kg_t_pheno, aes(x = fdr_3level, y = GRS2)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = "DAISY", x = "Family History") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min - 1, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min - 1, label = length(x)), 
               geom = "text") +
  annotate("text", label = sprintf("italic(P)~'='~'%s'", anova_fdr_d_kg_t_p),
            x = Inf, y = Inf, hjust = 1.2, vjust = 1.2, parse = T)
plot_fdr_d_kg_t

# ggsave("../Grants/HCT_maternal_epigenomic_targets/figures/daisy_grs2_1000g_teddy_snps_by_fdr.png",
#        plot_fdr_d_kg_t,
#        width = 3, height = 2.4)

```


### 5. TEDDY & DAISY GRS2 calculated from 1000G same 61 SNPs with 3 FDR groups

Combined TEDDY & DAISY plot with same SNPs, same as the plots above but
arranged in grid.

```{r}
plot_comb <- ggarrange(plot_fdr_d_kg_t,
                       plot_fdr_t  + labs(y = ""))
plot_comb

# ggsave("../Grants/HCT_maternal_epigenomic_targets/figures/daisy_teddy_grs2_1000g_same_snps_by_fdr.png",
#        plot_comb,
#        width = 6, height = 2.4)

```


# **Session Info**

```{r session_info}
sessionInfo()
```
