---
title: "5_genetic_risk_t1d_moms"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection/reports/risk_scores")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB")))

library(tidyverse)
library(table1)
library(sas7bdat)
library(kableExtra)
library(ggpubr)
library(forestplot)
library(khroma)
library(ggpattern)
library(readxl)

```

# **For IDS Abstract Submission**

Everything in this first section was updated prior to and used for IDS 2024
abstract submission.

## **Overview**

Analysis for IDS 2024 abstract. Main question is: Do kids born to moms with T1D
have lower genetic risk? Is the risk for certain groups?

Using DAISY & TEDDY 1000G-calculated GRS2 scores. Using the scores calculated
with the same set of SNPs for both studies. Subset of individuals is different
than what was used in 3_fam_hist_teddy_daisy report since that report used
individuals in the T1D studies and limited to DNAm. Here, using the IA case
control studies, limited to individuals with genetics and of majority EUR by
KING (used KING's auto-determined ancestry, seems to be 2/3 cutoff). Now also
using the 1:1 T1D case control studies.

1. GRS2 by Family History None-Sig Difference
2. GRS2 Slight Decline in Moms - Stratify by HLA
3. Linear Modeling (GRS2 ~ fdr_3level + outcome, adding HLAGRP, & fdr_3level*HLAGRP)

*Other notes:*

+ For now using "HLAGRP" variable for DAISY. Could do additional checks on
  other HLA-related variables or follow up with Fran in future to confirm.
+ For now, testing the overlapping HLA categories only for now: DR3/3, 3/4, 4/4,
  and 4/X. Not sure exactly how model will do with 0 Moms for DAISY 3/3, but
  will see.

## **Load Data**

Phenotype files for IA & T1D case control studies, filtered to those with genetic
data and with primary EUR ancestry made in 4.2_make_ia_phenos.R. Filtering on
ancestry since GRS2 is known to not work as well in non-NHW. No filtering has
been done of the HLA categories in the input files, but prior to linear
modeling, HLA categories were filtered to just DR3/3, DR3/4, DR4/4, & DR4/X.

```{r load_data, message = F}
# TEDDY EUR phenotype
pheno_t_ia <- readRDS(
  "Maternal_Protection/data/teddy/risk_scores/pheno_ia_with_grs2_eur.rds")
pheno_t_ia_dr <- readRDS(
  "Maternal_Protection/data/teddy/risk_scores/pheno_ia_with_grs2_sd_eur_dr_filt.rds")
pheno_t_t1d <- readRDS(
  "Maternal_Protection/data/teddy/risk_scores/pheno_t1d_with_grs2_eur.rds")
pheno_t_t1d_dr <- readRDS(
  "Maternal_Protection/data/teddy/risk_scores/pheno_t1d_with_grs2_sd_eur_dr_filt.rds")
# Loading IA with FDR 4-level for poster
pheno_t_ia_dr_4l <- readRDS(
  "Maternal_Protection/data/teddy/risk_scores/pheno_ia_with_grs2_sd_eur_dr_filt_fdr_4level.rds")
pheno_t_ia_dr_4l_hla <- readRDS(
  "Maternal_Protection/data/teddy/risk_scores/pheno_ia_with_grs2_sd_hla_only_eur_dr_filt_fdr_4level.rds")
pheno_t_ia_dr_4l_non_hla <- readRDS(
  "Maternal_Protection/data/teddy/risk_scores/pheno_ia_with_grs2_sd_non_hla_eur_dr_filt_fdr_4level.rds")

# DAISY EUR phenotype
pheno_d_ia <- readRDS(
  "Maternal_Protection/data/daisy/risk_scores/pheno_ia_with_grs2_eur.rds")
pheno_d_ia_dr <- readRDS(
  "Maternal_Protection/data/daisy/risk_scores/pheno_ia_with_grs2_sd_eur_dr_filt.rds")
pheno_d_t1d <- readRDS(
  "Maternal_Protection/data/daisy/risk_scores/pheno_t1d_with_grs2_eur.rds")
pheno_d_t1d_dr <- readRDS(
  "Maternal_Protection/data/daisy/risk_scores/pheno_t1d_with_grs2_sd_eur_dr_filt.rds")
# Loading IA with FDR 4-level for poster
pheno_d_ia_dr_4l <- readRDS(
  "Maternal_Protection/data/daisy/risk_scores/pheno_ia_with_grs2_sd_eur_dr_filt_fdr_4level.rds")
pheno_d_ia_dr_4l_hla <- readRDS(
  "Maternal_Protection/data/daisy/risk_scores/pheno_ia_with_grs2_sd_hla_only_eur_dr_filt_fdr_4level.rds")
pheno_d_ia_dr_4l_non_hla <- readRDS(
  "Maternal_Protection/data/daisy/risk_scores/pheno_ia_with_grs2_sd_non_hla_eur_dr_filt_fdr_4level.rds")

# TEDDY HLA DR-DQ generated by PRS toolkit
hla_gt_t <- read_delim(
  "Maternal_Protection/data/teddy/risk_scores/grs2/scoring_1000g/imp_1000g_haplogeno_hg19_Categorical.txt")

# DAISY HLA DR-DQ generated by PRS toolkit
hla_gt_d <- read_delim(
  "DAISY/genetics/risk_scores/data/grs2/1000g/imp_1000g_haplogeno_hg19_Categorical.txt")
map_d <- read_delim(
  "DAISY/genetics/daisy_ask_genetics/exomechip_id_key_plink_rename",
  col_names = c("lab_ID", "lab_ID2", "study_ID", "study_ID2"))

# Additional data for quick checks in final "Next Steps" section of GDM in TEDDY
pheno_t <- read_csv(
  "Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_demographic_masked.csv")
pheno_d <- read.sas7bdat("DAISY/pheno/ivyomicssample.sas7bdat")
ia_def_t <- read_csv(
  "Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_casecntrl_ialist3.csv")

# Load model results to make forest plot
mod <- read_csv(
"Maternal_Protection/data/results/risk_scores/lm_grs2_sd_fdr_hla_results_4level_dad_ref.csv")

# GRS2 SNP info
grs2 <- read_tsv(
  "Maternal_Protection/data/teddy/risk_scores/grs2/scoring_1000g/grs2_score_input_with_info.txt",
  col_types = c("cccnccnnc"))
grs2_full <- read_xlsx(
  "/Users/slacksa/repos/explore_matprot/risk_scores/hla-prs-toolkit/Tools/Snps/1000G/T1DGRS67_1000G_hg19.xlsx")

```

## **Study Table 1s**

### **IA Case/Control Studies**

Shown with primarily EUR (>65% by KING) individuals with genetic data in the
case/control study, and below with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **TEDDY - No DR Filter**

```{r}
table1(~ sex + fdr_3level + GRS2 + HLAGRP | IA,
       data = pheno_t_ia, overall = "Total")
```

#### **TEDDY - High-Risk DR**

```{r}
table1(~ sex + fdr_3level + GRS2 + GRS2_sd + HLAGRP | IA,
       data = pheno_t_ia_dr, overall = "Total")
```

Get GRS2 range by FDR category:

```{r}
table1(~ GRS2 | fdr_3level,
       data = pheno_t_ia_dr, overall = "Total")
```

Check TEDDY FDR categories:

Want to make sure DM mom is accurately counted as T1D mom for IDS abstract
analysis:

```{r}
pheno_t_ia_dr_pheno <- inner_join(
  pheno_t_ia_dr, pheno_t, by = c("ID" = "Norris_epigenetics_maskid"))

# print("FDR 3-level by dad:")
# table(pheno_t_ia_dr_pheno$fdr_3level, pheno_t_ia_dr_pheno$dad)
# print("FDR 3-level by mom:")
# table(pheno_t_ia_dr_pheno$fdr_3level, pheno_t_ia_dr_pheno$mom)
# print("FDR 3-level by sibling:")
# table(pheno_t_ia_dr_pheno$fdr_3level, pheno_t_ia_dr_pheno$sibling)

# table(pheno_t$family_mem)

pheno_t_ia_dr_pheno[
  which(pheno_t_ia_dr_pheno$family_mem==3 &
          pheno_t_ia_dr_pheno$diabetes==1),
  c("ID", "fdr_3level", "diabetes", "family_mem", "mom", "sibling")]
```

Yes, the mom diabetes = 1 (gestational diabetes) and family_mem = 3 (mom &
sibling), is correctly coded as mom in the fdr_3level variable.

#### **DAISY - No DR Filter**

```{r}
table1(~ sex + fdr_3level + GRS2 + HLAGRP | IA,
       data = pheno_d_ia, overall = "Total")
```

#### **DAISY - High-Risk DR**

```{r}
table1(~ sex + fdr_3level + GRS2 + GRS2_sd + HLAGRP | IA,
       data = pheno_d_ia_dr, overall = "Total")
```


Get GRS2 range by FDR category:

```{r}
table1(~ GRS2 | fdr_3level,
       data = pheno_d_ia_dr, overall = "Total")
```

Check DAISY FDR categories:

Want to make sure two IDs that had to previously be re-coded to Dad, based off
of previous script at: code/Step8.1_logistic_fdr_interaction.R.

The IDs "00275-0", "00459-0" should be "Dad/Sib":

```{r}
pheno_d_ia_dr$fdr_3level[which(pheno_d_ia_dr$ID %in% c("00275-0","00459-0"))]
```

They are! I didn't have to manually recode them in 4.2_make_ia_phenos.R - they
were already correct. Think that's because I pulled input data from the IVYOmics
sas7bdat instead of an older freeze of the data. Keeping this for IDS abstract
analysis.

*TODO: if in future, want to instead make input data match the input data*
*for report "FDR_interaction_summary" in DAISY, may need to check differences*
*in phenotype files.*

### **T1D Case/Control Studies**

Shown with primarily EUR (>65% by KING) individuals with genetic data in the
case/control study, and below with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **TEDDY - No DR Filter**

```{r}
table1(~ sex + fdr_3level + GRS2 + HLAGRP | T1D,
       data = pheno_t_t1d, overall = "Total")
```

#### **TEDDY - High-Risk DR**

```{r}
table1(~ sex + fdr_3level + GRS2 + HLAGRP | T1D,
       data = pheno_t_t1d_dr, overall = "Total")
```

Note the above two tables are the same, there are no additional exclusions for
TEDDY T1D.

#### **DAISY - No DR Filter**

```{r}
table1(~ sex + fdr_3level + GRS2 + HLAGRP | T1D,
       data = pheno_d_t1d, overall = "Total")
```

#### **DAISY - High-Risk DR**

```{r}
table1(~ sex + fdr_3level + GRS2 + HLAGRP | T1D,
       data = pheno_d_t1d_dr, overall = "Total")
```

## **HLA Categories**

### **IA Case/Control Studies**

Shown on top with primarily EUR individuals with genetic data in the case/
control study, and on bottom with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **TEDDY**

Haplogenotype definitions from phenotype file. Summarize to match DAISY:
*To note: ineligible HLA already removed when phenotype file was made.*

+ -1='HLA Results Pending'
+ 0='Not Eligible'
+ 1='DR4 030X/0302 DR3 0501/0201' = DR3/4
+ 2='DR4 030X/0302 DR4 030X/0302' = DR4/4
+ 3='DR4 030X/0302 DR4 030X/020X' = DR4/4b = DR4/?, excl if not FDR
  + **TO NOTE: not sure how to summarize this, but not handling now since**
  + **not present in our data.**
+ 4='DR4 030X/0302 DR8 0401/0402' = DR4/8 = DR4/X
+ 5='DR4 030X/0302 DR1 0101/0501' = DR4/DR1 = DR4/X, excl if not FDR
+ 6='DR4 030X/0302 DR13 0102/0604' = DR4/DR13 = DR4/X, excl if not FDR
+ 7='DR4 030X/0302 DR4 030X/0304' = DR4/4
+ 8='DR4 030X/0302 DR9 030X/0303' = DR4/DR9 = DR4/X, excl if not FDR
+ 9='DR3 0501/0201 DR3 0501/0201' = DR3/3
+ 10='DR3 0501/0201 DR9 030X/0303' = DR3/DR9 = DR3/X, excl if not FDR
+ 99='Results Under Review'

Summary of original categories:

```{r out.width = '50%', fig.show = 'hold'}
table(pheno_t_ia$HLA_Category)
table(pheno_t_ia_dr$HLA_Category)

```

Check categories with FDR status for removal of FDR "None" for HLA category
groups 3, 5, 6, 8, & 10:

```{r}
table(pheno_t_ia$fdr_3level, pheno_t_ia$HLA_Category)
table(pheno_t_ia_dr$fdr_3level, pheno_t_ia_dr$HLA_Category)
```

In FDR "None", there are 0 to be removed from 3, 5, 6, 8, or 10. There are no
individuals in group 3 at all. So no additional exclusions are needed here.

Summary after matching to DAISY categories:

```{r}
table(pheno_t_ia$HLAGRP)
table(pheno_t_ia_dr$HLAGRP)

```

#### **DAISY**

```{r}
table(pheno_d_ia$HLAGRP)
table(pheno_d_ia_dr$HLAGRP)
```



### **T1D Case/Control Studies**

Shown on top with primarily EUR individuals with genetic data in the case/
control study, and on bottom with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **TEDDY**

Summary of original categories:

```{r out.width = '50%', fig.show = 'hold'}
table(pheno_t_t1d$HLA_Category)
table(pheno_t_t1d_dr$HLA_Category)

```

*Note that there are no additional DR exlcusions made for TEDDY T1D, so below*
*only showing one version of each table.*

Check categories with FDR status for removal of FDR "None" for HLA category
groups 3, 5, 6, 8, & 10:

```{r}
table(pheno_t_t1d$fdr_3level, pheno_t_t1d$HLA_Category)
```

In FDR "None", there are 0 to be removed from 3, 5, 6, 8, or 10. There are no
individuals in group 3 at all. So no additional exclusions are needed here.

Summary after matching to DAISY categories:

```{r}
table(pheno_t_t1d$HLAGRP)

```

#### **DAISY**

```{r}
table(pheno_d_t1d$HLAGRP)
table(pheno_d_t1d_dr$HLAGRP)
```

## **HLA Categories with FDR**

### **IA Case/Control Studies**

Shown on top with primarily EUR individuals with genetic data in the case/
control study, and on bottom with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **TEDDY**

```{r}
table(pheno_t_ia$fdr_3level, pheno_t_ia$HLAGRP)
table(pheno_t_ia_dr$fdr_3level, pheno_t_ia_dr$HLAGRP)
```

#### **DAISY**

```{r}
table(pheno_d_ia$fdr_3level, pheno_d_ia$HLAGRP)
table(pheno_d_ia_dr$fdr_3level, pheno_d_ia_dr$HLAGRP)
```


### **T1D Case/Control Studies**

Shown on top with primarily EUR individuals with genetic data in the case/
control study, and on bottom with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **TEDDY**

```{r}
table(pheno_t_t1d$fdr_3level, pheno_t_t1d$HLAGRP)
table(pheno_t_t1d_dr$fdr_3level, pheno_t_t1d_dr$HLAGRP)
```

Again, no difference for TEDDY T1D.

#### **DAISY**

```{r}
table(pheno_d_t1d$fdr_3level, pheno_d_t1d$HLAGRP)
table(pheno_d_t1d_dr$fdr_3level, pheno_d_t1d_dr$HLAGRP)
```



## **HLA Categories from Genetics**

*TO NOTE: for now, this section only checks overlap with the IA case/control*
*studies without HLA group exclusion. Will update in future if needed.*

### **IA Case/Control Studies**

Step 1 of the HLA PRS toolkit used to calculated the 1000G GRS2 score, already
completes hard calling of HLA haplogenotypes from SNP proxies. It uses the HLA
LD-Tiebreak algorithm to resolve > 2 HLA calls per person. Read more on the
toolkit [GitHub](https://github.com/sethsh7/hla-prs-toolkit). 

**COMPONENT** = **DR** = **DQ**
HLA-DQ81    =   4   =   8
HLA-DQ25    =   3   =   2
HLA-DQ22
HLA-DQ75
HLA-DQ51    =   1	
HLA-DQ73
HLA-DQ63
HLA-DQ92
HLA-DQ42    =   8	
HLA-DQ53
HLA-DQ93    =   9	
HLA-DQ61
HLA-DQ69
HLA-DQ62

*TODO: confirm we want to use this output, vs. manually mapping of tag SNPs*
*which could result in > 2 per person.*

#### **TEDDY**

Show DQ assignments as given in initial file:

```{r}
hla_gt_t <- hla_gt_t %>%
  dplyr::mutate(FID = gsub("_.+$", "", IID),
                IID = gsub("^.+_", "", IID))
pheno_t_ia_gt <- inner_join(
  pheno_t_ia %>% dplyr::mutate(ID = as.character(ID)),
  hla_gt_t,
  by = c("ID" = "IID"))

table(pheno_t_ia_gt$fdr_3level, pheno_t_ia_gt$GENO1)
table(pheno_t_ia_gt$fdr_3level, pheno_t_ia_gt$GENO2)
```

Now, update them to match same DR categories as TEDDY & DAISY:

*Only converting those that map to DR3 & DR4.*

```{r}
pheno_t_ia_gt <- pheno_t_ia_gt %>%
  dplyr::mutate(HLAGRP_gt = case_when(
    GENO1 == "DQ25" & GENO2 == "DQ25" ~ "DR3/3",
    GENO1 == "DQ25" & GENO2 == "DQ81" ~ "DR3/4",
    GENO1 == "DQ25" & GENO2 == "X" ~ "DR3/X",
    GENO1 == "DQ81" & GENO2 == "DQ25" ~ "DR3/4",
    GENO1 == "DQ81" & GENO2 == "DQ81" ~ "DR4/4",
    GENO1 == "DQ81" & GENO2 == "X" ~ "DR4/X",
  ))
```

Check concordance with HLAGRP variable given by phenotype file:

```{r}
table(pheno_t_ia_gt$HLAGRP, pheno_t_ia_gt$HLAGRP_gt)
```

Actually I think this is pretty good!

#### **DAISY**

Show DQ assignments as given in initial file:

```{r}
hla_gt_d <- hla_gt_d %>%
  dplyr::mutate(FID = gsub("_.+$", "", IID),
                IID = gsub("^.+_", "", IID))
hla_gt_d <- inner_join(
  hla_gt_d,
  map_d,
  by = c("IID" = "lab_ID"))

pheno_d_ia_gt <- inner_join(
  pheno_d_ia,
  hla_gt_d,
  by = c("ID" = "study_ID"))

table(pheno_d_ia_gt$fdr_3level, pheno_d_ia_gt$GENO1)
table(pheno_d_ia_gt$fdr_3level, pheno_d_ia_gt$GENO2)
```

Now, update them to match same DR categories as TEDDY & DAISY:

*Only converting those that map to DR3 & DR4.*
*TODO: update when have more time to check concordance of all categories,*
*curious about the Moms in DRX/X.*

```{r}
pheno_d_ia_gt <- pheno_d_ia_gt %>%
  dplyr::mutate(HLAGRP_gt = case_when(
    GENO1 == "DQ25" & GENO2 == "DQ25" ~ "DR3/3",
    GENO1 == "DQ25" & GENO2 == "DQ81" ~ "DR3/4",
    GENO1 == "DQ25" & GENO2 == "X" ~ "DR3/X",
    GENO1 == "DQ81" & GENO2 == "DQ25" ~ "DR3/4",
    GENO1 == "DQ81" & GENO2 == "DQ81" ~ "DR4/4",
    GENO1 == "DQ81" & GENO2 == "X" ~ "DR4/X",
  ))
```


Check concordance with HLAGRP variable given by phenotype file:

```{r}
table(pheno_d_ia_gt$HLAGRP, pheno_d_ia_gt$HLAGRP_gt)
```

Even better concordance than TEDDY!


## **GRS2 Plots**

```{r}
# Plot prep
min <- floor(min(min(pheno_t_ia$GRS2), min(pheno_d_ia$GRS2),
                 min(pheno_t_t1d$GRS2), min(pheno_d_t1d$GRS2)))
max <- ceiling(max(max(pheno_t_ia$GRS2), max(pheno_d_ia$GRS2),
                   max(pheno_t_t1d$GRS2), max(pheno_d_t1d$GRS2)))

```

### **Check Normal Distribution**

*TO NOTE: only checking distribution in non-DR filtered data**

#### **IA - TEDDY**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(pheno_t_ia, aes(x = GRS2)) +
  geom_histogram()
ggplot(pheno_t_ia, aes(x = GRS2)) +
  geom_density()
```

#### **IA - DAISY**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(pheno_d_ia, aes(x = GRS2)) +
  geom_histogram()
ggplot(pheno_d_ia, aes(x = GRS2)) +
  geom_density()
```

#### **T1D - TEDDY**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(pheno_t_t1d, aes(x = GRS2)) +
  geom_histogram()
ggplot(pheno_t_t1d, aes(x = GRS2)) +
  geom_density()
```

#### **T1D - DAISY**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(pheno_d_t1d, aes(x = GRS2)) +
  geom_histogram()
ggplot(pheno_d_t1d, aes(x = GRS2)) +
  geom_density()
```


### **By Family History**

```{r}
plot_fdr <- function(df, study) {
  ggplot(df, aes(x = fdr_3level, y = GRS2)) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = study,
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))
}

anova_fdr <- function(df) {
  oneway.test(
    GRS2 ~ fdr_3level,
    data = df,
    var.equal = F)
}
```


Shown on the left with primarily EUR individuals with genetic data in the case/
control study, and on the right with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **IA - TEDDY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr(pheno_t_ia, "TEDDY")
plot_fdr(pheno_t_ia_dr, "TEDDY")
```

One-way ANOVA of GRS2 ~ 3-level FDR. Did not assume equal variances.

```{r}
anova_fdr(pheno_t_ia)
anova_fdr(pheno_t_ia_dr)
```

Difference here is barely significant. Surprised by this with how the plot
looks. Maybe it's the "None" group driving?

One add DR group exclusion, the difference becomes a lot more significant. Why?

#### **IA - DAISY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr(pheno_d_ia, "DAISY")
plot_fdr(pheno_d_ia_dr, "DAISY")
```

One-way ANOVA of GRS2 ~ 3-level FDR. Did not assume equal variances.

```{r}
anova_fdr(pheno_d_ia)
anova_fdr(pheno_d_ia_dr)
```

Again, result becomes more significant once add DR exclusion.


#### **T1D - TEDDY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr(pheno_t_t1d, "TEDDY")
plot_fdr(pheno_t_t1d_dr, "TEDDY")
```

*Note that above are identical, so only including one ANOVA output.*

One-way ANOVA of GRS2 ~ 3-level FDR. Did not assume equal variances.

```{r}
anova_fdr(pheno_t_t1d)
```

Difference here is not significant - so a fairly different result from IA.

#### **T1D - DAISY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr(pheno_d_t1d, "DAISY")
plot_fdr(pheno_d_t1d_dr, "DAISY")
```

One-way ANOVA of GRS2 ~ 3-level FDR. Did not assume equal variances.

```{r}
anova_fdr(pheno_d_t1d)
anova_fdr(pheno_d_t1d_dr)
```

Again, result becomes more significant once add DR exclusion. Also interesting
shift up in GRS2 scores for Moms and Dad/Sibs.


### **By IA Case/Control and Family History**

```{r}
plot_fdr_outcome <- function(df, study, outcome) {
  ggplot(df, aes(x = fdr_3level, y = GRS2, fill = !!sym(outcome))) +
  geom_boxplot(alpha = 0.5, varwidth = TRUE) +
  labs(title = study,
       fill = outcome,
       x = "FDR") + 
  theme_bw() +
  scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min, label = length(x)), 
               geom = "text", position = position_dodge(width = 0.4))
  
}
```

Shown on the left with primarily EUR individuals with genetic data in the case/
control study, and on the right with the additional exclusion to DR3/3, DR3/4,
DR4/4, & DR4/X.

#### **IA - TEDDY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr_outcome(pheno_t_ia, "TEDDY", "IA")
plot_fdr_outcome(pheno_t_ia_dr, "TEDDY", "IA")
```


#### **IA - DAISY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr_outcome(pheno_d_ia, "DAISY", "IA")
plot_fdr_outcome(pheno_d_ia_dr, "DAISY", "IA")
```

#### **T1D - TEDDY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr_outcome(pheno_t_t1d, "TEDDY", "T1D")
plot_fdr_outcome(pheno_t_t1d_dr, "TEDDY", "T1D")
```


#### **T1D - DAISY**

```{r out.width = '50%', fig.show = 'hold'}
plot_fdr_outcome(pheno_d_t1d, "DAISY", "T1D")
plot_fdr_outcome(pheno_d_t1d_dr, "DAISY", "T1D")
```

## **Models**

*TO NOTE: this moved to 4.3_lm_grs2_fdr_hla.R, with results in*
*data/risk_scores/lm_grs2_fdr_hla_results.csv &*
*data/risk_scores/ids_results_summary.xlsx.*

Using linear models for power.

For TEDDY, clinical center as covariate accounts for matching as well as
population structure. For DAISY, there is no site. Using the HLA cateogry
variable from the phenotype files.

Want to test the overlapping HLAGRP categories only for now: DR3/3, 3/4, 4/4,
and 4/X. Not sure exactly how model will do with 0 Moms for DAISY 3/3, but
will see.

*TODO: still may want to run with and without Finland?*

Clinical center variable coded as 1-6, with the following values: 1 = Colorado,
2 = Georgia/Florida, 3 = Washington, 4 = Finland, 5 = Germany, 6 = Sweden.
Want to set Colorado as the reference level since it is generally a genetically
mixed site, not very different like Finland or Sweden, didn't have different
recruitment like Germany, and is larger than the other US sites.


# **After IDS Abstract Submission**

## **Next Steps**

These things not included in abstract, but were Randi's ideas for next steps if
we want to publish this on its own in a short report:

1. We should look at the feasibility of splitting the father/sibling group.
  We’d need to look at frequencies of those with single relative (mom, dad, sib,
  none), and multiple relatives with T1D.
  + **Update 20241021 - am splitting this for IDS poster, see For IDS Poster**
    **section.**
2. We also might want to consider a comparison with T1D-mom vs. GDM in TEDDY.
  Or look at frequencies of the T1D fam hist categories by GDM and decide
  whether GDM could be its own level in the T1D famhist variable.
  + **Update 20241021 - still sorting out re-definition of famhist/GDMs, so**
    **not looking into this for IDS poster.**
  
### **Check Feasibility of Splitting Dad/Sib Group**

**In IA case-control filtered to high risk HLA-DR categories, with only EUR >**
**65%, with genetic data.**

Want to check frequencies of single relatives and multiple relatives with T1D.

#### **TEDDY IA High-Risk DR**

First, give each individual only one relative, where Mom trumps Dad and Dad
trumps Sib and Sib trumps None:

Check this with family_mem, where: 0=none, 1=all, 2=mom & dad, 3=mom & sibling,
4=dad & sibling, 5=mom, 6=dad, 7=sibling

```{r}
pheno_t_ia_dr_pheno$fdr_sing =
  ifelse(pheno_t_ia_dr_pheno$mom == 1, "Mom",
         ifelse(pheno_t_ia_dr_pheno$dad == 1, "Dad",
                ifelse(pheno_t_ia_dr_pheno$sibling == 1, "Sib", "None")))

table(pheno_t_ia_dr_pheno$fdr_sing)
table(pheno_t_ia_dr_pheno$fdr_sing, pheno_t_ia_dr_pheno$family_mem)
```

  
Next, accurately account for individuals with more than one relative with T1D:

```{r}
pheno_t_ia_dr_pheno$fdr_mult =
  case_when(
    pheno_t_ia_dr_pheno$mom == 1 & pheno_t_ia_dr_pheno$dad == 1 & pheno_t_ia_dr_pheno$sibling == 1 ~ "Mom/Dad/Sib",
    pheno_t_ia_dr_pheno$mom == 1 & pheno_t_ia_dr_pheno$dad == 1 ~ "Mom/Dad",
    pheno_t_ia_dr_pheno$mom == 1 & pheno_t_ia_dr_pheno$sibling == 1 ~ "Mom/Sib",
    pheno_t_ia_dr_pheno$dad == 1 & pheno_t_ia_dr_pheno$sibling == 1 ~ "Dad/Sib",
    pheno_t_ia_dr_pheno$mom == 1 ~ "Mom",
    pheno_t_ia_dr_pheno$dad == 1 ~ "Dad",
    pheno_t_ia_dr_pheno$sibling == 1 ~ "Sib")

table(pheno_t_ia_dr_pheno$fdr_mult)
```

Very few individuals have multiple family members with T1D. After discussing
with Randi, she thinks we could exclude those with multiple family members from
the next step of analysis.

#### **DAISY IA High-Risk DR**

First, give each individual only one relative, where Mom trumps Dad and Dad
trumps Sib and Dib trumps None:

```{r}
pheno_d_dedup <- pheno_d %>%
  dplyr::select(ID, T1Dgroup, IAgroup2, NHW, dr34, SEX, DMfamily) %>%
  dplyr::filter(!duplicated(.))

pheno_d_ia_dr_pheno <- left_join(
  pheno_d_ia_dr,
  pheno_d_dedup,
  by = c("ID"))

pheno_d_ia_dr_pheno$fdr_sing =
  ifelse(pheno_d_ia_dr_pheno$DMfamily == "", "None",
         ifelse(pheno_d_ia_dr_pheno$DMfamily == "mom/dad" |
                  pheno_d_ia_dr_pheno$DMfamily == "mom", "Mom",
                ifelse(pheno_d_ia_dr_pheno$DMfamily == "dad", "Dad", "sibling")))

table(pheno_d_ia_dr_pheno$fdr_sing)
table(pheno_d_ia_dr_pheno$fdr_sing, pheno_d_ia_dr_pheno$DMfamily)
```

Next, accurately account for individuals with more than one relative with T1D:

```{r}
table(pheno_d_ia_dr_pheno$DMfamily)
```

Only 1 individual with multiple family members with T1D.

**Decision: exclude individuals with multiple family members, re-run model 1 in**
**script 4.3 with dad group as reference.**

### **Check for GDMs in TEDDY**

**In IA case-control filtered to high risk HLA-DR categories, with only EUR >**
**65%, with genetic data.**

Use variable "diabetes": What type of diabetes did mother have?  4 groups:

+ 1=Had gestational diabetes
+ 2=Has Type 1 diabetes
+ 3=Has Type 2 diabetes,
+ 4=No diabetes of any kind

*Note: If more than one type of diabetes indicated on mother questionnaire –*
*1=Gestational diabetes came before T1D/T2D, 2=Gestational diabetes came after*
*T1D, 3=Gestational diabetes came after T2D, . (missing) = Can’t determine*
*which came first/data show age mismatch.*


```{r}
table(pheno_t_ia_dr_pheno$diabetes)
```

Check T1D family history by diabetes variable:

Using both fdr_3level made by me, and family_mem variable: 0=none, 1=all,
2=mom & dad, 3=mom & sibling, 4=dad & sibling, 5=mom, 6=dad, 7=sibling.

```{r}
table(pheno_t_ia_dr_pheno$diabetes, pheno_t_ia_dr_pheno$fdr_3level)
table(pheno_t_ia_dr_pheno$diabetes, pheno_t_ia_dr_pheno$family_mem)
```

Use Charlie's code to check for individuals where FDR = mom and mom classified
as GDM:

```{r}
pheno_t_ia_dr_pheno[which(
  pheno_t_ia_dr_pheno$mom==1 &
    pheno_t_ia_dr_pheno$diabetes==1),
  c("ID", "fdr_3level", "diabetes", "family_mem", "mom", "sibling")]
```

Now, recode this individual so that diabetes = 2 (recode from GDM to T1D):

```{r}
pheno_t_ia_dr_pheno[which(
  pheno_t_ia_dr_pheno$mom==1 &
    pheno_t_ia_dr_pheno$diabetes==1), "diabetes"] = 2

pheno_t_ia_dr_pheno[which(
  pheno_t_ia_dr_pheno$mom==1 &
    pheno_t_ia_dr_pheno$diabetes==1),
  c("ID", "fdr_3level", "diabetes", "family_mem", "mom", "sibling")]
# To note: this should print 0 rows if recoding was correct
```

*TODO: ask Randi - do we expect more than 1 mom to need re-coding?*


# **For IDS Poster**

## Table 1s

For abstract, only present analysis in IA case/control restricted to high risk
DRs for TEDDY and DAISY and with 4-level FDR variable (this version NOT shown
above in table 1 section). Write out table 1s here to be used in poster.

TEDDY "cc" = clinical center: 1=COL, 2=GEO, 3=WAS, 4=FIN, 5=GER, 6=SWE, 133=NBD, 134=CHP

For both, need to add age at IA here.
*TODO: confirm pulling this correctly!*

Note that converting TEDDY age in days to years using 365.25. If revisit, may
want to use package combined with year of birth to do this more accurately.

```{r}
# Add age at IA
pheno_d_dedup <- pheno_d %>%
  dplyr::select(ID, Sv_age) %>%
  dplyr::filter(!duplicated(.))
# n_distinct(pheno_d_dedup$ID)   # 426/426

pheno_d_ia_dr_4l <- inner_join(
  pheno_d_ia_dr_4l,
  pheno_d_dedup,
  by = c("ID"))
# Dedup IA case def
# n_distinct(ia_def_t$Norris_epigenetics_maskid)  # 1621 / 1671
ia_def_t_dup_list <- ia_def_t %>%
  dplyr::filter(duplicated(Norris_epigenetics_maskid))
ia_def_t_dedup <- ia_def_t %>%
  dplyr::filter(
    !(Norris_epigenetics_maskid %in% ia_def_t_dup_list$Norris_epigenetics_maskid &
        outcome == 0))
pheno_t_ia_dr_4l <- inner_join(
  pheno_t_ia_dr_4l,
  ia_def_t_dedup %>% dplyr::select(Norris_epigenetics_maskid, case_endptage),
  by = c("ID" = "Norris_epigenetics_maskid"))
# Convert age in days to years
pheno_t_ia_dr_4l$case_endptage_yr <- pheno_t_ia_dr_4l$case_endptage / 365.25


pheno_t_ia_dr_4l <- pheno_t_ia_dr_4l %>%
  dplyr::mutate(cc_name = case_when(
    cc == 1 ~ "COL",
    cc == 2 ~ "GEO",
    cc == 3 ~ "WAS",
    cc == 4 ~ "FIN",
    cc == 5 ~ "GER",
    cc == 6 ~ "SWE"))

table1_t <- table1(~ sex + case_endptage_yr + fdr_4level + GRS2 + GRS2_sd + HLAGRP + cc_name | IA,
       data = pheno_t_ia_dr_4l, overall = "Total")
table1_t

table1_d <- table1(~ sex + Sv_age + fdr_4level + GRS2 + GRS2_sd + HLAGRP | IA,
       data = pheno_d_ia_dr_4l, overall = "Total")
table1_d
```

## Boxplots

Want 4 FDR groups, faceted by TEDDY and DAISY and also by full GRS2 and non-HLA
GRS2. Guiding question is around what's driving?

Colors = CU gold #CFB97D, Paul Tol's colorblind-friendly blue #4477AA

### Combined - with IA Case/Control

```{r}
# all
pheno_d_ia_dr_4l$study <- "DAISY"
pheno_t_ia_dr_4l$study <- "TEDDY"
merge_4l <- rbind(
  pheno_d_ia_dr_4l %>% dplyr::select(-Sv_age),
  pheno_t_ia_dr_4l %>% dplyr::select(-cc, -cc_name, -HLA_Category, -case_endptage, -case_endptage_yr)
)

# HLA
pheno_d_ia_dr_4l_hla$study <- "DAISY"
pheno_t_ia_dr_4l_hla$study <- "TEDDY"
merge_4l_hla <- rbind(
  pheno_d_ia_dr_4l_hla,
  pheno_t_ia_dr_4l_hla %>% dplyr::select(-cc, -HLA_Category)
)

# Non-HLA
pheno_d_ia_dr_4l_non_hla$study <- "DAISY"
pheno_t_ia_dr_4l_non_hla$study <- "TEDDY"
merge_4l_non_hla <- rbind(
  pheno_d_ia_dr_4l_non_hla,
  pheno_t_ia_dr_4l_non_hla %>% dplyr::select(-cc, -HLA_Category)
)

merge_4l$type <- "All"
merge_4l_hla$type <- "HLA"
merge_4l_non_hla$type <- "Non-HLA"

merge_4l_all <- rbind(merge_4l, merge_4l_hla, merge_4l_non_hla)

boxplot_all <- ggplot(merge_4l_all, aes(x = fdr_4level, y = GRS2_sd, fill = IA)) +
  geom_boxplot(alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "FDR") + 
  theme_bw() +
  # scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min(merge_4l_all$GRS2_sd - 0.2), label = length(x)), 
               geom = "text", position = position_dodge(width = 0.8)) +
  facet_wrap(vars(study, type)) +
  scale_fill_manual(values = c("#CFB97D", "#4477AA")) +
  scale_x_discrete(labels = c(
    "None" = "None", "Dad" = "Dad", "Mom" = "Mom", "Sib" = "Sibling"),
    expand = c(0, 0))

boxplot_all
```

### Combined - No IA Case/Control

```{r}

boxplot_all_no_ia <- ggplot(merge_4l_all, aes(x = fdr_4level, y = GRS2_sd)) +
  geom_boxplot(alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "FDR") + 
  theme_bw() +
  # scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min(merge_4l_all$GRS2_sd - 0.2), label = length(x)), 
               geom = "text", position = position_dodge(width = 0.8)) +
  facet_wrap(vars(study, type)) +
  scale_fill_manual(values = c("#CFB97D", "#4477AA")) +
  scale_x_discrete(labels = c(
    "None" = "None", "Dad" = "Dad", "Mom" = "Mom", "Sib" = "Sibling"),
    expand = c(0, 0))

boxplot_all_no_ia
```

### Combined - DR3/4 and DR4/4 Only

*Note that for now showing actual GRS2 because SD would need to be recalculated.*

```{r}

merge_4l_all_hr <- merge_4l_all %>%
  dplyr::filter(HLAGRP == "DR3/4" | HLAGRP == "DR4/4")

boxplot_all_hr <- ggplot(merge_4l_all_hr, aes(x = fdr_4level, y = GRS2, fill = IA)) +
  geom_boxplot(alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "FDR") + 
  theme_bw() +
  # scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min(merge_4l_all$GRS2 - 0.2), label = length(x)), 
               geom = "text", position = position_dodge(width = 0.8)) +
  facet_wrap(vars(study, type), scales = "free_y") +
  scale_fill_manual(values = c("#CFB97D", "#4477AA")) +
  scale_x_discrete(labels = c(
    "None" = "None", "Dad" = "Dad", "Mom" = "Mom", "Sib" = "Sibling"),
    expand = c(0, 0))

# boxplot_all_hr
```


### Combined - DR3/4 and DR4/4 Only, No IA Case/Control

*Note that for now showing actual GRS2 because SD would need to be recalculated.*

```{r}

merge_4l_all_hr <- merge_4l_all %>%
  dplyr::filter(HLAGRP == "DR3/4" | HLAGRP == "DR4/4")

boxplot_all_hr_no_ia <- ggplot(merge_4l_all_hr, aes(x = fdr_4level, y = GRS2)) +
  geom_boxplot(alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "FDR") + 
  theme_bw() +
  # scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min(merge_4l_all$GRS2 - 0.2), label = length(x)), 
               geom = "text", position = position_dodge(width = 0.8)) +
  facet_wrap(vars(study, type), scales = "free_y") +
  scale_fill_manual(values = c("#CFB97D", "#4477AA")) +
  scale_x_discrete(labels = c(
    "None" = "None", "Dad" = "Dad", "Mom" = "Mom", "Sib" = "Sibling"),
    expand = c(0, 0))

# boxplot_all_hr_no_ia
```

### As Modeled, Moms Only by DR

*Note also need to use actual GRS2 score here.*

```{r}

merge_4l_mom <- merge_4l %>%
  dplyr::filter(fdr_4level == "Mom")

boxplot_mom <- ggplot(merge_4l_mom, aes(x = HLAGRP, y = GRS2, fill = IA)) +
  geom_boxplot(alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "FDR") + 
  theme_bw() +
  # scale_y_continuous(breaks = seq(min, max, by = 2), limits = c(min, max)) +
  stat_summary(fun.data = function(x) data.frame(y = min(merge_4l_mom$GRS2 - 0.2), label = length(x)), 
               geom = "text", position = position_dodge(width = 0.8)) +
  facet_wrap(vars(study)) +
  scale_fill_manual(values = c("#CFB97D", "#4477AA")) +
  scale_x_discrete(labels = c(
    "None" = "None", "Dad" = "Dad", "Mom" = "Mom", "Sib" = "Sibling"),
    expand = c(0,0))

# boxplot_mom

```


### Grouped Fig1 for Poster

1. GRS2 raw by FDR in DAISY & TEDDY without IA
2. GRS2 raw by FDR in controls only
3. IA case/control in DAISY & TEDDY
4. Same as 1. but for non-HLA GRS2

```{r}
# Colors from Paul Tol
muted <- color("muted")
colors <- muted(6)
fdr_colors <- c(
  "None" = colors[1], "Mom" = colors[2], "Dad" = colors[3], "Sib" = colors[4])



boxplot_main <- ggplot(merge_4l, aes(x = fdr_4level, y = GRS2, fill = fdr_4level)) +
  geom_boxplot(
    alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "") + 
  theme_bw() +
  stat_summary(
    fun.data = function(x) data.frame(y = min(merge_4l$GRS2 - 0.2), label = length(x)),
    geom = "text") +
  scale_fill_manual(values = fdr_colors) +
  scale_x_discrete(labels = element_blank()) +
  theme(legend.position = "none") +
  facet_wrap(~ study)

boxplot_main
```

```{r}
# Colors from Paul Tol
muted <- color("muted")
colors <- muted(6)
fdr_colors <- c(
  "None" = colors[1], "Mom" = colors[2], "Dad" = colors[3], "Sib" = colors[4])

boxplot_main_control <- ggplot(
  merge_4l %>% dplyr::filter(IA == 0),
  aes(x = fdr_4level, y = GRS2, fill = fdr_4level)) +
  geom_boxplot(
    alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "") + 
  theme_bw() +
  stat_summary(
    fun.data = function(x) data.frame(y = min(merge_4l$GRS2 - 0.2), label = length(x)),
    geom = "text") +
  scale_fill_manual(values = fdr_colors) +
  scale_x_discrete(labels = element_blank()) +
  theme(legend.position = "none") +
  facet_wrap(~ study)

boxplot_main_control
```

```{r}
boxplot_ia <- ggplot(merge_4l, aes(x = IA, y = GRS2, fill = IA)) +
  geom_boxplot(alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "") +
  theme_bw() +
  stat_summary(fun.data = function(x) data.frame(y = min(merge_4l$GRS2 - 0.2), label = length(x)), 
               geom = "text", position = position_dodge(width = 0.8)) +
  scale_fill_manual(values = c("white", "darkgrey"),
                    labels = c("0" = "Control", "1" = "Case")) +
  scale_x_discrete(labels = element_blank()) +
  theme(legend.position = "bottom") +
  facet_wrap(~ study)

boxplot_ia
```

```{r}
boxplot_main_non_hla <- ggplot(merge_4l_non_hla, aes(x = fdr_4level, y = GRS2, fill = fdr_4level)) +
  geom_boxplot(
    alpha = 0.8, varwidth = TRUE, width = 1) +
  labs(x = "",
       y = "Non-HLA GRS2",
       fill = "FDR") + 
  theme_bw() +
  stat_summary(
    fun.data = function(x) data.frame(y = min(merge_4l_non_hla$GRS2 - 0.2), label = length(x)),
    geom = "text") +
  scale_fill_manual(values = fdr_colors) +
  scale_x_discrete(labels = element_blank()) +
  theme(legend.position = "bottom") +
  facet_wrap(~ study)

boxplot_main_non_hla
```


Combined:

```{r}
main_fig_bottom <- ggarrange(
  boxplot_main_non_hla, boxplot_ia,
  labels = c("C", "D"), nrow=1, widths = c(.5, .5))
main_fig_top <- ggarrange(
  boxplot_main, boxplot_main_control,
  labels = c("A", "B"), nrow=1, widths = c(.5, .5))
main_fig <- ggarrange(
  main_fig_top, main_fig_bottom, nrow = 2, heights = c(0.85, 1))
ggsave(
  "Maternal_Protection/data/results/risk_scores/daisy_teddy_grs2_fdr_4level_for_poster.png",
  main_fig, units = c("in"), width = 6.5, height = 5.2)
```




### Fig2 Modeling Forest Plot for Poster

```{r}
# TODO: revisit this and clean up when have more time!
# Had to re-run model with CIs extracted
excl <- grep("cc|DR|Intercept|IA|T1D", colnames(mod), value = T)
mod_1 <- mod %>%
  dplyr::filter(model == "1" & outcome == "IA") %>%
  dplyr::select(study, starts_with("conf") | starts_with("estimate")) %>%
  dplyr::select(!any_of(excl))
colnames(mod_1) <- gsub("\\.", "_", colnames(mod_1))
colnames(mod_1) <- gsub("_n", "N", colnames(mod_1))
colnames(mod_1) <- gsub("_m", "M", colnames(mod_1))
colnames(mod_1) <- gsub("_s", "S", colnames(mod_1))

mod_1_lg <- pivot_longer(
  mod_1,
  cols = -study,
  names_to = c("measure", "fdr_type"),
  names_pattern = "(.*)_(.*)",
  values_to = "value"
)

mod_1_wd <- pivot_wider(
  mod_1_lg,
  names_from = "measure",
  values_from = "value")
mod_1_wd$fdr_type = factor(mod_1_wd$fdr_type, levels = c("fdrNone", "fdrMom", "fdrSib"))

# forestplot(
#   mean = mod_1_wd$estimate,
#   lower = mod_1_wd$conf_low,
#   upper = mod_1_wd$conf_high,
#   labeltext = paste0(mod_1_wd$study, "_", mod_1_wd$fdr_type))

fdr_colors_mod <- fdr_colors
names(fdr_colors_mod) <- paste0("fdr", names(fdr_colors_mod))

fig2 <- ggplot(mod_1_wd, aes(
  x=study, y=estimate, ymin=conf_low, ymax=conf_high, color=fdr_type, fill=fdr_type)) + 
  geom_linerange(position=position_dodge(width = 0.5), show.legend = F) +
  geom_hline(yintercept=0, lty=2) +
  geom_point(size=3, shape=21, colour="white", stroke = 0.5, position=position_dodge(width = 0.5)) +
  scale_fill_manual(values = fdr_colors_mod,
                    labels = c("fdrSib" = "Sib", "fdrMom" = "Mom", "fdrNone" = "None"))+
  scale_color_manual(values = fdr_colors_mod,
                     labels = c("fdrSib" = "Sib", "fdrMom" = "Mom", "fdrNone" = "None"))+
  coord_flip() +
  theme_bw() +
  labs(
    x = "",
    y = "Change in GRS2",
    fill = "FDR",
  ) +
  scale_x_discrete(labels = c("teddy" = "TEDDY","daisy" = "DAISY")) +
  theme(legend.position = "bottom")

fig2

ggsave(
  "Maternal_Protection/data/results/risk_scores/daisy_teddy_grs2_fdr_4level_forest_plot_for_poster.png",
  fig2, units = c("in"), width = 3, height = 2)
```


### SNPs in GRS2 Weight Chart

Without haplogenotypes since those are scored based on their pairwise
combinations. Positions plotted are hg19.

```{r}
grs2_full_mod <- grs2_full %>%
  dplyr::mutate(present = ifelse(RSID %in% grs2$RSID, "yes", "no")) %>%
  dplyr::mutate(
    POS = as.numeric(gsub("^.+:", "", POSITION_DBSNP151)),
    CHR = as.numeric(gsub(":.+$", "", POSITION_DBSNP151)),
    var_id = factor(POSITION_DBSNP151, levels = unique(POSITION_DBSNP151[order(CHR, POS, decreasing = T)]))) %>%
  dplyr::filter(!str_detect(COMPONENT, "DQ")) %>%  # remove HLA-DQ
  dplyr::mutate(SCORE = as.numeric(SCORE))

fig_wt <- ggplot(grs2_full_mod,
       aes(x = SCORE, y = var_id, fill = present)) +
  geom_col(color = "black") +
  theme_bw() + 
  # geom_text(aes(label = paste0("chr", CHR), x = -1), hjust = 0) +
  labs(
    x = "Weight",
    y = "Variant Position (hg19)",
    fill = "Included"
  ) +
  scale_fill_manual(values = c("white", "darkgrey"),
                    labels = c("no" = "No", "yes" = "Yes")) +
  facet_grid(vars(COMPONENT), scales = "free", space="free") +
  theme(legend.position = "bottom",
        axis.text.y = element_text(size = 6))

fig_wt

ggsave(
  "Maternal_Protection/data/results/risk_scores/daisy_teddy_grs2_weights_for_poster.png",
  fig_wt, units = c("in"), width = 2.5, height = 5)
```


## **Write Out**

```{r}
write_tsv(
  as.data.frame(table1_t),
  "Maternal_Protection/data/results/risk_scores/teddy_table1_ia_with_grs2_sd_eur_dr_filt_fdr_4level.txt")

write_tsv(
  as.data.frame(table1_d),
  "Maternal_Protection/data/results/risk_scores/daisy_table1_ia_with_grs2_sd_eur_dr_filt_fdr_4level.txt")


```


# **Session Info**

```{r session_info}
sessionInfo()
```

