---
title: "6_risk_score_documentation"
author: "SDS"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = paste0(Sys.getenv("RKJCOLLAB"), "/Maternal_Protection/reports")) })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = paste0(Sys.getenv("RKJCOLLAB")))

library(tidyverse)
library(table1)
library(sas7bdat)
library(kableExtra)
library(ggpubr)

```


# **Overview**

Report briefly documents the distributions of the various T1D risk scores
calculated in DAISY and in the Maternal Protection TEDDY data release. Also
includes comparisons between the different versions of GRS2 scores calculated
in DAISY.

**Companion Excel sheet is:**
**Maternal_Protection/documentation/teddy_daisy_risk_score_documentation.xlsx**

Excel sheet includes ALL details on the risk scores used, the input SNPs,
imputation type and quality, etc.

IA was chosen as the outcome to look at score distribution below as it allowed
for the largest sample size.

# **Load Data**

```{r load_data, message = F}

# All risk score results
score_d <- read_delim(
  "DAISY/genetics/risk_scores/data/results/daisy_t1d_genetic_risk_scores.csv")
score_t <- read_delim(
  "Maternal_Protection/data/risk_scores/teddy/teddy_t1d_genetic_risk_scores.csv")

# Pheno
pheno_d <- read.sas7bdat(
  "DAISY/snapshots_of_data/Outputdata08102023_live/casedef_daisy.sas7bdat")
pheno_t <- read_csv(
  "Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_demographic_masked.csv")
ia_def_t <- read_csv(
  "Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_casecntrl_ialist3.csv")

```

# **DAISY**

```{r}
score_pheno_d <- inner_join(
  score_d, pheno_d, by = c("ID")) %>%
  dplyr::mutate(IAcase = as.factor(IAcase))
```


IA case/control status in the `r nrow(score_pheno_d)` individuals scored.
All participants with risk score results also had non-NA IA status.

## **Distributions**

### **AA 7 SNP PRS**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_d, aes(x = AA_PRS)) +
  geom_density()

ggplot(score_pheno_d, aes(x = AA_PRS, fill = IAcase)) +
  geom_density(alpha = 0.5)


table1(~ AA_PRS | IAcase, data = score_pheno_d)
```

### **GRS2 1000G All SNPs**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_d, aes(x = GRS2_1000G)) +
  geom_density()

ggplot(score_pheno_d, aes(x = GRS2_1000G, fill = IAcase)) +
  geom_density(alpha = 0.5)


table1(~ GRS2_1000G | IAcase, data = score_pheno_d)
```

### **GRS2 1000G TEDDY SNPs**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_d, aes(x = GRS2_1000G_TEDDY_SNPS)) +
  geom_density()

ggplot(score_pheno_d, aes(x = GRS2_1000G_TEDDY_SNPS, fill = IAcase)) +
  geom_density(alpha = 0.5)


table1(~ GRS2_1000G_TEDDY_SNPS | IAcase, data = score_pheno_d)
```

### **GRS2 TOPMed R2**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_d, aes(x = GRS2_TM_R2)) +
  geom_density()

ggplot(score_pheno_d, aes(x = GRS2_TM_R2, fill = IAcase)) +
  geom_density(alpha = 0.5)

table1(~ GRS2_TM_R2 | IAcase, data = score_pheno_d)
```


## **Comparisons**

### **GRS2 1000G All SNPs & GRS2 1000G TEDDY SNPs**

Comparison of GRS2 calculated from 1000G with all SNPs to calculation from
just SNPs also available in TEDDY. All SNPs = 67, with TEDDY SNPs = 61. See
more details in Excel document described at top of report.

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_d,
       aes(x = GRS2_1000G, y = GRS2_1000G_TEDDY_SNPS)) +
  geom_point(aes(color = IAcase), alpha = 0.5) +
  xlim(c(1, 20)) +
  ylim(c(1, 20)) +
  stat_cor()
```


### **GRS2 TOPMed R2 & Both GRS2 1000G Versions**

TOPMed R2 GRS2 calculated and used by Henley, compared to GRS2 calculated with
all 1000G SNPs and with subset to match TEDDY 1000G.

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_d,
       aes(x = GRS2_TM_R2, y = GRS2_1000G)) +
  geom_point(aes(color = IAcase), alpha = 0.5) +
  xlim(c(1, 20)) +
  ylim(c(1, 20)) +
  stat_cor()

ggplot(score_pheno_d,
       aes(x = GRS2_TM_R2, y = GRS2_1000G_TEDDY_SNPS)) +
  geom_point(aes(color = IAcase), alpha = 0.5) +
  xlim(c(1, 20)) +
  ylim(c(1, 20)) +
  stat_cor()

```



# **TEDDY**

```{r}
# Prep pheno
# Dedup IA case def
# n_distinct(ia_def_t$Norris_epigenetics_maskid)  # 1621 / 1671
ia_def_t_dup_list <- ia_def_t %>%
  dplyr::filter(duplicated(Norris_epigenetics_maskid))
ia_def_t_dedup <- ia_def_t %>%
  dplyr::filter(
    !(Norris_epigenetics_maskid %in% ia_def_t_dup_list$Norris_epigenetics_maskid &
    outcome == 0))

# Confirm all dups still present have outcome == 1
# check <- ia_def_t_dedup %>%
#   dplyr::filter(Norris_epigenetics_maskid %in% ia_def_t_dup_list$Norris_epigenetics_maskid)

# Merge pheno and IA case def
pheno_t_ia <- full_join(
  pheno_t,
  ia_def_t_dedup,
  by = c("Norris_epigenetics_maskid")) %>%
  dplyr::rename(IAcase = outcome)

# Merge with score, print case/control 
score_pheno_t <- inner_join(
  score_t, pheno_t_ia,
  by = c("ID" = "Norris_epigenetics_maskid")) %>%
  dplyr::mutate(IAcase = as.factor(IAcase))

# Make version filtered to non-NA IA
score_pheno_t_filt <- score_pheno_t %>% dplyr::filter(!is.na(IAcase))
```


`r nrow(score_pheno_t)` individuals were scored, but IA case/control status
was only availabe for `r nrow(score_pheno_t_filt)`, which is the subset of
individuals shown in all following plots.


## **Distributions**

### **AA 7 SNP PRS**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_t, aes(x = AA_PRS)) +
  geom_density()

ggplot(score_pheno_t_filt, aes(x = AA_PRS, fill = IAcase)) +
  geom_density(alpha = 0.5)


table1(~ AA_PRS | IAcase, data = score_pheno_t_filt)
```

### **GRS2 1000G TEDDY SNPs**

```{r out.width = '50%', fig.show = 'hold'}
ggplot(score_pheno_t_filt, aes(x = GRS2_1000G)) +
  geom_density()

ggplot(score_pheno_t_filt, aes(x = GRS2_1000G, fill = IAcase)) +
  geom_density(alpha = 0.5)

table1(~ GRS2_1000G | IAcase, data = score_pheno_t_filt)
```

## **Comparisons**

Not including any plots for TEDDY since no multiple versions of the same score
were generated in TEDDY.


# **Session Info**

```{r session_info}
sessionInfo()
```

