---
title: "Teddy Mat prot GRS2x"
author: "NV"
date: "Last compiled on `r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: kable
    code_folding: hide
knit: (function(inputFile, encoding) {
  source("config.R")
  rmarkdown::render(inputFile, encoding = encoding,
  output_dir = risk_scores_report_dir()) })
---

```{r setup, include=FALSE}
source("config.R")

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(
  root.dir = risk_scores_collab_root()) #setting up pathway

# Load libraries
library(tidyverse)
library(haven)
library(conflicted)
library(dplyr)
library(ggplot2)
library(patchwork)
library(openxlsx)
library(data.table)
library(GGally)
library(ggcorrplot)
```


#**Overview**

Files
Genetics: TEDDY MatProt delivery TOPMed r3 imputation: Maternal_Protection/data/teddy/genetics/tm_r3_imp/imputed_clean_maf0_rsq0.3/chr_all_concat
Code: explore_matprot/risk_scores
You will push this to the RKJcollab explore_matprot GitHub!
Make sure “teddy” is in the file name somewhere since this folder combines TEDDY & DAISY analyses
Results: Maternal_Protection/data/teddy/risk_scores/grs2x/tm_r3_imp
Pheno:  Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_demographic_masked.csv



# **Load Data** 

```{r, message = F}
# .csv file created from log file (GRS2x calculated from PRSedm)
SNPs_logfile <- read.csv("Maternal_Protection/data/teddy/genetics/risk_scores/grs2x/tm_r3_imp/snp_cluster_status_from_logfile.csv")


# Importing .pvar file and comparing for the quality of SNPs used.
SNPs_input_impute <- fread(
  "Maternal_Protection/data/teddy/genetics/imputation/tm_r3_imp/hwe_in_all/imputed_clean_maf0_rsq0.3/chr_all_concat.pvar",
  skip = "#CHROM"
)
# Fix the column name
setnames(SNPs_input_impute, "#CHROM", "CHROM")


# Unfiltered imputation file chr6
info_file <- "Maternal_Protection/data/teddy/genetics/imputation/tm_r3_imp/hwe_in_all/imputed_info/chr6.info.gz" 

info_chr6 <- fread(info_file,
                 skip = "#CHROM")
# Fix the column name
setnames(info_chr6, "#CHROM", "CHROM")


# Importing .pvar file and comparing for the quality of SNPs used in DAISY and TEDDY.
SNPs_daisy <- fread(
  "DAISY/genetics/daisy_ask_genetics/tm_r3_imp/hwe_in_all/imputed_clean_maf0_rsq0.3/chr_all_concat_study_ids.pvar",
  skip = "#CHROM"  # skip comment line to get clean column names
)


# Phenotype file
pheno_raw<- read.csv("Maternal_Protection/raw/TEDDYarrays_v2/additional_data/Norris_demographic_masked.csv")


#GRS2x file without imputation
prs_scores <- read.csv("Maternal_Protection/data/teddy/genetics/risk_scores/grs2x/tm_r3_imp/score_results.csv")

#GRS2x file with imputation
prs_scores_imputed<- read.csv("Maternal_Protection/data/teddy/genetics/risk_scores/grs2x/tm_r3_imp/score_results_imputed.csv")


# GRS2_1000G and AA_PRS 
grs2 <- read.csv("Maternal_Protection/data/teddy/famhist_prs/archive/teddy_t1d_genetic_risk_scores.csv")


```



# **GRS2x Calculation**

```{bash engine.opts= '-i', eval=F}

# mamba install sethsh7::prsedm

# Define input and output paths
in="Maternal_Protection/data/teddy/genetics/imputation/tm_r3_imp/hwe_in_all/imputed_clean_maf0_rsq0.3/chr_all_concat.vcf.gz"
out="Maternal_Protection/data/teddy/genetics/risk_scores/grs2x/tm_r3_imp/score_results"

# Create output directory if it doesn't exist
mkdir -p "$out"

# Run PRSedm
prsedm --vcf "$in" \
       --col GP \
       --build hg38 \
       --scores t1dgrs2-luckett25 \
       --output "$out" > "$out/run.log" 2>&1


```
For genotype column to score (default: GT, options: GT for WGS, GP for imputed data). But GP didn't work even though we have imputed data. So, trying to run GT.


### Rerunnig

```{bash engine.opts= '-i', eval=F}

# mamba install sethsh7::prsedm
# Define the input and output directory
in="Maternal_Protection/data/teddy/genetics/imputation/tm_r3_imp/hwe_in_all/imputed_clean_maf0_rsq0.3/chr_all_concat.vcf.gz"
out="Maternal_Protection/data/teddy/genetics/risk_scores/grs2x/tm_r3_imp/score_results"


# Run PRSedm
prsedm --vcf "$in" \
       --col GT \
       --build hg38 \
       --scores t1dgrs2-luckett25 \
       --output "${out}.csv" > "${out}.log" 2>&1 # > "${out}.log" 2>&1 - to get log file in same output folder
```
created .csv file with SNPs processed using the above log file.


### .csv file created from log file
```{r}
head(SNPs_logfile)
nrow(SNPs_logfile)
sum(SNPs_logfile$Processed=="missing")

missing_snps <- SNPs_logfile %>%
  dplyr::filter(Processed == "missing") %>%
  select(CHR, POS, Cluster, Processed)
print(missing_snps)
```
So, out of 67 SNPs, 3 were neither processed nor imputed.So, checking the quality of these 3 SNPs.

# **SNP Quality Check**

### Comparing with .pvar file from input data for the quality of SNPs used.

```{r}

# column names
names(SNPs_input_impute)


```
```{r}
head(SNPs_input_impute)
```


```{r}
# See if the missing SNPs exist at all in SNPs_input_impute
missing_positions <- c(32644083, 32706288, 32615766)

SNPs_input_impute %>%
  dplyr::filter(POS %in% missing_positions & CHROM == 6)

```
This confirms that all these 3 SNPs are not in the input file

### Quality of these missing SNPs from unfiltered imputation file

```{r}

head(info_chr6)

```
 
```{r}
info_chr6$CHROM <- as.integer(info_chr6$CHROM)

#Join with SNPs_input_impute on POS
missing_snps_info <- missing_snps %>%
  left_join(info_chr6, by = c("POS" = "POS")) %>%
  select(CHR, POS, Cluster, Processed, INFO)

# View result
missing_snps_info
```
### Compare the missing SNPs qualities from TEDDY and DAISY 

```{r}

head(SNPs_daisy)
```


```{r}
# Ensure CHROM is integer to match CHR
SNPs_daisy$CHROM <- as.integer(SNPs_daisy$CHROM)

# Join on POS
combined_snps <- missing_snps_info %>%
  left_join(SNPs_daisy, by = c("POS" = "POS"))

# Then rename the INFO columns
combined_snps <- combined_snps %>%
  rename(INFO_teddy = INFO.x, INFO_daisy = INFO.y) %>%
  select(CHR, POS, Cluster, Processed,REF, ALT, INFO_teddy, INFO_daisy)

# View result
combined_snps
```
The imputation quality of these 3 missing SNPs are better in daisy than teddy.

### Compare the SNPs quality of remaining 64 SNPs which are processed

```{r}
# Join to get Cluster info alongside INFO field
SNP_quality_check <- SNPs_logfile %>%
  left_join(SNPs_input_impute, by = c("CHR" = "CHROM", "POS")) %>%
  distinct(CHR, POS, .keep_all = TRUE) %>%
  dplyr::filter(Processed== "processed")

# View quality info
head(SNP_quality_check[, c("CHR", "POS", "Cluster", "INFO")])

```

```{r}
# Extract MAF and R2 from INFO column using regular expressions
SNP_quality_check <- SNP_quality_check %>%
  mutate(
    MAF_value = as.numeric(str_extract(INFO, "(?<=MAF=)[0-9.]+")),
    R2_value  = as.numeric(str_extract(INFO, "(?<=R2=)[0-9.]+"))
  )

# View summary stats
summary(SNP_quality_check$MAF_value)
summary(SNP_quality_check$R2_value)

# Or get exact min and max
min_MAF <- min(SNP_quality_check$MAF_value, na.rm = TRUE)
max_MAF <- max(SNP_quality_check$MAF_value, na.rm = TRUE)

min_R2 <- min(SNP_quality_check$R2_value, na.rm = TRUE)
max_R2 <- max(SNP_quality_check$R2_value, na.rm = TRUE)

cat("MAF range:", min_MAF, "to", max_MAF, "\n")
cat("R2  range:", min_R2, "to", max_R2, "\n")


```
### Table: SNPs - MAF
```{r}
# Table 1: All SNPs with their CHR, POS, Cluster, and MAF
MAF_table_all <- SNP_quality_check %>%
  select(CHR, POS, Cluster, MAF_value) %>%
  arrange(MAF_value)

# Print full MAF table
print(MAF_table_all)

# Table 2: Top 10 SNPs with lowest MAF
MAF_table_lowest10 <- MAF_table_all %>%
  slice_min(MAF_value, n = 10)

# Print top 10 lowest MAF SNPs
cat("\nTop 10 SNPs with lowest MAF:\n")
print(MAF_table_lowest10)



```

### To asses how many TYPED and IMPUTED SNPs are present
```{r}

# Detect "TYPED" and "IMPUTED" in INFO column
SNP_quality_check$Typed <- grepl("TYPED", SNP_quality_check$INFO)
SNP_quality_check$Imputed <- grepl("IMPUTED", SNP_quality_check$INFO)

# Create a new column for SNP category
SNP_quality_check$SNP_Category <- with(SNP_quality_check, 
  ifelse(Typed & Imputed, "Typed and Imputed",
  ifelse(Typed & !Imputed, "Typed only",
  ifelse(!Typed & Imputed, "Imputed only", "Other"))))

# View summary table with relevant columns
summary_SNP_quality_check <- SNP_quality_check[, c("CHR", "POS", "Cluster", "INFO", "SNP_Category")]

# Print first few rows to check
head(summary_SNP_quality_check)

# Print counts for each category
cat("Number of Typed only SNPs:", sum(SNP_quality_check$SNP_Category == "Typed only"), "\n")
cat("Number of Imputed only SNPs:", sum(SNP_quality_check$SNP_Category == "Imputed only"), "\n")
cat("Number of Typed and Imputed SNPs:", sum(SNP_quality_check$SNP_Category == "Typed and Imputed"), "\n")

```



# **GRS2x calculation (imputed dataset for 3 missing SNPs)**
```{bash engine.opts='-i', eval=F}
# Define input/output paths
in="Maternal_Protection/data/teddy/genetics/imputation/tm_r3_imp/hwe_in_all/imputed_clean_maf0_rsq0.3/chr_all_concat.vcf.gz"
refvcf="Maternal_Protection/raw/genome_maps/TOPMed_freeze8/chr6.TOPMed_freeze_8.all.fixed.vcf.gz" 
out="Maternal_Protection/data/teddy/genetics/risk_scores/grs2x/tm_r3_imp/score_results_imputed"

# Run PRSedm with imputation enabled
prsedm --vcf "$in" \
       --col GT \
       --build hg38 \
       --scores t1dgrs2-luckett25 \
       --impute \
       --refvcf "$refvcf" \
       --output "${out}.csv" > "${out}.log" 2>&1

```
3 SNPs which were missing in previous calculation were imputed. 

Quality of these 3 imputed SNPs- ???

# **Compare GRS2x and GRS2x_imputed across cases and control** 
### Phenotype file

```{r}


head(pheno_raw)
colnames(pheno_raw)

#Subset phenotype file to IDs and t1d status
pheno<- pheno_raw %>%
  select(ID= Norris_epigenetics_maskid, t1d)
  
pheno$ID <- as.character(pheno$ID)

# Label t1d group
pheno <- pheno %>%
  mutate(t1d= ifelse(t1d == 0, "control", "case"))


```
### GRS2x score file (with and without imputation)

```{r}


head(prs_scores)
prs_scores <- prs_scores %>%
  rename(GRS2x_TM_R3 = t1dgrs2.luckett25_total)
head(prs_scores)


head(prs_scores_imputed)
prs_scores_imputed <- prs_scores_imputed %>%
  rename(GRS2x_TM_R3 = t1dgrs2.luckett25_total)
head(prs_scores_imputed)

```

```{r}

prs_scores$ID <- sub(".*_", "", prs_scores$IID)

df <- left_join(prs_scores, pheno, by = "ID")
df <- df %>%
   dplyr::filter(!is.na(GRS2x_TM_R3), !is.na(t1d))
dim(df)         # Check dimensions
colnames(df)    # See available columns

prs_scores_imputed$ID <- sub(".*_", "", prs_scores_imputed$IID)

df_imputed <- left_join(prs_scores_imputed, pheno, by = "ID")
df_imputed <- df_imputed %>%
   dplyr::filter(!is.na(GRS2x_TM_R3), !is.na(t1d))
dim(df_imputed)         # Check dimensions
colnames(df_imputed)    # See available columns


```

### **Visual representation of GRS2x and GRS2x_imputed across cases and control** 


```{r}
# Prepare medians
medians <- df %>%
  group_by(t1d) %>%
  summarise(median_val = median(GRS2x_TM_R3, na.rm = TRUE))

# Plot with ggplot2
ggplot(df, aes(x = factor(t1d), y = GRS2x_TM_R3, fill = factor(t1d))) +
  geom_boxplot() +
  geom_text(data = medians, aes(x = factor(t1d), y = median_val, label = round(median_val, 3)),
            vjust = -0.5, color = "black", size = 3.5) +
  scale_fill_manual(values = c("lightblue", "lightcoral")) +
  labs(title = "GRS2x_tedmp by T1D Case/Control",
       x = "T1D Status",
       y = "GRS2x_TM_R3 Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "none")

# Prepare medians
medians <- df_imputed %>%
  group_by(t1d) %>%
  summarise(median_val = median(GRS2x_TM_R3, na.rm = TRUE))

# Plot with ggplot2
ggplot(df_imputed, aes(x = factor(t1d), y = GRS2x_TM_R3, fill = factor(t1d))) +
  geom_boxplot() +
  geom_text(data = medians, aes(x = factor(t1d), y = median_val, label = round(median_val, 3)),
            vjust = -0.5, color = "black", size = 3.5) +
  scale_fill_manual(values = c("lightblue", "lightcoral")) +
  labs(title = "GRS2x_imputed_tedmp by T1D Case/Control",
       x = "T1D Status",
       y = "GRS2x_TM_R3 Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),
        legend.position = "none")





```


```{r}

ggplot(df, aes(x = GRS2x_TM_R3, fill = t1d)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of GRS2x_tedmp by T1D group",
       x = "GRS2x_TM_R3 Score", y = "Density") +
  theme_minimal()

ggplot(df_imputed, aes(x = GRS2x_TM_R3, fill = t1d)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density of GRS2x_imputed_tedmp by T1D group",
       x = "GRS2x_TM_R3 Score", y = "Density") +
  theme_minimal()


```

```{r}
ggplot(df, aes(x = GRS2x_TM_R3, fill = t1d)) +
  geom_histogram(alpha = 0.6, bins = 20, position = "identity") +
  labs(title = "Histogram of GRS2x_tedmp Scores by T1D group",
       x = "GRS2x_TM_R3 Score", y = "Count") +
  theme_minimal()


ggplot(df_imputed, aes(x = GRS2x_TM_R3, fill = t1d)) +
  geom_histogram(alpha = 0.6, bins = 20, position = "identity") +
  labs(title = "Histogram of GRS2x_imputed_tedmp Scores by T1Dgroup",
       x = "GRS2x_TM_R3 Score", y = "Count") +
  theme_minimal()


```
```{r}
ggplot(df, aes(x = GRS2x_TM_R3, fill = t1d)) +
  geom_histogram(aes(y = after_stat(density)), alpha = 0.6, bins = 20, position = "identity") +
  geom_density(alpha = 0.4) +
  labs(title = "Histogram + Density of GRS2x_tedmp by T1D",
       x = "GRS2x_TM_R3 Score", y = "Density") +
  theme_minimal()

ggplot(df_imputed, aes(x = GRS2x_TM_R3, fill = t1d)) +
  geom_histogram(aes(y = after_stat(density)), alpha = 0.6, bins = 20, position = "identity") +
  geom_density(alpha = 0.4) +
  labs(title = "Histogram + Density of GRS2x_imputed_tedmp by T1D",
       x = "GRS2x_TM_R3 Score", y = "Density") +
  theme_minimal()

```
# **Compare AA_PRS, GRS2_1000G, GRS2x_TM_R3 and GRS2x_TM_R3_imputed **

**GRS2 file**
```{r}

head(grs2)

grs2$ID <- as.character(grs2$ID)

```

### Generate graphs for each GRS2 version


```{r}


# ---------------- DATA PREP ----------------

# Merge grs2 scores into df and df_imputed
merged_df <- df %>% left_join(grs2, by = "ID")
merged_df_imputed <- df_imputed %>% left_join(grs2, by = "ID")

# Filter to keep only complete rows
df_clean <- merged_df %>%
  dplyr::filter(!is.na(t1d), !is.na(GRS2x_TM_R3), !is.na(AA_PRS), !is.na(GRS2_1000G))

df_imputed_clean <- merged_df_imputed %>%
  dplyr::filter(!is.na(t1d), !is.na(GRS2x_TM_R3), !is.na(AA_PRS), !is.na(GRS2_1000G))

# ---------------- FIND GLOBAL LIMITS ----------------

# Boxplot y-axis limits (all scores together)
ymin <- min(c(df_clean$AA_PRS, df_clean$GRS2_1000G,
              merged_df$GRS2x_TM_R3, merged_df_imputed$GRS2x_TM_R3), na.rm = TRUE)
ymax <- max(c(df_clean$AA_PRS, df_clean$GRS2_1000G,
              merged_df$GRS2x_TM_R3, merged_df_imputed$GRS2x_TM_R3), na.rm = TRUE)

# Density x-axis limits
xmin <- min(c(df_clean$AA_PRS, df_clean$GRS2_1000G,
              merged_df$GRS2x_TM_R3, merged_df_imputed$GRS2x_TM_R3), na.rm = TRUE)
xmax <- max(c(df_clean$AA_PRS, df_clean$GRS2_1000G,
              merged_df$GRS2x_TM_R3, merged_df_imputed$GRS2x_TM_R3), na.rm = TRUE)

# ---------------- PLOTTING HELPERS ----------------

# Boxplot helper with fixed y axis
plot_box <- function(df, score, title) {
  ggplot(df, aes(x = t1d, y = {{ score }}, fill = t1d)) +
    geom_boxplot() +
    stat_summary(fun = median, geom = "text", aes(label = round(..y.., 2)),
                 position = position_nudge(y = 0.3), size = 3, color = "black") +
    labs(title = title, x = "T1D status", y = "Score") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = c("skyblue", "coral")) +
    coord_cartesian(ylim = c(ymin, ymax))
}

# Density plot helper with fixed x axis
plot_density <- function(df, score, title) {
  ggplot(df, aes(x = {{ score }}, fill = t1d)) +
    geom_density(alpha = 0.5) +
    labs(title = title, x = "Score", y = "Density", fill = "T1D status") +
    theme_minimal() +
    scale_fill_manual(values = c("skyblue", "coral")) +
    coord_cartesian(xlim = c(xmin, xmax))
}

# ---------------- BOXPLOTS ----------------

# First pair: AA_PRS and GRS2_1000G
box1 <- plot_box(df_clean, AA_PRS, "AA_PRS by T1D group")
box2 <- plot_box(df_clean, GRS2_1000G, "GRS2_1000G by T1D group")

# Second pair: GRS2x observed and imputed
box3 <- plot_box(merged_df, GRS2x_TM_R3, "GRS2x_TM_R3_tedmp (3 missing SNPs)")
box4 <- plot_box(merged_df_imputed, GRS2x_TM_R3, "GRS2x_TM_R3_tedmp (Imputed)")

# Combine boxplots into 2 figures
boxplot_figure1 <- box1 | box2 +
  plot_annotation(title = "Boxplots: AA_PRS and GRS2_1000G",
                  theme = theme(plot.title = element_text(size = 16, face = "bold")))

boxplot_figure2 <- box3 | box4 +
  plot_annotation(title = "Boxplots: GRS2x (3 missing SNPs) vs (Imputed)",
                  theme = theme(plot.title = element_text(size = 16, face = "bold")))

# ---------------- DENSITY PLOTS ----------------

# First pair: AA_PRS and GRS2_1000G
d1 <- plot_density(df_clean, AA_PRS, "AA_PRS by T1D group")
d2 <- plot_density(df_clean, GRS2_1000G, "GRS2_1000G by T1D group")

# Second pair: GRS2x observed and imputed
d3 <- plot_density(merged_df, GRS2x_TM_R3, "GRS2x_TM_R3_tedmp (3 missing SNPs)")
d4 <- plot_density(merged_df_imputed, GRS2x_TM_R3, "GRS2x_TM_R3_tedmp (Imputed)")

# Combine density plots into 2 figures
density_figure1 <- d1 | d2 +
  plot_annotation(title = "Density: AA_PRS and GRS2_1000G",
                  theme = theme(plot.title = element_text(size = 16, face = "bold")))

density_figure2 <- d3 | d4 +
  plot_annotation(title = "Density: GRS2x (3 missing SNPs) vs (Imputed)",
                  theme = theme(plot.title = element_text(size = 16, face = "bold")))

# ---------------- PRINT PLOTS ----------------
print(boxplot_figure1)
print(boxplot_figure2)
print(density_figure1)
print(density_figure2)



```


### Correlation
```{r}

# Create correlation-ready dataframe
corr_df <- df %>%
  select(ID, GRS2x_TM_R3_tedmp_missing = GRS2x_TM_R3) %>%
  left_join(df_imputed %>% select(ID, GRS2x_TM_R3_tedmp_imp = GRS2x_TM_R3), by = "ID") %>%
  left_join(grs2 %>% select(ID, AA_PRS, GRS2_1000G), by = "ID") %>%
  dplyr::filter(complete.cases(.))  # keep only complete rows

# Correlation matrix
cor_matrix <- cor(corr_df %>% select(-ID))

# Plot
ggcorrplot(cor_matrix, 
           method = "circle", 
           type = "lower", 
           lab = TRUE, 
           title = "Correlation among GRS2 Scores",
           colors = c("red", "white", "blue"),
           lab_size = 4,
           show.diag = TRUE)

```
```{r}

# Plot pairwise correlations and scatter plots
ggpairs(corr_df %>% select(-ID),
        title = "Pairwise Correlations and Scatter Plots for GRS2 Scores")
```



# **Session Info**

```{r session_info}
sessionInfo()
```

